### 2019-03-20 Results.Rmd
Add files via upload

---
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Loading enviroment
## Will change to a enviroment specifically containing only results.
load("C:/Users/Martin/transfer-effect-mistriage/code/globalenv.RData")
```

# Results

We analyseed data from `r sum(data.sets$multi.centre.vs.single.centre$multi.centre$.imp == 0) ` trauma patients. The high and low centre samples contains `r sum(data.sets$high.volume.vs.low.volume$high.volume$.imp==0) ` and `r sum(data.sets$high.volume.vs.low.volume$low.volume$.imp==0)` trauma patients respectively (Table 1). The metropolitan sample and the non-metropolitan sample contains `r sum(data.sets$metropolitan.vs.non.metropolitan$metropolitan$.imp==0) ` and `r sum(data.sets$metropolitan.vs.non.metropolitan$non.metropolitan$.imp==0) ` respectively (Table 2). The multi centre sample constitutes the total number of trauma patients. We identified only one valid single centre. This single centre was identified as Karolinska Sjukhuset in Stockholm and contains `r  sum(data.sets$multi.centre.vs.single.centre$single.centre.11001$.imp==0) ` trauma patients (Table 3). The data presented in each table is multiple imputed. The number of imputations used is `r number.of.imputations  `, which is equal to the maximum percentage of missing data found across the data sets.

All samples were similar regarding age, GCS, SBP and RR. The percentage of missing data in the high and low volume sampes were `r NA.info.sample$high.volume.vs.low.volume$high.volume$percent.NA`% and `r NA.info.sample$high.volume.vs.low.volume$low.volume$percent.NA`%. ...

### 2019-04-23 functions/FuncPack.R
Add doParallel and foreach

@@ -8,6 +8,9 @@ FuncPack <- function() {
     library(mice)
     library(Hmisc)
     library(boot)
+    library(bengaltiger)
+    library(doParallel)
+    library(foreach)
   
     ## Source functions
     invisible(lapply(list.files("./Functions", full.names = TRUE, pattern = ".R$"), source))

### 2019-04-23 functions/RunStudy.R
Add test argument

@@ -1,148 +1,142 @@
 #' RunStudy
 #'
-#' Inputs a data frame with the data required to run the study.
-#' Potentially runs with boostraps, augemented by foreach.
-#' Outputs a data frame with medians and IQR.
-#' If bootstraps are used, the function will NOT return values to the Results.env
-#' or pring sample characteristics tables.
-#' If bootstraps are used, the output created will be a row-binded data frame of
-#' all the data frames produced from the bootstraps.
-#' @param data Data used to run study
-#' @param  Logical. If TRUE, the study will run with 1000 bootstraps. Defaults to
-#'     FALSE.
-RunStudy <- function(data, boot = FALSE) {
-  
-  ## Error handling
-  if (!is.list(data))
-    stop ("Input has to be a list")
-  
-  ## BOOTSTRAP
-  ## Apply bootstrap to run
-  if (boot == TRUE) {
-     ## Botstrapping code
-  }
-  
-  ## RUN STUDY
-  ## Select cases with age > 15 or age = NA
-  selected.data <- InclusionSelection(selected.data)
-  ## Data Cleaning
-  selected.data <- DataCleaning(selected.data)
-  
-  ## DATA SETS AND SAMPLES
-  ## Mark entries as High Volume or Low Volume
-  selected.data.vol.mark <- HighVolumeCheck(selected.data)
-  ## Create High Volume Sample and Low Volume Sample
-  High.Volume.Sample <- selected.data.vol.mark[selected.data.vol.mark$High_Volume_Centre == "Yes", ]
-  Low.Volume.Sample <- selected.data.vol.mark[selected.data.vol.mark$High_Volume_Centre == "No", ]
-  rm(selected.data.vol.mark)
-  ## Mark entries as Metropolitan or Non-Metropolitan
-  selected.data.metro.mark <- Metrocheck(selected.data)
-  ## Create Metropolitan Sample and Non-Metropolitan Sample
-  Metropolitan.Sample <- selected.data.metro.mark[selected.data.metro.mark$metropolitan == "Yes", ]
-  Non.Metropolitan.Sample <- selected.data.metro.mark[selected.data.metro.mark$metropolitan == "No", ]
-  rm(selected.data.metro.mark)
-  ## Mark entries as valid (>170 events) individual centres
-  selected.data.ind.mark <- IndividualCentreCheck(selected.data)
-  ## Create Multi Centre Sample (All data)
[…]

### 2019-04-23 functions/FuncPack.R
Add argument to return packages and functions instead of loading

@@ -1,17 +1,31 @@
-#' Loads packages and functions
+#' Loads or return packages and functions
 #' 
-#' No parameters or variables needed to run.
-FuncPack <- function() {
-  
+#' @param return.only Logical. If TRUE packages and functions are not loaded,
+#'     but returned. Defaults to FALSE.
+#' @export
+FuncPack <- function(return.only = FALSE) {
+    
     ## Load required packages
-    library(data.table)
-    library(mice)
-    library(Hmisc)
-    library(boot)
-    library(bengaltiger)
-    library(doParallel)
-    library(foreach)
-  
+    packages <- c("data.table",
+                  "mice",
+                  "Hmisc",
+                  "boot",
+                  "bengaltiger",
+                  "doParallel",
+                  "foreach")
+
+    if (!return.only)
+        for(package in packages) library(package, character.only = TRUE)
+    
     ## Source functions
-    invisible(lapply(list.files("./Functions", full.names = TRUE, pattern = ".R$"), source))
+    files <- list.files("./Functions", full.names = TRUE, pattern = ".R$")
+    functions <- gsub("./Functions/|.R", "", files)
+    if (!return.only)
+        invisible(lapply(files, source))
+
+    ## Return packages and functions
+    return.object <- list(packages = packages,
+                          functions = functions)
+    if (return.only)
+        return(return.object)
 }

### 2019-04-23 functions/FuncPack.R
Fix regex pattern in gsub

@@ -19,7 +19,7 @@ FuncPack <- function(return.only = FALSE) {
     
     ## Source functions
     files <- list.files("./Functions", full.names = TRUE, pattern = ".R$")
-    functions <- gsub("./Functions/|.R", "", files)
+    functions <- gsub("^./Functions/|.R$", "", files)
     if (!return.only)
         invisible(lapply(files, source))
 

### 2019-04-23 functions/RunStudy.R
Remove bootstrap stats

@@ -127,16 +127,6 @@ RunStudy <- function(data, boot = FALSE, test = FALSE) {
         Results$results.data.frames <- results.data.frames
         Results$stats.calculated <- stats.calculated
     }
-    
-    ## If run with bootstraps, return combined outcome
-    if (boot == TRUE) {
-        ## Combine stats obtained by row to new data frme
-        stats.calculated.boot <- rbind(stats.calculated.boot, stats.calculated)
-        ## Output boot specific outcome
-        return(stats.calculated.boot)
-    }
-    
     ## Return output
     return(stats.calculated)
-    
 }

### 2019-04-23 MainCodeRun.R
Add test argument and implement parallel analyses of bootstraping

@@ -2,23 +2,45 @@
 ## 
 ## This file was created using RStudio 1.1.463
 ## Cloned from GitHub 190422 23:00
-MainCodeRun <- function() {
-  
-  ## INITIALIZING
-  ## Load required packages and source functions
-  ## FuncPack() first needs to be sourced to run
-  source("./functions/FuncPack.R")
-  FuncPack()
-  ## Set random seed
-  set.seed(-41892)
-  ## Create Results enviroment
-  Results <- new.env()  
-  ## Import data
-  raw.data <- ImportStudyData(data.file.name =  "simulated-swetrau-data 2.csv", data.path = "C:/Users/Martin/transfer-effect-mistriage/code/data/")
-  ## Create study sample from selected variables
-  selected.data <- VariableSelection(raw.data)
-  
-  ## Run study
-  stats.calculated <- RunStudy(data = selected.data , boot = FALSE)
-
+MainCodeRun <- function(test = FALSE) {
+    
+    ## INITIALIZING
+    ## Load required packages and source functions
+    ## FuncPack() first needs to be sourced to run
+    source("./functions/FuncPack.R")
+    FuncPack()
+    ## Set random seed
+    set.seed(-41892)
+    ## Create Results enviroment
+    Results <- new.env()  
+    ## Import data
+    raw.data <- ImportStudyData("simulated-swetrau-data.csv")
+    ## Create study sample from selected variables
+    selected.data <- VariableSelection(raw.data)
+    ## Run study
+    stats.calculated <- RunStudy(data = selected.data , boot = FALSE, test = test)
+    ## Create bootstrap samples
+    number.of.bootstrap.samples = 1000
+    if (test)
+        number.of.bootstrap.samples = 5
+    bootstrap.samples <- bengaltiger::CreateBootstrapSamples(selected.data,
+                                                             save.to.disk = FALSE,
+                                                             return.samples = TRUE,
+                                                             number.of.bootstrap.samples = number.of.bootstrap.samples)
+    ## Run bootstrap analysis
+    study.cluster <- makeCluster(detectCores())
+    registerDoParallel(study.cluster)
+    bootstrap.results <- foreach(sample = bootstrap.samples,
+                                 .packages = FuncPack(return.only = TRUE)$packages) %dopar%
+        RunStudy(data = selected.data, boot = TRUE, test = test)
+    stopCluster(study.cluster)
+    ## Save results
[…]

### 2019-04-23 MainCodeRun.R
Fix assignment to Results environment

@@ -12,13 +12,14 @@ MainCodeRun <- function(test = FALSE) {
     ## Set random seed
     set.seed(-41892)
     ## Create Results enviroment
-    Results <- new.env()  
+    Results <- new.env(parent = .GlobalEnv)
+    attach(Results)
     ## Import data
     raw.data <- ImportStudyData("simulated-swetrau-data.csv")
     ## Create study sample from selected variables
     selected.data <- VariableSelection(raw.data)
     ## Run study
-    stats.calculated <- RunStudy(data = selected.data , boot = FALSE, test = test)
+    stats.calculated <- RunStudy(selected.data = selected.data , boot = FALSE, test = test)
     ## Create bootstrap samples
     number.of.bootstrap.samples = 1000
     if (test)
@@ -36,6 +37,7 @@ MainCodeRun <- function(test = FALSE) {
     stopCluster(study.cluster)
     ## Save results
     saveRDS(list(original.results = stats.calculated, bootstrap.results = bootstrap.results), "test.results.Rds")
+    saveRDS(Results, "Results.env.Rds")
     message("results saved to disk")
 }
 start.time <- Sys.time()

### 2019-04-23 functions/RunStudy.R

@@ -4,16 +4,16 @@
 #' runs with boostraps, augemented by foreach.  Outputs a data frame with
 #' medians and IQR.  If bootstraps are used, the function will NOT return values
 #' to the Results.env or pring sample characteristics tables.
-#' @param data Data used to run study. No default.
+#' @param selected.data Data used to run study. No default.
 #' @param boot Logical. If TRUE, the study will run with 1000
 #'     bootstraps. Defaults to FALSE.
 #' @param test Logical. If TRUE only multiple imputed datasets are created and 5
 #'     bootstraps are used to estimate the linear shrinkage factor. Passed to
 #'     MICEImplement and DevelopmentModelCreator. Defaults to FALSE.
-RunStudy <- function(data, boot = FALSE, test = FALSE) {
+RunStudy <- function(selected.data, boot = FALSE, test = FALSE) {
     
     ## Error handling
-    if (!is.data.frame(data))
+    if (!is.data.frame(selected.data))
         stop ("Input has to be a data frame")
 
     ## RUN STUDY
@@ -65,9 +65,9 @@ RunStudy <- function(data, boot = FALSE, test = FALSE) {
         ## Extraxt missing data information from each variable in each sample
         NA.info.variable <- lapply(data.sets, function(sample) lapply(sample, NACounterVariable))                 
         ## Save information to Results enviroment
-        Results$data.sets.before.imputations <- data.sets
-        Results$NA.info.sample <- NA.info.sample
-        Results$NA.info.variable <- NA.info.variable
+        assign("data.sets.before.imputations", data.sets, pos = "Results")
+        assign("NA.info.sample", NA.info.sample, pos = "Results")
+        assign("NA.info.variable", NA.info.variable, pos = "Results")
         rm(NA.info.sample, NA.info.variable)
     }
     ## Impute missing data
@@ -124,8 +124,8 @@ RunStudy <- function(data, boot = FALSE, test = FALSE) {
     ## If not a bootstrap run, send results data frames to Results
     if (boot == FALSE) {
         ## Send to results
-        Results$results.data.frames <- results.data.frames
-        Results$stats.calculated <- stats.calculated
+        assign("results.data.frames", results.data.frames, pos = "Results")
+        assign("stats.calculated", stats.calculated, pos = "Results")
     }
     ## Return output
     return(stats.calculated)

### 2019-04-24 MainCodeRun.R
Update MainCodeRun.R

@@ -20,5 +20,8 @@ MainCodeRun <- function() {
   
   ## Run study
   stats.calculated <- RunStudy(data = selected.data , boot = FALSE)
+  
+  ## Save Results enviroment to working directory
+  save(Results, file = "Results.RData")
 
 }

### 2019-04-24 functions/RunStudy.R
Update RunStudy.R

@@ -142,7 +142,9 @@ RunStudy <- function(data, boot = FALSE) {
     return(stats.calculated.boot)
   }
   
-  ## Return output
-  return(stats.calculated)
-  
+  ## If run without bootstraps, return standard outcome
+  if (boot == FALSE) {
+    ## Return standard output
+    return(stats.calculated)
+  }
 }

### 2019-04-24 MainCodeRun.R
Rewrite to get bootstraps to work with saving results locally

@@ -2,46 +2,58 @@
 ## 
 ## This file was created using RStudio 1.1.463
 ## Cloned from GitHub 190422 23:00
-MainCodeRun <- function(test = FALSE) {
+MainCodeRun <- function(test = FALSE, clean.start = TRUE) {
     
     ## INITIALIZING
     ## Load required packages and source functions
     ## FuncPack() first needs to be sourced to run
     source("./functions/FuncPack.R")
     FuncPack()
+    ## Delete output directory if clean.start is TRUE
+    if (clean.start)
+        unlink("output", recursive = TRUE)
     ## Set random seed
     set.seed(-41892)
-    ## Create Results enviroment
-    Results <- new.env(parent = .GlobalEnv)
-    attach(Results)
     ## Import data
     raw.data <- ImportStudyData("simulated-swetrau-data.csv")
     ## Create study sample from selected variables
     selected.data <- VariableSelection(raw.data)
-    ## Run study
-    stats.calculated <- RunStudy(selected.data = selected.data , boot = FALSE, test = test)
+    ## Select cases with age > 15 or age = NA
+    selected.data <- InclusionSelection(selected.data)
+    ## Data Cleaning
+    selected.data <- DataCleaning(selected.data)
+    ## DATA SETS AND SAMPLES
+    ## Mark entries as High Volume or Low Volume
+    selected.data <- HighVolumeCheck(selected.data)
+    ## Mark entries as Metropolitan or Non-Metropolitan
+    selected.data <- Metrocheck(selected.data)
+    ## Mark entries as valid (>170 events) individual centres
+    selected.data <- IndividualCentreCheck(selected.data)
+    
+    ## Get original results
+    RunStudy(selected.data = selected.data , boot = FALSE, test = test)
     ## Create bootstrap samples
     number.of.bootstrap.samples = 1000
     if (test)
         number.of.bootstrap.samples = 5
     bootstrap.samples <- bengaltiger::CreateBootstrapSamples(selected.data,
-                                                             save.to.disk = FALSE,
+                                                             strata = "res_survival",
+                                                             save.to.disk = TRUE,
                                                              return.samples = TRUE,
                                                              number.of.bootstrap.samples = number.of.bootstrap.samples)
-    ## Run bootstrap analysis
+        ## Get bootstrap results
     study.cluster <- makeCluster(detectCores())
     registerDoParallel(study.cluster)
[…]

### 2019-04-24 functions/RunStudy.R

@@ -11,40 +11,27 @@
 #'     bootstraps are used to estimate the linear shrinkage factor. Passed to
 #'     MICEImplement and DevelopmentModelCreator. Defaults to FALSE.
 RunStudy <- function(selected.data, boot = FALSE, test = FALSE) {
-    
+
     ## Error handling
     if (!is.data.frame(selected.data))
         stop ("Input has to be a data frame")
+    ## Create results list
+    Results <- list()
 
     ## RUN STUDY
-    ## Select cases with age > 15 or age = NA
-    selected.data <- InclusionSelection(selected.data)
-    ## Data Cleaning
-    selected.data <- DataCleaning(selected.data)
-    
-    ## DATA SETS AND SAMPLES
-    ## Mark entries as High Volume or Low Volume
-    selected.data.vol.mark <- HighVolumeCheck(selected.data)
     ## Create High Volume Sample and Low Volume Sample
-    High.Volume.Sample <- selected.data.vol.mark[selected.data.vol.mark$High_Volume_Centre == "Yes", ]
-    Low.Volume.Sample <- selected.data.vol.mark[selected.data.vol.mark$High_Volume_Centre == "No", ]
-    rm(selected.data.vol.mark)
-    ## Mark entries as Metropolitan or Non-Metropolitan
-    selected.data.metro.mark <- Metrocheck(selected.data)
+    High.Volume.Sample <- selected.data[selected.data$High_Volume_Centre == "Yes", ]
+    Low.Volume.Sample <- selected.data[selected.data$High_Volume_Centre == "No", ]
     ## Create Metropolitan Sample and Non-Metropolitan Sample
-    Metropolitan.Sample <- selected.data.metro.mark[selected.data.metro.mark$metropolitan == "Yes", ]
-    Non.Metropolitan.Sample <- selected.data.metro.mark[selected.data.metro.mark$metropolitan == "No", ]
-    rm(selected.data.metro.mark)
-    ## Mark entries as valid (>170 events) individual centres
-    selected.data.ind.mark <- IndividualCentreCheck(selected.data)
+    Metropolitan.Sample <- selected.data[selected.data$metropolitan == "Yes", ]
+    Non.Metropolitan.Sample <- selected.data[selected.data$metropolitan == "No", ]
     ## Create Multi Centre Sample (All data)
-    Multi.Centre.Sample <- selected.data.ind.mark
+    Multi.Centre.Sample <- selected.data
[…]

### 2019-04-24 functions/RunStudy.R
Update RunStudy.R

@@ -42,9 +42,7 @@ RunStudy <- function(selected.data, boot = FALSE, test = FALSE) {
                       multi.centre.vs.single.centre = c(list(multi.centre = Multi.Centre.Sample),
                                                         Single.Centre.Samples))
     
-    ## RESTRICTED CUBIC SPLINES AND MISSING DATA
-    ## Create restricted cubic splines
-    data.sets <- lapply(data.sets, function(sample) lapply(sample, RCSplineConvert))
+    ## RESTRICTED CUBIC SPLINES AND MISSING DATA 
     ## If not a bootstrap run, obtain NA-data and send to results
     if (boot == FALSE) {
         ## Extract missing data information from each sample
@@ -57,6 +55,9 @@ RunStudy <- function(selected.data, boot = FALSE, test = FALSE) {
         Results$NA.info.variable <- NA.info.variable
         rm(NA.info.sample, NA.info.variable)
     }
+    ## Create restricted cubic splines
+    data.sets <- lapply(data.sets, function(sample) lapply(sample, RCSplineConvert))
+
     ## Impute missing data
     data.sets <- lapply(data.sets, MICEImplement, test = test)
     ## If not a bootstrap run, create table one

### 2019-04-26 Old stuff/Results.Rmd
Rename Results.Rmd to Old stuff/Results.Rmd

--- 2019-03-20 Results.Rmd
@@
 We identified only one valid single centre.
-This single centre was identified as Karolinska Sjukhuset in Stockholm and contains `r  sum(data.sets$multi.centre.vs.single.centre$single.centre.11001$.imp==0) ` trauma patients (Table 3).
+which contains `r  sum(data.sets$multi.centre.vs.single.centre$single.centre.11001$.imp==0) ` trauma patients (Table 3).
 The data presented in each table is multiple imputed.

### 2019-04-26 Results.Rmd
Add files via upload

@@
 knitr::opts_chunk$set(echo = TRUE)
+## LOADING ENVIROMENT
+results <- readRDS("C:/Users/Martin/transfer-effect-mistriage/code/output/original.results.Rds")
+## PREPARED CODE SNIPPETS
+## Total patients and NA
+total.number.of.patients <- nrow(results$data.sets.before.imputations$multi.centre.vs.single.centre$multi.centre)
+total.number.of.na <- results$NA.info.sample$multi.centre.vs.single.centre$multi.centre$total.number.of.na
+## Sample NA
+na.vector <- unlist(results$NA.info.sample)
+highest.percent.missing.number <- max(na.vector[na.vector<100])
+highest.percent.missing.name <- names((na.vector[na.vector == highest.percent.missing.number]))
+## Variable NA
+number.variable.na <- max(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na)
+percent.variable.na <- results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$percentage[results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na == number.variable.na]
+name.variable.na <- row.names( results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na == number.variable.na,] )
+## Development and validation
+max.imp <- max(results$original.stats$Total.imputations)
+best.validation <- results$original.stats[results$original.stats$Validation.mistriage.median == min(results$original.stats$Validation.mistriage.median),]
+## Comparison
+best.transfer <- results$original.stats[results$original.stats$Transfer.mistriage.median == min(results$original.stats$Transfer.mistriage.median),]
+worst.transfer <- results$original.stats[results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median == min(min(results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median)),]
 
-## Loading enviroment
-## Will change to a enviroment specifically containing only results.
-load("C:/Users/Martin/transfer-effect-mistriage/code/globalenv.RData")
 ```
@@
 # Results
+We analysed data from `r total.number.of.patients ` trauma patients to investigate the effects of transfer on clinical prediction models (Table 1).
+The total number of missing observations was `r total.number.of.na ` in the entire study cohort.
+The sample with the highest percentage of missing observations was the `r highest.percent.missing.name ` sample with `r highest.percent.missing.number `%.
+The variable with the highest number of missing observations was `r name.variable.na `, with `r number.variable.na ` missing observations, which constituted `r percent.variable.na ` % of the total observations for this variable.
 
-We analyseed data from `r sum(data.sets$multi.centre.vs.single.centre$multi.centre$.imp == 0) ` trauma patients.
-The high and low centre samples contains `r sum(data.sets$high.volume.vs.low.volume$high.volume$.imp==0) ` and `r sum(data.sets$high.volume.vs.low.volume$low.volume$.imp==0)` trauma patients respectively (Table 1).
-The metropolitan sample and the non-metropolitan sample contains `r sum(data.sets$metropolitan.vs.non.metropolitan$metropolitan$.imp==0) ` and `r sum(data.sets$metropolitan.vs.non.metropolitan$non.metropolitan$.imp==0) ` respectively (Table 2).
-The multi centre sample constitutes the total number of trauma patients.
-We identified only one valid single centre.
-This single centre was identified as Karolinska Sjukhuset in Stockholm and contains `r  sum(data.sets$multi.centre.vs.single.centre$single.centre.11001$.imp==0) ` trauma patients (Table 3).
-The data presented in each table is multiple imputed.
-The number of imputations used is `r number.of.imputations  `, which is equal to the maximum percentage of missing data found across the data sets.
+Variable outliers?  
+Divergent samples or variables.  
+NA for predictors and outcome??  
+Table 1 = Sample Characteristics table
 
-All samples were similar regarding age, GCS, SBP and RR.
-The percentage of missing data in the high and low volume sampes were `r NA.info.sample$high.volume.vs.low.volume$high.volume$percent.NA`% and `r NA.info.sample$high.volume.vs.low.volume$low.volume$percent.NA`%.
-...
+## Development and validation
+During model development, the maximum amount of imputations used was `r max.imp`, which is equal to the highest percentage of missing observations.
+Restricted dubic splines were implemented when modelling sysytolic blood pressure and respiratory rate, in both cases 4 knots were used.
+Model validation performance was obtained for each model (Table 2).
+The sample with the best validation model performance was the `r best.validation$Sample.name` sample with a median mistriage rate of `r best.validation$Validation.mistriage.median` (IQR `r best.validation$Validation.mistriage.lb` - `r best.validation$Validation.mistriage.ub`).
+Model validation performance in general was good/poor with a median mistriage rate range of `r min(results$original.stats$Validation.mistriage.median)` - `r max(results$original.stats$Validation.mistriage.median)`.
+
+Imputations  
+Shrinkage factor?  
+Optimal cut-off?  
+RCS for different variables  
+Bootstraps?  
+Local mistriage values.  
+Table 2 = Medians and IQR for validation
+
+## Comparison
+Model performance after transfer was determined in each data set (Table 3).
+When transferred, the model with the best performance was `r best.transfer$Sample.name` with a median mistriage of `r best.transfer$Transfer.mistriage.median` (IQR `r best.transfer$Transfer.mistriage.lb` - `r best.transfer$Transfer.mistriage.ub`).
+When transferred model performance was compared to the validation performance in the same sample, the transferred models generally performed worse/better/same than their corresponding local models (Table 4).
+The highest decline in performance was found in the `r worst.transfer$Sample.name` sample, with a median mistriage difference of `r worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median` (IQR `r worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.lb` - `r worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.ub`), meaning that this model was the most negativly affected by being transferred.
+In clinical terms, this model transfer means that among 100 trauma patients, `r abs(worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median) * 100` would be wrongly classified as major or minor trauma.
+
+Comparison model performances.  
+Most? Less?   
+Significance! P-values.  
+Table 3 = Medians and IQR for transfer.  
+Table 4 = Medians and IQR for the difference in performance

### 2019-04-27 Results.Rmd
Add line to import result tables and print one

@@
 ## LOADING ENVIROMENT
-results <- readRDS("C:/Users/Martin/transfer-effect-mistriage/code/output/original.results.Rds")
+results <- readRDS("./output/original.results.Rds")
 ## PREPARED CODE SNIPPETS
@@
 worst.transfer <- results$original.stats[results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median == min(min(results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median)),]
+## Result tables
+result.tables <- readRDS("result.tables.Rds")
 
@@
 Table 4 = Medians and IQR for the difference in performance
+
+```{r, echo = FALSE, results = "asis"} 
+result.tables$high.volume
+```

### 2019-05-17 ResultsMarkdown.Rmd
Add files via upload

---
title: "Mistriage"
author: "Martin Henriksson"
date: "190429"
output: word_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
## LOADING ENVIROMENT AND TABLES
results <- readRDS("compiled output/original.results.Rds")
result.tables <- readRDS("result.tables.Rds")
# table.one <-

## PREPARED CODE
## Total patients and NA
total.number.of.patients <- nrow(results$data.sets.before.imputations$multi.centre.vs.single.centre$multi.centre)
total.number.of.na <- results$NA.info.sample$multi.centre.vs.single.centre$multi.centre$total.number.of.na
## Sample NA
na.vector <- unlist(results$NA.info.sample)
highest.percent.missing.number <- max(na.vector[na.vector<100])
highest.percent.missing.name <- names((na.vector[na.vector == highest.percent.missing.number]))
## Variable NA
number.variable.na <- max(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na)
percent.variable.na <- results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$percentage[results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na == number.variable.na]
name.variable.na <- row.names( results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na == number.variable.na,] )
gcs.na <- (results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[row.names(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre) == "ed_gcs_sum",])
sbp.na <- (results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[row.names(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre) == "ed_sbp_value",])
## Development and validation
high.low.imp <- results$original.stats$Total.imputations[results$original.stats$Sample.name == "high.volume"]
metro.non.imp <- results$original.stats$Total.imputations[results$original.stats$Sample.name == "metropolitan"]
multi.single.imp <- results$original.stats$Total.imputations[results$original.stats$Sample.name == "multi.centre"]
max.imp <- max(results$original.stats$Total.imputations)
best.validation <- results$original.stats[results$original.stats$Validation.mistriage.median == min(results$original.stats$Validation.mistriage.median),]
## Comparison
best.transfer <- results$original.stats[results$original.stats$Transfer.mistriage.median == min(results$original.stats$Transfer.mistriage.median),]
worst.transfer <- results$original.stats[results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median == max(results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median),]
## Figure and table setup
# Figure 1 = Figure 1: High- and low volume centre data set. Initial split based on number of patients. Temporal split made using date of traumatic event.
# Table 1 = Sample characteristics table
# Table 2 = result.tables, cropped to only show validation
# Table 3 = Cropped to only show transfer
# Table 4 = Cropped to only show transfer differance
```
# Introduction
Trauma results in millions of deaths annually across the globe, and road traffic injuries are among the top 10 leading causes of death worldwide (1, 2). Each year 9% of the world’s deaths are because of trauma, and predictions indicate that this number is likely to increase (3).

Trauma care involves many levels of the health care system. In a typical, high resource setting, the initial management of patients is performed on the scene of the trauma by Emergency Medical Services (EMS). In this prehospital phase; patient data, vital signs and scene information is collected and transferred to the receiving hospital. This information is then evaluated with the help of a system to determine the level of trauma and prepare adequate resources (4). The level of trauma dictates if a full or limited trauma team is activated (5).

Systems to determine the level of trauma can be based on clinical prediction models. Such models are continuously being developed and studied, examining different parameters and different contexts concerning trauma care. These models differ in quality and characteristics, but they generally perform well when predicting survival (6). Many of these models are developed in a single standardized setting such as a large metropolitan area or a major trauma centre but are then supposed to be implemented in different settings, such as non-trauma centres.

What is not fully understood is how this transfer affects the performance of these models. Previous research that studied the transfer of clinical prediction models between settings has shown that model performance in terms of calibration can be adversely affected (7). However, the aforementioned research studied model transfers between substantially different settings and did not assess more clinically relevant performance measures, such as misclassification.

In trauma, misclassification is often referred to as mistriage. Here, triage refers to the classification of severity as minor or major trauma. Mistriage can be subdivided into overtriage; the incorrect classification of a patient with minor trauma as major trauma, or undertriage; the incorrect classification of a patient with major trauma as minor trauma. Both are detrimental in terms of patient care and distribution of resources.

Thus, the effect of model transfers between contexts within the setting of a single health care system, as well as the effect of such transfers on mistriage have not been studied and represent substantial knowledge gaps. Therefore, the aim of this study is to assess how transfers of clinical prediction models for early trauma care between different contexts within a single health system affect mistriage rates.

# Materials and methods
## Study design
We conducted a registry-based cohort study. Sweden has a nationally encompassing register for registering trauma called SweTrau. We used SweTrau data to create clinical prediction models, which we then transferred between different contexts to study the effects on mistriage. Please refer to appendix I for a brief explanatory list of the statistical terminology.

## Setting
In Sweden, trauma is the most common cause of death before 44 years of age. It is estimated that up to 10% of all major trauma patients suffer some form of disability. The societal effects of trauma are also substantial; with the loss of work days comparable to those lost from cardiovascular and malignant diseases combined (5).

The trauma care organization in Sweden is usually clearly defined in major metropolitan areas (albeit the specific routines differ), and trauma patients are transported directly to a predesignated hospital with specific trauma competency and capacity. In large, more rural parts of the country, trauma patients are instead transported to the nearest hospital for stabilization. Once stabilized, they may then be transported to hospitals with more advanced trauma care capacity (19). Most hospitals in Sweden use some form of system to categorize the level of trauma and necessary response. These systems are usually based on the patient's vital signs and injury mechanism, as reported by the EMS before patient arrival, or as registered in the emergency department if the patient arrives by other means (16).

We used data from SweTrau between 2011 and 2016, as this study’s inception was in 2016 and data obtained the same year. Today 95,5% (52 of 55) of Swedish hospitals record trauma cases in SweTrau. Currently the registry includes 55 000 cases, and the board encourages its use in both academic research and more local quality improvement initiatives (20). Included hospitals are both university and non-university hospitals, all of which receive trauma patients.

Patient data is recorded upon arrival to the hospital. All hospital personnel can perform patient registration, but it is recommended they received Abbreviated Injury Scale (AIS) training. The SweTrau inclusion criteria are: Traumatic event with subsequent activation of hospital trauma protocol, admitted patients with NISS (New Injury Severity Score) > 15 and patients transferred to the hospital within 7 days of traumatic event with NISS > 15. SweTrau excludes patients if the only traumatic event is chronic subdural hematoma or if hospital trauma protocol is activated without traumatic event (21).

## Participants 
The eligibility criteria are patients registered in SweTrau, and age 15 or above. This age was decided as the study aims to study adult trauma and not paediatric trauma which differs in physiology, triage and initial care (22). We recognize this age cut-off is not completely without controversy. However, in Sweden patients age ≥15 go to the adult emergency rooms and the Swedish guidelines for trauma activation define children as age <15 (16). Also, several guidelines and protocols listed in ATLS use age ≥15 as cut-off (9). Patient age was obtained from the SweTrau register.

## Variables
### Model predictors
Our clinical prediction models included the predictors SBP, RR and GCS on arrival to hospital. The rationale for including these three predictors is that they are part of many established clinical prediction models for early trauma care, such as the Revised Trauma Score (23).

### Model outcome
The outcome that we used to develop the clinical prediction models was all cause mortality within 30 days of the trauma.

### Participant characteristics
To describe the patient cohort we reported age, sex, ASA physical status classification system, Injury Severity Score (ISS) and New ISS (NISS).

### Study outcome
We used ISS > 15 as the gold standard to define trauma severity as major trauma, and hence patients with ISS ≤ 15 were considered minor trauma (24). We defined overtriage as the event when a clinical prediction model classified a patient with ISS ≤ 15 as major trauma, and undertriage as the event when a clinical prediction model classified a patient with ISS > 15 as minor trauma. Clinically, in the event of overtriage, a full trauma team is activated and tasked with the care of a minimally injured patient. The opposite is true with undertriage, where a patient with serious injuries is managed with (potentially) inadequate resources  (10). We defined the overtriage rate as the number of overtriaged patients divided by all patients. We defined the undertriage rate as the number of undertriaged patients divided by all patients. The mistriage rate was defined as the sum of the over- and undertriage rates.

## Data sources and measurements
Model predictors, outcome, participant characteristics and study outcome as outlined above will all be obtained from the SweTrau registry. The method of measurement for the model predictors is not specified (for example, if SBP is measured using an automated cuff or manually) in the registry entries. In Swedish emergency rooms, patient parameters are usually obtained by a registered nurse or assistant nurse and are assumed to be accurate. 
Whether the patient is dead or alive 30 days after trauma is manually entered into the registry locally, by each respective hospital. Foreign citizens discharged within 30 days are registered as survivors. Participant characteristics; sex and age are obtained from the patient file before being registered in SweTrau. ASA, NISS and ISS are calculated by hospital personal based on the patient’s injuries, and registered.

## Bias
Data analysis was conducted in a step-by-step fashion, and according to a prearranged analysis plan. The analysis plan and statistical analysis code was finalised using simulated data instead of the real data. These efforts were taken to avoid confirmation bias when conducting the data analysis. The analysis plan was reviewed by an experienced statistician and programmer prior to implementation to ensure objectivity. Neither outcome nor variable were blinded when conducting data analysis, which made a structured objective approach even more important.

## Study size
All patients matching eligibility criteria are listed above. Four data sets were used to study the transfer of clinical prediction models, each data set representing a different setting. Data set, and sample size considerations are outlined below.

## Quantitative variables
SBP and RR was modelled using restricted cubic splines with four knots whenever possible, placed at equally spaced percentiles and GCS as a continuous linear term. When describing the participant characteristics all quantitative variables were to be presented as continuous. ISS was also presented as dichotomous using ISS > 15 as the cutoff.

## Statistical methods
### Data sets
The complete SweTrau cohort was split into four sets of data. High and low volume centres, metropolitan and non-metropolitan centres, multi and single centre data and individual centres.

### High and low volume centres
Based on number of patients, two samples were derived from this data set. High volume centres were those with in the top quartile of number of patients received. The rest were low volume centres.

### Metropolitan and non-metropolitan centres
This data set was also split into two samples. The metropolitan sample consisted of greater Stockholm, greater Gothenburg and greater Malmö, as defined by statistics Sweden. The other sample was patients from non-metropolitan areas.

### Multi and single centre data
In this data set multiple samples were created. Each centre with large enough sample size to develop and validate a model were their own sample. The multi-centre sample consisted of the combined data from all centres.

### Indicidual centres
This data set was also be split into multiple samples. Each centre with large enough sample size to develop and validate a model constituted its own sample. This data set was removed in the final version of the study due to a lack of centres with a large enough sample size.

### Development and validation sample
Each data set included at least two samples. The samples were then split into two subsamples using a temporal split based on the date of traumatic event. The earlier subsample was the development sample, and the later subsample the validation sample. The development sample contained 70 events (events being patients who died within 30 days of the trauma) and all non-events (non-events being patient survival 30 days past the trauma) during the same time. The rationale for including 70 events is that we needed at least 10 events per free parameter in the logistic regression to obtain stable coefficient estimates (25). The validation sample instead contained 100 events and at least 100 non-events, which was suggested as the minimum number by Vergouwe in 2005 for external validation samples (26). See figure 1 for example using the high- and low volume centre data set, with high volume centres being those in the top quartile of number of patients registered.

```{r, echo = FALSE, results = "asis"} 
## Figure 1 = Figure 1: High- and low volume centre data set. Initial split based on number of patients. Temporal split made using date of traumatic event. 
```

The minimum sample size of development and validation samples was 140 and 200 respectively. The total number of events and non-events per data set was therefore at least 680. We performed analyses only on data sets for which all samples included at least the minimum number of patients.

### Sequence of analysis
The programming language R was used for all analyses (27), and all progress was uploaded to a GitHub repository for inspection (28). We performed the analyses in the sequence of model development, model validation and finally model comparison. These steps were repeated in each data set. Below we used the transfer of a model from a high-volume sample to a low volume sample as an example to describe the complete procedure.

### Model development
In the model development step a clinical prediction model was developed in the high-volume centre development sample. The model was developed using logistic regression as implemented in the R function glm. The dependent variable was all cause mortality within 30 days of trauma and independent variables SBP, RR, and GCS modelled as previously described. To avoid overfitting the model we used a bootstrap procedure to estimate a linear shrinkage factor that we then applied to the model coefficients (29). The shrunk model was then used to estimate the probability of all cause 30-day mortality in the development sample. We then performed a gridsearch across estimated probabilities in the development sample to identify the cut-off that optimised overtriage keeping undertriage at less than 5% (11). This cut-off was then used to classify patients as major or minor trauma.

### Model validation
In the model validation step, the model performance was assessed in the high-volume centre validation sample and in the low volume centre validation data. First the model was used to estimate the probability of all cause 30-day mortality in each sample. Then the probability cutoff identified in the development sample was applied to the validation samples, patients were classified as major or minor trauma, and model performance was estimated.

### Model comparison
Finally, in the model comparison step, the difference in model performance between the high and low volume centre validation samples was calculated. We used an empirical bootstrap to estimate 95% confidence intervals (CI) around performance and differences in performance estimates. Both bootstrap procedures used 1000 bootstrap samples drawn with replacement of the same size as the original samples. The three steps of model development, model validation, and model comparison were repeated in all four sets of data.

### Performance measures
Model performance was assessed in terms of over-, under-, and mistriage rates as defined above.

### Missing data
We used multiple imputation using chained equations, as implemented in the R package mice, to handle missing data (30). The number of imputations created for each data set was equal to the highest percentage of missing data in that data set. Quantitative variables were imputed using predictive mean matching and qualitative variables were imputed using logistic regression. SBP and RR was transformed as restricted cubic splines before imputation and imputed as just another variable. All analyses outlined above were then conducted separately in each imputed dataset. We present the combined results as the median point estimate across imputations along with an empirical bootstrap of the 25th and 75th percentiles across imputations, i.e. the lower bound of the presented interval is the lower bound of a 95% CI of the 25th percentile and the upper bound is the upper bound of a 95% CI of the 75th percentile. This combined CI was referred to as an Uncertainty Interval (UI) and was used to express the added uncertainty associated with the imputation procedure to handle missing data, as such it is more conservative than a standard 95% CI.

## Ethical considerations
This study was conducted in accordance with the four major principles of medical ethics and The Declaration of Helsinki (31, 32).

### Respect for autonomy
The respect for autonomy was upheld as Swedish Privacy Law requires the informed consent of all participants. Patients included in SweTrau receive letters detailing that data obtained may be used for scientific purposes. They are also informed that participation is voluntary and that they may withdraw at any time.

### The Principle of Beneficence
Throughout the entire study we attempted to act in the best interest of the patients. In the long run we hope this research will improve the care of trauma patients, and therefore prove beneficial to those affected. We also believe that the patients who consented to research want to contribute to this improvement of care, and as thus by performing this study we hope to honour their wishes.

### The Principle of Nonmaleficence
All attempts were made not to harm any patients included in this study, either intentionally or unintentionally.  Data leakage and patient identification constitutes the primary risk. All data obtain was depersonalized and handled with care to minimize risk of unauthorised access and subsequent patient identification. By implementing these actions, we determined that the risk to the patient population is minimal.

### The Principle of Justice
As this study contained no intervention, the principle of justice does not truly apply. However, all cases included from the registry were treated as equal. It is also our wish that any scientific gains this study may contribute will be implemented in a fair and equal fashion.

### Ethical permit
This study is approved by the regional ethics review board in Stockholm, Sweden. Ethical review numbers are 2015/426-31 and 2016/461-32.

# Results
We analysed data from `r total.number.of.patients ` trauma patients to investigate the effects of transfer on clinical prediction models (Table 1). The total number of missing observations across all variables was `r total.number.of.na ` in the entire study cohort. The sample with the highest percentage of missing observations was the `r highest.percent.missing.name ` sample with `r highest.percent.missing.number `% incomplete observations. The variable with the highest number of missing values was `r name.variable.na `, with `r number.variable.na ` missing values, which constituted `r round(percent.variable.na) `% of the total values for this variable. The percentages of missing values for the other model predictors, GCS and SBP was `r round(gcs.na$percentage)`% and `r round(sbp.na$percentage)`% respectively.

```{r, echo = FALSE, results = "asis"} 
## Table 1 = Sample characteristics table
##result.tables
```

## Development and validation
During model development, the number of imputations used for each data set was `r high.low.imp`, `r metro.non.imp` and `r multi.single.imp` for the high and low volume data set, the metropolitan and non-metropolitan data set and the multi and single centre data set respectively. This number was equal to the maximum percentage of missing observations for each sample in the respective data sets. Restricted cubic splines with four knots placed at equally spaced percentiles were implemented when modelling systolic blood pressure and respiratory rate. 1000 bootstrap samples drawn with replacement was used to estimate a linear shrinkage factor that was applied to the developed models to avoid model overfitting.

Table 2 shows model validation performance. Model validation performance refers to the model performance when applied in the validation subsample from the sample in which it was developed. The sample with the best validation model performance was the `r best.validation$Sample.name` sample with a median mistriage rate of `r best.validation$Validation.mistriage.median`. Performance in terms of under and overtriage for the same model was `r best.validation$Validation.undertriage.median`, and `r best.validation$Validation.overtriage.median`. Model validation performance varied across all samples with a median mistriage rate range of `r min(results$original.stats$Validation.mistriage.median)` - `r max(results$original.stats$Validation.mistriage.median)`. Uncertainty interval(UI) for validation mistriage, undertriage and overtriage ranged from 0 to 0.830 across all samples.

```{r, echo = FALSE, results = "asis"} 
## Table 2 = Cropped to only show validation
result.tables
```

## Comparison
Model performance after transfer was determined for each model being transferred to the other sample in the data set in all data sets (Table 3). When transferred, the model with the best performance i.e. lowest mistriage rate was the `r best.transfer$Sample.name` model with a median mistriage of `r best.transfer$Transfer.mistriage.median`. Concerning the UI for transfer mistriage, undertriage and overtriage, this too ranged from 0 to 0.830 across all samples.

```{r, echo = FALSE, results = "asis"} 
## Table 3 = Cropped to only show transfer
result.tables
```

The main purpose of this study was to examine the transferability of clinical prediction models for trauma. Table 4 shows the transferred model performance when compared to the validation performance in the same sample, transferred model performance varied between samples and data sets.

The lowest model performance when compared to the sample validation performance was found in the `r worst.transfer$Sample.name` sample, with a median mistriage difference of `r worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median`, meaning this model performed worse when transferred to the other sample in the same data set, when compared with the model developed and validated in this sample. In clinical terms, this model transfer means that among 100 trauma patients, `r abs(worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median) * 100` more patients would be wrongly classified as major or minor trauma. However, when uncertainty intervals (UI) were obtained no observed differences between transferred model performances and validation model performances in any sample were found to be significant.

```{r, echo = FALSE, results = "asis"} 
## Table 4 = Cropped to only show transfer differance
result.tables
```

# Discussion (Needs update with code snippets) 
This study aimed to assess how transfers of clinical prediction models for early trauma care between different contexts within a single health system affect mistriage rates. The most notable effects on model performance following model transfer compared to the original sample model was seen when the metropolitan model was transferred to the non-metropolitan sample. The performance of the metropolitan model transferred to the non-metropolitan sample was a mistriage rate of 0,51, and the validation mistriage rate of the non-metropolitan model was 0,20. This indicates that the mistriage increased by 0,31 when using the metropolitan model in a non-metropolitan setting. Mainly contributing to this increase in mistriage was the marked increase in overtriage when performing this model transfer. The UI of this difference however ranged from -0.055 to 0.630, indicating that our findings are compatible with a -0.055 decrease to 0.630 increase in mistriage. This means that with the negative UI (-0.055), the effect of transfer would be an improvement in model performance. The upper limit of the UI (0.630) would mean a marked decrease in model performance, especially compared to the point estimate of 0.31.

The transfer of clinical prediction models in trauma care has not previously been extensively studied. In 2016, Gerdin et al. also found model transfer to adversely affect model performance (18). They found the transferred model to perform poorly in terms of calibration, and that this could be improved by updating the model. In our current study we did not examine model performance in terms of calibration, neither did we examine the effects of updating the model. What the two studies have in common is an observed decline in model performance, when transferring the models between certain settings. Therefore, we believe model transfer can potentially lead to a decline in model performance compared to the setting in which it was originally developed.  Model transfer should be performed with caution, and the effects examined before implementing the transferred model in a clinical setting.

External validation of clinical prediction models, meaning the evaluation of a clinical prediction model in a setting in which it was not originally developed has however been studied more extensively. The concepts of external validation are applicable and comparable to our current study, with the exception being that out models were technically externally validated using a validation subsample before being transferred to an entirely different sample. Studies using simulated clinical prediction models based on different predictors often show a decline in model performance when being externally validated (33, 34), both underlining the necessity of testing models in other populations before clinical use. A study from 2018 examining the external validity of prediction models for coronary artery disease (CAD) (35), showing models significantly underestimating the probability of CAD. Our study did not show the same dramatic decrease in model performance after transfer but combined with the results of other studies of external validation it further warrants caution when transferring models between different settings.

## Strengths and limitations
This study examined model transfer in a limited setting (Sweden, as a single health care setting), which has advantages and disadvantages. The study design realistically reflects potential transfers: clinical prediction models for trauma are usually developed in large metropolitan centres and then transferred as clinical recommendations to minor non-metropolitan centres in which traumatic injuries are not as common.

In terms of disadvantages these results are restricted to a single health care system in Sweden, and the generalizability to other health care systems might be limited. Also, the small number of patients (27043) weakens the applicability of the results on a more general setting. We originally intended to study all single centres with a valid number of events, but when performing the analysis only one centre provided a sufficient number of events for study. Leading to the loss of the fourth data set: Individual centres, and the comparison between these. We recognize that the results this data set might have provided could reflect realistic situations as major trauma centres regularly exchange data to improve patient care.

In Sweden major trauma is relatively uncommon, especially in non-metropolitan and minor centres, and therefore it is reasonable to assume this leads to inconsistencies when performing SweTrau patient registration. Such as using NISS instead of ISS, registering SBP by category and not by absolute value, not following up on patient survival, or other local differences. This might lead to exaggerated performance differences between the different settings, as the transferred models are based on registry data.

We developed models using logistic regression based on registry data, instead of using pre-developed models such as the Revised Trauma Score, eliminating the risk of potentially misrepresenting already established models. The models used in this study used GCS, RR and SBP as predictors for 30-day mortality. Potentially, the use of additional predictors such as age, mechanism of injury, or newer predictors as chock index might increase model performance. Furthermore, we used 30-day mortality as an outcome. We recognize that late mortality, time in hospital, functional impairment, and morbidity are also important outcome factors in the context of trauma care.

Statistical analysis and comparisons were performed using R. We aimed to follow a predetermined analysis plan in a step-by-step fashion with to aim of avoiding confirmation bias. However, programming challenges and late-time code inclusions required a flexible approach to this analysis plan, with revisions of previous steps. We recognize that this is not optimal but aimed to maintain objectivity by not deviating in any major way from the analysis plan. Using GitHub, all code was made public for inspection and to ensure reproducibility (28).

## Clinical applications
Clinical prediction models are constantly being developed in different contexts, and often transferred to settings in which they were not originally developed. The results of this study show that mistriage can potentially increase following model transfer, in this case more prominently when transferring from a metropolitan setting to a non-metropolitan setting. This supports the view that model transfers should be done with caution, as performance might decline.

Recognizing these risks and updating the models accordingly will lead to an increased accuracy during the triage of trauma patients. This could potentially save lives by minimizing undertriage, and likely save resources by minimizing overtriage. Potentially, because of registries such as SweTrau, analysis of model transfer could be performed continuously to optimize models currently in use and advise which model transfers might be detrimental (or potentially beneficial).

## Future studies
To increase the generalisability, studies of this type could be performed in different settings e.g. assessing the transferability between different countries, such as Sweden and the United Kingdom. Also, increasing the number of patients would also increase the power of the study results, one potential way of doing this would be to replicate this study using the very latest data in the SweTrau registry, as many hospitals have joined recently which in turn has led to an increase in the number of patients registered.

Model transfer could potentially be studied using different methods. Quite often, as mentioned, clinical models are transferred from a major metropolitan centre to a smaller rural centre. Staff opinions considering how applicable this new model is could be obtained using an interview or questionnaire approach Clinical trials using different models could also be performed, however, appropriate ethical considerations are paramount before performing this type of study.

In terms of equity, patient sex and age were obtained to characterize the patient cohort: 66 % male, and a median age of 41 (mean 44). No actions were undertaken in response to this division. Further studies might aim to examine if model performance differs using sex or age as a predictor.

## Conclusion (Needs update with code snippets) 
A clinical prediction model developed using logistic regression and GCS, RR and SBP as predictors can be used to predict 30-day mortality in different settings. Model transfer in the metropolitan and non-metropolitan data set lead to increased mistriage rate (95% UI -0.055 - 0.630). This was mainly due to a marked increase in overtriage. In all samples, 95% UI for model transfer comparison included 0. Poor model transfer can potentially lead to increased patient mortality and a misappropriation of resources. We recognize the limitations of this study, but believe further studies are warranted due to the potentially severe patient consequences.

# References

### 2019-05-17 ManuscriptMarkdown.Rmd
Rename ResultsMarkdown.Rmd to ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

--- 2019-05-17 ResultsMarkdown.Rmd
@@
 author: "Martin Henriksson"
-date: "190429"
+date: "190517"
 output: word_document
@@
 Model validation performance varied across all samples with a median mistriage rate range of `r min(results$original.stats$Validation.mistriage.median)` - `r max(results$original.stats$Validation.mistriage.median)`.
-Uncertainty interval(UI) for validation mistriage, undertriage and overtriage ranged from 0 to 0.830 across all samples.
+Uncertainty interval(UI) for validation mistriage, undertriage and overtriage ranged from ...
+to ...
+across all samples.
 
@@
 The most notable effects on model performance following model transfer compared to the original sample model was seen when the metropolitan model was transferred to the non-metropolitan sample.
-The performance of the metropolitan model transferred to the non-metropolitan sample was a mistriage rate of 0,51, and the validation mistriage rate of the non-metropolitan model was 0,20.
-This indicates that the mistriage increased by 0,31 when using the metropolitan model in a non-metropolitan setting.
+The performance of the metropolitan model transferred to the non-metropolitan sample was a mistriage rate of ..., and the validation mistriage rate of the non-metropolitan model was ....
+This indicates that the mistriage increased by ...
+when using the metropolitan model in a non-metropolitan setting.
 Mainly contributing to this increase in mistriage was the marked increase in overtriage when performing this model transfer.
-The UI of this difference however ranged from -0.055 to 0.630, indicating that our findings are compatible with a -0.055 decrease to 0.630 increase in mistriage.
-This means that with the negative UI (-0.055), the effect of transfer would be an improvement in model performance.
-The upper limit of the UI (0.630) would mean a marked decrease in model performance, especially compared to the point estimate of 0.31.
+The UI of this difference however ranged from ...
+to ..., indicating that our findings are compatible with a ...
+decrease to ...
+increase in mistriage.
+This means that with the negative UI (...), the effect of transfer would be an improvement in model performance.
+The upper limit of the UI (...) would mean a marked decrease in model performance, especially compared to the point estimate of ....
 
@@
 A clinical prediction model developed using logistic regression and GCS, RR and SBP as predictors can be used to predict 30-day mortality in different settings.
-Model transfer in the metropolitan and non-metropolitan data set lead to increased mistriage rate (95% UI -0.055 - 0.630).
-This was mainly due to a marked increase in overtriage.
+Model transfer in the metropolitan and non-metropolitan data set lead to ...
+mistriage rate (95% UI ...).
+This was mainly due to ....
 In all samples, 95% UI for model transfer comparison included 0.

### 2019-05-20 functions/CompileResults.R
Update CompileResults.R

@@ -80,10 +80,12 @@ CompileResults <- function() {
     })
     names(point.estimate.with.uncertainty) <- sample.names
     ## Create result tables
-    result.tables <- lapply(point.estimate.with.uncertainty, function(result) {
-        kable(result, format = "markdown")
-    })
+    result.tables.data.frame <- point.estimate.with.uncertainty
+    # result.tables <- lapply(point.estimate.with.uncertainty, function(result) {
+    #     kable(result, format = "markdown")
+    # })
     ## Save result tables
-    saveRDS(result.tables, "result.tables.Rds")
+    saveRDS(result.tables.data.frame , "result.tables.data.frame.Rds")
+    # saveRDS(result.tables, "result.tables.Rds")
     message("Result tables saved")
 }

### 2019-05-20 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 ## LOADING ENVIROMENT AND TABLES
-results <- readRDS("compiled output/original.results.Rds")
+## Original results and UI tables
+results <- readRDS("output/original.results.Rds")
 result.tables <- readRDS("result.tables.Rds")
+result.tables.data.frame <- readRDS("result.tables.data.frame.Rds")
+## Sample characteristics table
 # table.one <-
@@
 worst.transfer <- results$original.stats[results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median == max(results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median),]
-## Figure and table setup
-# Figure 1 = Figure 1: High- and low volume centre data set. Initial split based on number of patients. Temporal split made using date of traumatic event.
+
+## Table setup
 # Table 1 = Sample characteristics table
+#
 # Table 2 = result.tables, cropped to only show validation
+table.2 <- lapply(result.tables.data.frame, function(x) x[1:3,])
+table.2 <- lapply(table.2, function(result) {
+           kable(result, format = "markdown")
+           })
 # Table 3 = Cropped to only show transfer
+table.3 <- lapply(result.tables.data.frame, function(x) x[4:6,])
+table.3 <- lapply(table.3, function(result) {
+           kable(result, format = "markdown")
+           })
 # Table 4 = Cropped to only show transfer differance
+table.4 <- lapply(result.tables.data.frame, function(x) x[7:9,])
+table.4 <- lapply(table.4, function(result) {
+           kable(result, format = "markdown")
+           })
 ```
@@
 
-```{r, echo = FALSE, results = "asis"} 
-## Figure 1 = Figure 1: High- and low volume centre data set. Initial split based on number of patients. Temporal split made using date of traumatic event. 
-```
+![**Figure 1** High- and low volume centre data set. Initial split based on number of patients. Temporal split made using date of traumatic event.](fig1.jpg)
 
@@
 ## Table 1 = Sample characteristics table
-##result.tables
+#
 ```
@@
 ## Table 2 = Cropped to only show validation
-result.tables
+table.2
 ```
@@
 ## Table 3 = Cropped to only show transfer
-result.tables
+table.3
 ```
@@
 ## Table 4 = Cropped to only show transfer differance
-result.tables
+table.4
 ```

### 2019-05-28 ManuscriptMarkdown.Rmd
Add files via upload

@@
 author: "Martin Henriksson"
-date: "190517"
+date: "190429"
 output: word_document
@@
 
-Trauma care involves many levels of the health care system.
 In a typical, high resource setting, the initial management of patients is performed on the scene of the trauma by Emergency Medical Services (EMS).
-In this prehospital phase; patient data, vital signs and scene information is collected and transferred to the receiving hospital.
-This information is then evaluated with the help of a system to determine the level of trauma and prepare adequate resources (4).
-The level of trauma dictates if a full or limited trauma team is activated (5).
+Patient data, vital signs and scene information is transferred to the receiving hospital.
+This information is evaluated using a system to determine the level of trauma and prepare adequate resources (4), dictating if a full or limited trauma team is activated (5).
 
 Systems to determine the level of trauma can be based on clinical prediction models.
-Such models are continuously being developed and studied, examining different parameters and different contexts concerning trauma care.
-These models differ in quality and characteristics, but they generally perform well when predicting survival (6).
-Many of these models are developed in a single standardized setting such as a large metropolitan area or a major trauma centre but are then supposed to be implemented in different settings, such as non-trauma centres.
+Such models are continuously being developed and studied, examining different parameters and different contexts.
+Models differ in quality and characteristics, but generally perform well predicting survival (6).
+Many of these models are developed in a single standardized setting such as a large metropolitan area or a major trauma centre but are then implemented in different settings, such as non-trauma centres.
 
-What is not fully understood is how this transfer affects the performance of these models.
-Previous research that studied the transfer of clinical prediction models between settings has shown that model performance in terms of calibration can be adversely affected (7).
+What is not fully understood is how this transfer affects model performance.
+Previous research has shown that model performance in terms of calibration can be adversely affected (7).
 However, the aforementioned research studied model transfers between substantially different settings and did not assess more clinically relevant performance measures, such as misclassification.
@@
 In trauma, misclassification is often referred to as mistriage.
-Here, triage refers to the classification of severity as minor or major trauma.
-Mistriage can be subdivided into overtriage; the incorrect classification of a patient with minor trauma as major trauma, or undertriage; the incorrect classification of a patient with major trauma as minor trauma.
+Triage refers to the classification of severity as minor or major trauma.
+Mistriage can be subdivided into overtriage; incorrect classification of a patient with minor trauma as major trauma, or undertriage; incorrect classification of a patient with major trauma as minor trauma.
 Both are detrimental in terms of patient care and distribution of resources.
@@
 Sweden has a nationally encompassing register for registering trauma called SweTrau.
-We used SweTrau data to create clinical prediction models, which we then transferred between different contexts to study the effects on mistriage.
-Please refer to appendix I for a brief explanatory list of the statistical terminology.
+SweTrau data was used to create clinical prediction models, which were transferred between different contexts to study the effects on mistriage.
 
@@
 
-The trauma care organization in Sweden is usually clearly defined in major metropolitan areas (albeit the specific routines differ), and trauma patients are transported directly to a predesignated hospital with specific trauma competency and capacity.
+The trauma care organization in Sweden is usually clearly defined in major metropolitan areas: trauma patients are transported directly to a predesignated hospital with specific trauma competency.
 In large, more rural parts of the country, trauma patients are instead transported to the nearest hospital for stabilization.
 Once stabilized, they may then be transported to hospitals with more advanced trauma care capacity (19).
-Most hospitals in Sweden use some form of system to categorize the level of trauma and necessary response.
-These systems are usually based on the patient's vital signs and injury mechanism, as reported by the EMS before patient arrival, or as registered in the emergency department if the patient arrives by other means (16).
+Most hospitals in Sweden use some form of system to categorize the level of trauma.
+These systems are usually based on patient vital signs and injury mechanism, as reported by the EMS, or as registered in the emergency department (16).
 
-We used data from SweTrau between 2011 and 2016, as this study’s inception was in 2016 and data obtained the same year.
+Data was obtained from SweTrau between 2011 and 2016.
 Today 95,5% (52 of 55) of Swedish hospitals record trauma cases in SweTrau.
 Currently the registry includes 55 000 cases, and the board encourages its use in both academic research and more local quality improvement initiatives (20).
-Included hospitals are both university and non-university hospitals, all of which receive trauma patients.
 
-Patient data is recorded upon arrival to the hospital.
-All hospital personnel can perform patient registration, but it is recommended they received Abbreviated Injury Scale (AIS) training.
 The SweTrau inclusion criteria are: Traumatic event with subsequent activation of hospital trauma protocol, admitted patients with NISS (New Injury Severity Score) > 15 and patients transferred to the hospital within 7 days of traumatic event with NISS > 15.
@@
 ## Participants 
-The eligibility criteria are patients registered in SweTrau, and age 15 or above.
+The eligibility criteria were patients registered in SweTrau, age 15 or above.
 This age was decided as the study aims to study adult trauma and not paediatric trauma which differs in physiology, triage and initial care (22).
-We recognize this age cut-off is not completely without controversy.
-However, in Sweden patients age ≥15 go to the adult emergency rooms and the Swedish guidelines for trauma activation define children as age <15 (16).
+The Swedish guidelines for trauma activation define children as age <15 (16).
 Also, several guidelines and protocols listed in ATLS use age ≥15 as cut-off (9).
@@
 ### Model predictors
-Our clinical prediction models included the predictors SBP, RR and GCS on arrival to hospital.
+The clinical prediction models included the predictors SBP, RR and GCS on arrival to hospital.
 The rationale for including these three predictors is that they are part of many established clinical prediction models for early trauma care, such as the Revised Trauma Score (23).
@@
 ### Model outcome
-The outcome that we used to develop the clinical prediction models was all cause mortality within 30 days of the trauma.
+The outcome used to develop the clinical prediction models was all cause mortality within 30 days of the trauma.
 
 ### Participant characteristics
-To describe the patient cohort we reported age, sex, ASA physical status classification system, Injury Severity Score (ISS) and New ISS (NISS).
+To describe the patient cohort we reported age, sex, Injury Severity Score (ISS) and New ISS (NISS).
 
 ### Study outcome
-We used ISS > 15 as the gold standard to define trauma severity as major trauma, and hence patients with ISS ≤ 15 were considered minor trauma (24).
-We defined overtriage as the event when a clinical prediction model classified a patient with ISS ≤ 15 as major trauma, and undertriage as the event when a clinical prediction model classified a patient with ISS > 15 as minor trauma.
-Clinically, in the event of overtriage, a full trauma team is activated and tasked with the care of a minimally injured patient.
-The opposite is true with undertriage, where a patient with serious injuries is managed with (potentially) inadequate resources  (10).
-We defined the overtriage rate as the number of overtriaged patients divided by all patients.
-We defined the undertriage rate as the number of undertriaged patients divided by all patients.
-The mistriage rate was defined as the sum of the over- and undertriage rates.
+ISS > 15 was used as the gold standard to define trauma severity as major trauma, and hence patients with ISS ≤ 15 were considered minor trauma (24).
+Overtriage was defined as the event when a clinical prediction model classified a patient with ISS ≤ 15 as major trauma, and undertriage as the event when a clinical prediction model classified a patient with ISS > 15 as minor trauma.
+Overtriage rate was defined as the number of overtriaged patients divided by all patients.
+Undertriage rate was defined as the number of undertriaged patients divided by all patients.
+Mistriage rate was defined as the sum of the over- and undertriage rates.
 
 ## Data sources and measurements
-Model predictors, outcome, participant characteristics and study outcome as outlined above will all be obtained from the SweTrau registry.
+Model predictors, outcome, participant characteristics and study outcome as outlined above were all obtained from SweTrau.
 The method of measurement for the model predictors is not specified (for example, if SBP is measured using an automated cuff or manually) in the registry entries.
@@
 Whether the patient is dead or alive 30 days after trauma is manually entered into the registry locally, by each respective hospital.
-Foreign citizens discharged within 30 days are registered as survivors.
-Participant characteristics; sex and age are obtained from the patient file before being registered in SweTrau.
-ASA, NISS and ISS are calculated by hospital personal based on the patient’s injuries, and registered.
+Participant characteristics are obtained from the patient file before being registered in SweTrau.
+NISS and ISS are calculated by hospital personal based on the patient’s injuries.
 
 ## Bias
-Data analysis was conducted in a step-by-step fashion, and according to a prearranged analysis plan.
-The analysis plan and statistical analysis code was finalised using simulated data instead of the real data.
-These efforts were taken to avoid confirmation bias when conducting the data analysis.
-The analysis plan was reviewed by an experienced statistician and programmer prior to implementation to ensure objectivity.
-Neither outcome nor variable were blinded when conducting data analysis, which made a structured objective approach even more important.
+Data analysis was conducted according to a prearranged analysis plan.
+The analysis plan and statistical analysis code was finalised using simulated data.
+These efforts were taken to avoid confirmation bias.
+The analysis plan was reviewed by an experienced statistician and programmer prior to implementation, to ensure objectivity.
+Neither outcome nor variable were blinded when conducting data analysis, which made a structured objective approach important.
 
 ## Study size
-All patients matching eligibility criteria are listed above.
+All patients matching eligibility criteria listed above.
 Four data sets were used to study the transfer of clinical prediction models, each data set representing a different setting.
@@
 ## Quantitative variables
-SBP and RR was modelled using restricted cubic splines with four knots whenever possible, placed at equally spaced percentiles and GCS as a continuous linear term.
-When describing the participant characteristics all quantitative variables were to be presented as continuous.
-ISS was also presented as dichotomous using ISS > 15 as the cutoff.
+SBP and RR was modelled using restricted cubic splines with four knots whenever possible placed at equally spaced percentiles, and GCS as a continuous linear term.
+When describing the participant characteristics quantitative variables were presented as continuous.
+ISS was presented as dichotomous using ISS > 15 as the cutoff.
 
@@
 
-### Indicidual centres
-This data set was also be split into multiple samples.
+### Individual centres
+This data set was also split into multiple samples.
 Each centre with large enough sample size to develop and validate a model constituted its own sample.
@@
 The minimum sample size of development and validation samples was 140 and 200 respectively.
-The total number of events and non-events per data set was therefore at least 680.
 We performed analyses only on data sets for which all samples included at least the minimum number of patients.
@@
 ### Sequence of analysis
-The programming language R was used for all analyses (27), and all progress was uploaded to a GitHub repository for inspection (28).
-We performed the analyses in the sequence of model development, model validation and finally model comparison.
+The programming language R was used for all analyses (27), all progress was uploaded to a GitHub repository (28).
+Analyses were performed in the sequence of model development, model validation and finally model comparison.
 These steps were repeated in each data set.
@@
 The shrunk model was then used to estimate the probability of all cause 30-day mortality in the development sample.
-We then performed a gridsearch across estimated probabilities in the development sample to identify the cut-off that optimised overtriage keeping undertriage at less than 5% (11).
+A gridsearch was performed across estimated probabilities in the development sample to identify the cut-off that optimised overtriage keeping undertriage at less than 5% (11).
 This cut-off was then used to classify patients as major or minor trauma.
@@
 Finally, in the model comparison step, the difference in model performance between the high and low volume centre validation samples was calculated.
-We used an empirical bootstrap to estimate 95% confidence intervals (CI) around performance and differences in performance estimates.
+Empirical bootstrap was used to estimate 95% confidence intervals (CI) around performance and differences in performance estimates.
 Both bootstrap procedures used 1000 bootstrap samples drawn with replacement of the same size as the original samples.
@@
 ## Ethical considerations
-This study was conducted in accordance with the four major principles of medical ethics and The Declaration of Helsinki (31, 32).
-
-### Respect for autonomy
-The respect for autonomy was upheld as Swedish Privacy Law requires the informed consent of all participants.
-Patients included in SweTrau receive letters detailing that data obtained may be used for scientific purposes.
-They are also informed that participation is voluntary and that they may withdraw at any time.
-
-### The Principle of Beneficence
-Throughout the entire study we attempted to act in the best interest of the patients.
-In the long run we hope this research will improve the care of trauma patients, and therefore prove beneficial to those affected.
-We also believe that the patients who consented to research want to contribute to this improvement of care, and as thus by performing this study we hope to honour their wishes.
-
-### The Principle of Nonmaleficence
-All attempts were made not to harm any patients included in this study, either intentionally or unintentionally.
-Data leakage and patient identification constitutes the primary risk.
-All data obtain was depersonalized and handled with care to minimize risk of unauthorised access and subsequent patient identification.
-By implementing these actions, we determined that the risk to the patient population is minimal.
-
-### The Principle of Justice
-As this study contained no intervention, the principle of justice does not truly apply.
-However, all cases included from the registry were treated as equal.
-It is also our wish that any scientific gains this study may contribute will be implemented in a fair and equal fashion.
-
-### Ethical permit
-This study is approved by the regional ethics review board in Stockholm, Sweden. Ethical review numbers are 2015/426-31 and 2016/461-32.
+This study is approved by the regional ethics review board in Stockholm, Sweden.
+Ethical review numbers are 2015/426-31 and 2016/461-32.
+All patients included recived letters detainling SweTrau inclusion, and the possibility of their data being used for scientific purposes.
+It was assumed that the patients who consented to research want to contribute to this improvement of care, and as thus by performing this study the authors hope to honour their wishes.
 
@@
 ## Table 1 = Sample characteristics table
-#
+##result.tables
 ```
@@
 During model development, the number of imputations used for each data set was `r high.low.imp`, `r metro.non.imp` and `r multi.single.imp` for the high and low volume data set, the metropolitan and non-metropolitan data set and the multi and single centre data set respectively.
-This number was equal to the maximum percentage of missing observations for each sample in the respective data sets.
-Restricted cubic splines with four knots placed at equally spaced percentiles were implemented when modelling systolic blood pressure and respiratory rate.
-1000 bootstrap samples drawn with replacement was used to estimate a linear shrinkage factor that was applied to the developed models to avoid model overfitting.
 
 Table 2 shows model validation performance.
-Model validation performance refers to the model performance when applied in the validation subsample from the sample in which it was developed.
-The sample with the best validation model performance was the `r best.validation$Sample.name` sample with a median mistriage rate of `r best.validation$Validation.mistriage.median`.
-Performance in terms of under and overtriage for the same model was `r best.validation$Validation.undertriage.median`, and `r best.validation$Validation.overtriage.median`.
-Model validation performance varied across all samples with a median mistriage rate range of `r min(results$original.stats$Validation.mistriage.median)` - `r max(results$original.stats$Validation.mistriage.median)`.
-Uncertainty interval(UI) for validation mistriage, undertriage and overtriage ranged from ...
-to ...
-across all samples.
+The sample with the lowest mistriage was the `r best.validation$Sample.name` sample with a median mistriage rate of `r best.validation$Validation.mistriage.median`.
+Performance in terms of under and overtriage for the same model was `r best.validation$Validation.undertriage.median`, and `r best.validation$Validation.overtriage.median` respectivly.
+Across all samples model validation performance in terms of mistriage ranged from `r min(results$original.stats$Validation.mistriage.median)` to `r max(results$original.stats$Validation.mistriage.median)`.
 
+**Table 2** Model validation performance  
 ```{r, echo = FALSE, results = "asis"} 
@@
 Model performance after transfer was determined for each model being transferred to the other sample in the data set in all data sets (Table 3).
-When transferred, the model with the best performance i.e.
-lowest mistriage rate was the `r best.transfer$Sample.name` model with a median mistriage of `r best.transfer$Transfer.mistriage.median`.
-Concerning the UI for transfer mistriage, undertriage and overtriage, this too ranged from 0 to 0.830 across all samples.
+When transferred, the model with the lowest mistriage rate was the `r best.transfer$Sample.name` model with a median mistriage of `r best.transfer$Transfer.mistriage.median`.
 
+**Table** 3 Model transfer performance  
 ```{r, echo = FALSE, results = "asis"} 
@@
 
-The main purpose of this study was to examine the transferability of clinical prediction models for trauma.
-Table 4 shows the transferred model performance when compared to the validation performance in the same sample, transferred model performance varied between samples and data sets.
+Table 4 shows the transferred model performance when compared to the validation performance in the sample reciving the transferred model, e.g.
+high volume transfer performance minus low volume validation performance.
 
-The lowest model performance when compared to the sample validation performance was found in the `r worst.transfer$Sample.name` sample, with a median mistriage difference of `r worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median`, meaning this model performed worse when transferred to the other sample in the same data set, when compared with the model developed and validated in this sample.
+The highest mistriage when compared to the reciveing sample validation model performance was found in the `r worst.transfer$Sample.name` sample, with a median mistriage difference of `r worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median`, meaning this model performed worse when transferred to the other sample in the same data set..
 In clinical terms, this model transfer means that among 100 trauma patients, `r abs(worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median) * 100` more patients would be wrongly classified as major or minor trauma.
-However, when uncertainty intervals (UI) were obtained no observed differences between transferred model performances and validation model performances in any sample were found to be significant.
 
+**Table 4** Comparison mistriage (Sample validation performance subtracted from transferred performance)  
 ```{r, echo = FALSE, results = "asis"} 
@@
 This study aimed to assess how transfers of clinical prediction models for early trauma care between different contexts within a single health system affect mistriage rates.
-The most notable effects on model performance following model transfer compared to the original sample model was seen when the metropolitan model was transferred to the non-metropolitan sample.
-The performance of the metropolitan model transferred to the non-metropolitan sample was a mistriage rate of ..., and the validation mistriage rate of the non-metropolitan model was ....
-This indicates that the mistriage increased by ...
+The most notable effects on model performance following model transfer were seen after transferring the metropolitan model.
+Mistriage increased by ...
 when using the metropolitan model in a non-metropolitan setting.
 Mainly contributing to this increase in mistriage was the marked increase in overtriage when performing this model transfer.
-The UI of this difference however ranged from ...
-to ..., indicating that our findings are compatible with a ...
-decrease to ...
-increase in mistriage.
-This means that with the negative UI (...), the effect of transfer would be an improvement in model performance.
-The upper limit of the UI (...) would mean a marked decrease in model performance, especially compared to the point estimate of ....
+For this particular transfer 95% UI ranged from ...
+to ..., indicating that our findings are compatible with a -0.055 decrease to 0.630 increase in mistriage.
+The negative 95% UI of ...
+would signifiy an improvement in model performance.
+The higher 95% UI of ...
+would mean a marked increase in model mistriage, compared to the point estimate of ....
 
@@
 In 2016, Gerdin et al.
-also found model transfer to adversely affect model performance (18).
-They found the transferred model to perform poorly in terms of calibration, and that this could be improved by updating the model.
-In our current study we did not examine model performance in terms of calibration, neither did we examine the effects of updating the model.
-What the two studies have in common is an observed decline in model performance, when transferring the models between certain settings.
-Therefore, we believe model transfer can potentially lead to a decline in model performance compared to the setting in which it was originally developed.
-Model transfer should be performed with caution, and the effects examined before implementing the transferred model in a clinical setting.
+found model transfer to adversely affect model performance in terms of calibration, but that this could be improved by updating the model (18).
+This study did not examine model performance in terms of calibration, neither was the effects of updating the model explored.
 
-External validation of clinical prediction models, meaning the evaluation of a clinical prediction model in a setting in which it was not originally developed has however been studied more extensively.
-The concepts of external validation are applicable and comparable to our current study, with the exception being that out models were technically externally validated using a validation subsample before being transferred to an entirely different sample.
-Studies using simulated clinical prediction models based on different predictors often show a decline in model performance when being externally validated (33, 34), both underlining the necessity of testing models in other populations before clinical use.
-A study from 2018 examining the external validity of prediction models for coronary artery disease (CAD) (35), showing models significantly underestimating the probability of CAD.
-Our study did not show the same dramatic decrease in model performance after transfer but combined with the results of other studies of external validation it further warrants caution when transferring models between different settings.
+External validation as the evaluation of a clinical prediction model in a setting in which it was not originally developed has been studied more extensively.
+Studies using simulated clinical prediction models based on different predictors showed a decline in model performance when being externally validated (33, 34).
+A study from 2018 examining the external validity of prediction models for coronary artery disease (CAD) (35), showed models significantly underestimating the probability of CAD.
+This study did not find the same dramatic decrease in model performance after transfer.
+However, the combined results of this study, the studies on model transfer and the studies on external validation warrants further caution when transferring models between different settings.
 
 ## Strengths and limitations
-This study examined model transfer in a limited setting (Sweden, as a single health care setting), which has advantages and disadvantages.
-The study design realistically reflects potential transfers: clinical prediction models for trauma are usually developed in large metropolitan centres and then transferred as clinical recommendations to minor non-metropolitan centres in which traumatic injuries are not as common.
+This study design realistically reflects potential transfers: e.g.
+clinical prediction models for trauma are usually developed in large metropolitan centres and then transferred as clinical recommendations to minor non-metropolitan centres in which traumatic injuries are not as common.
 
-In terms of disadvantages these results are restricted to a single health care system in Sweden, and the generalizability to other health care systems might be limited.
-Also, the small number of patients (27043) weakens the applicability of the results on a more general setting.
-We originally intended to study all single centres with a valid number of events, but when performing the analysis only one centre provided a sufficient number of events for study.
+Originally all single centres with a valid number of events were to constitute one data set, but when performing the analysis only one centre provided a sufficient number of events.
 Leading to the loss of the fourth data set: Individual centres, and the comparison between these.
-We recognize that the results this data set might have provided could reflect realistic situations as major trauma centres regularly exchange data to improve patient care.
+Possibly, the results this data set might have provided results that could reflect realistic situations as major trauma centres regularly exchange data to improve patient care.
 
-In Sweden major trauma is relatively uncommon, especially in non-metropolitan and minor centres, and therefore it is reasonable to assume this leads to inconsistencies when performing SweTrau patient registration.
-Such as using NISS instead of ISS, registering SBP by category and not by absolute value, not following up on patient survival, or other local differences.
-This might lead to exaggerated performance differences between the different settings, as the transferred models are based on registry data.
+In Sweden major trauma is relatively uncommon, this might lead to inconsistencies across centres when performing SweTrau registration.
+E.g.
+using NISS instead of ISS, registering SBP by category and not by absolute value or not following up on patient survival.
+This might lead to exaggerated performance differences between the settings.
 
-We developed models using logistic regression based on registry data, instead of using pre-developed models such as the Revised Trauma Score, eliminating the risk of potentially misrepresenting already established models.
-The models used in this study used GCS, RR and SBP as predictors for 30-day mortality.
-Potentially, the use of additional predictors such as age, mechanism of injury, or newer predictors as chock index might increase model performance.
-Furthermore, we used 30-day mortality as an outcome.
-We recognize that late mortality, time in hospital, functional impairment, and morbidity are also important outcome factors in the context of trauma care.
+The models in this study used GCS, RR and SBP as predictors for 30-day mortality.
+Potentially, the use of additional predictors such as age, mechanism of injury, or newer predictors as shock index might increase model performance.
+Furthermore, we used 30-day mortality as an outcome, recognizing that late mortality, time in hospital, functional impairment, and morbidity are also important outcome factors in the context of trauma care.
 
-Statistical analysis and comparisons were performed using R.
-We aimed to follow a predetermined analysis plan in a step-by-step fashion with to aim of avoiding confirmation bias.
-However, programming challenges and late-time code inclusions required a flexible approach to this analysis plan, with revisions of previous steps.
-We recognize that this is not optimal but aimed to maintain objectivity by not deviating in any major way from the analysis plan.
+The predetermined analysis plan was to be followed in a step-by-step fashion, to avoiding confirmation bias.
+Programming challenges and late-time code inclusions required a flexible approach to this analysis plan, sometimes with revisions of previous steps.
+This is not optimal, and efforts were made to not deviate in any major way.
 Using GitHub, all code was made public for inspection and to ensure reproducibility (28).
@@
 ## Clinical applications
-Clinical prediction models are constantly being developed in different contexts, and often transferred to settings in which they were not originally developed.
 The results of this study show that mistriage can potentially increase following model transfer, in this case more prominently when transferring from a metropolitan setting to a non-metropolitan setting.
@@
 
-Recognizing these risks and updating the models accordingly will lead to an increased accuracy during the triage of trauma patients.
-This could potentially save lives by minimizing undertriage, and likely save resources by minimizing overtriage.
-Potentially, because of registries such as SweTrau, analysis of model transfer could be performed continuously to optimize models currently in use and advise which model transfers might be detrimental (or potentially beneficial).
+Recognizing these risks and updating models accordingly will lead to increased accuracy during trauma patient triage.
+Potentially saving lives by minimizing undertriage, and likely save resources by minimizing overtriage.
+With registries such as SweTrau, analysis of model transfer could be performed continuously to optimize models currently in use and advise which model transfers might be detrimental (or potentially beneficial).
 
 ## Future studies
-To increase the generalisability, studies of this type could be performed in different settings e.g.
-assessing the transferability between different countries, such as Sweden and the United Kingdom.
-Also, increasing the number of patients would also increase the power of the study results, one potential way of doing this would be to replicate this study using the very latest data in the SweTrau registry, as many hospitals have joined recently which in turn has led to an increase in the number of patients registered.
+Studies on model transfer can be performed in different settings, e.g.
+assessing the transferability between different but similar countries, such as Sweden and Norway.
+Also, increasing the number of patients would increase the power of the results, one way of doing this would be to replicate this study with the latest data in the SweTrau, as many hospitals have joined recently leading to an increased number of patients registered.
 
@@
 
-In terms of equity, patient sex and age were obtained to characterize the patient cohort: 66 % male, and a median age of 41 (mean 44).
-No actions were undertaken in response to this division.
-Further studies might aim to examine if model performance differs using sex or age as a predictor.
-
 ## Conclusion (Needs update with code snippets) 
-A clinical prediction model developed using logistic regression and GCS, RR and SBP as predictors can be used to predict 30-day mortality in different settings.
-Model transfer in the metropolitan and non-metropolitan data set lead to ...
-mistriage rate (95% UI ...).
-This was mainly due to ....
-In all samples, 95% UI for model transfer comparison included 0.
-Poor model transfer can potentially lead to increased patient mortality and a misappropriation of resources.
-We recognize the limitations of this study, but believe further studies are warranted due to the potentially severe patient consequences.
+Model transfer from the metropolitan sample to the non-metropolitan sample led to an increased mistriage rate of ...
+(95% UI ...).
+Mainly due to a marked increase in overtriage.
+Across all samples, 95% UI for model transfer comparison included ....
+Model transfer can potentially lead to increased patient mortality and a misappropriation of resources, due to increased undertriage and overtriage respectively.
+Recognizing the limitations of this study, the authors believe further studies are warranted due to the potential economic costs and patient consequences of poor model transfers.
 

### 2019-05-29 ManuscriptMarkdown.Rmd
Add files via upload

@@
 knitr::opts_chunk$set(echo = TRUE)
-## LOADING ENVIROMENT AND TABLES
+## LOAD FUNCTIONS, ENVIROMENT AND TABLES
+source("functions/WorkTableCreator.R")
 ## Original results and UI tables
@@
 result.tables.data.frame <- readRDS("result.tables.data.frame.Rds")
+## Create table for ease of data retrival
+results.all.data <- WorkTableCreator(result.tables.data.frame)
 ## Sample characteristics table
@@
 
-## PREPARED CODE
+## RESULTS
+## Characteristics paragraph
   ## Total patients and NA
@@
   na.vector <- unlist(results$NA.info.sample)
+  names(na.vector) <- c("high volume", "high volume", "low volume", "low volume", "metropolitan", "metropolitan", "non metropolitan", "non metropolitan", "multi centre", "multi centre", "single centre", "single centre")
   highest.percent.missing.number <- max(na.vector[na.vector<100])
@@
 ## Development and validation
+  ## Imputation data
   high.low.imp <- results$original.stats$Total.imputations[results$original.stats$Sample.name == "high.volume"]
@@
   max.imp <- max(results$original.stats$Total.imputations)
-best.validation <- results$original.stats[results$original.stats$Validation.mistriage.median == min(results$original.stats$Validation.mistriage.median),]
+  ## Validation data
+  sorted.validation.mistriage <- sort(results.all.data[row.names(results.all.data) == "Validation.mistriage",], decreasing = FALSE)
+  lowest.val.mistriage <- sorted.validation.mistriage[1]
+  val.undertriage <- results.all.data["Validation.undertriage", colnames(lowest.val.mistriage)]
+  val.overtriage <- results.all.data["Validation.overtriage", colnames(lowest.val.mistriage)]
+  highest.val.mistriage <- sorted.validation.mistriage[6]
 ## Comparison
-best.transfer <- results$original.stats[results$original.stats$Transfer.mistriage.median == min(results$original.stats$Transfer.mistriage.median),]
+  ## Transfer data
+  sorted.transfer.mistriage <- sort(results.all.data[row.names(results.all.data) == "Transfer.mistriage",], decreasing = FALSE)
+  lowest.tran.mistriage <- sorted.transfer.mistriage[1]
+  highest.tran.mistriage <- sorted.transfer.mistriage[6]
+  ## Comparison data
+  sorted.comp.mistriage <- sort(results.all.data[row.names(results.all.data) == "Transferred.mistriage.minus.buddy.local.mistriage",], decreasing = FALSE)
+  highest.comp.mistriage <- sorted.comp.mistriage[6]
   worst.transfer <- results$original.stats[results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median == max(results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median),]
-
+## DISCUSSION
+  ## First paragraph
+  UI.interval <- sub("\\).*", "", sub(".*\\(", "", format(highest.comp.mistriage)))
+  UI.top.bot.list <- strsplit(UI.interval, split = " - ")
+  UI.top.bot <- unlist(UI.top.bot.list)
 ## Table setup
@@
 Table 2 shows model validation performance.
-The sample with the lowest mistriage was the `r best.validation$Sample.name` sample with a median mistriage rate of `r best.validation$Validation.mistriage.median`.
-Performance in terms of under and overtriage for the same model was `r best.validation$Validation.undertriage.median`, and `r best.validation$Validation.overtriage.median` respectivly.
-Across all samples model validation performance in terms of mistriage ranged from `r min(results$original.stats$Validation.mistriage.median)` to `r max(results$original.stats$Validation.mistriage.median)`.
+The model with the lowest validation mistriage rate was the `r colnames(lowest.val.mistriage)` model with a median mistriage rate with 95% UI of `r lowest.val.mistriage`.
+Performance in terms of undertriage and overtriage rate for the same model was `r val.undertriage`, and `r val.overtriage` respectively.
+Worst validation performance i.e.
+highest mistriage rate was found in the `r colnames(highest.val.mistriage)` sample, with a model median mistriage rate of `r highest.val.mistriage`.
 
@@
 ## Comparison
-Model performance after transfer was determined for each model being transferred to the other sample in the data set in all data sets (Table 3).
-When transferred, the model with the lowest mistriage rate was the `r best.transfer$Sample.name` model with a median mistriage of `r best.transfer$Transfer.mistriage.median`.
+Model performance after transfer was determined for each model being transferred to the other sample in the data set, in all data sets (Table 3).
+When transferred, the model with the lowest mistriage rate was the `r colnames(lowest.tran.mistriage)` model with a median mistriage rate of `r lowest.tran.mistriage`.
+The model with the highest mistriage rate after transfer was the `r colnames(highest.tran.mistriage)` with a median mistriage rate of `r highest.tran.mistriage`
 
@@
 
-The highest mistriage when compared to the reciveing sample validation model performance was found in the `r worst.transfer$Sample.name` sample, with a median mistriage difference of `r worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median`, meaning this model performed worse when transferred to the other sample in the same data set..
+The highest mistriage rate when compared to the reciveing sample validation model performance was found in the `r colnames(highest.comp.mistriage)` model, with a median mistriage rate difference of `r highest.comp.mistriage`, meaning this model performed worse when transferred to the other sample in the same data set.
 In clinical terms, this model transfer means that among 100 trauma patients, `r abs(worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median) * 100` more patients would be wrongly classified as major or minor trauma.
@@
 
-# Discussion (Needs update with code snippets) 
+# Discussion
 This study aimed to assess how transfers of clinical prediction models for early trauma care between different contexts within a single health system affect mistriage rates.
-The most notable effects on model performance following model transfer were seen after transferring the metropolitan model.
-Mistriage increased by ...
-when using the metropolitan model in a non-metropolitan setting.
+The most notable effects on model performance following model transfer were seen after transferring the `r colnames(highest.comp.mistriage)` model.
+The mistriage rate increased by `r substr(format(highest.comp.mistriage), 1, 5)` when using the `r colnames(highest.comp.mistriage)` model in the other sample in the same data set.
 Mainly contributing to this increase in mistriage was the marked increase in overtriage when performing this model transfer.
-For this particular transfer 95% UI ranged from ...
-to ..., indicating that our findings are compatible with a -0.055 decrease to 0.630 increase in mistriage.
-The negative 95% UI of ...
-would signifiy an improvement in model performance.
-The higher 95% UI of ...
-would mean a marked increase in model mistriage, compared to the point estimate of ....
+For this transfer the 95% UI ranged from `r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating that our findings are compatible with a `r UI.top.bot[1]` decrease to a `r UI.top.bot[2]` increase in mistriage.
+The negative 95% UI of `r UI.top.bot[1]` would signifiy an improvement in model performance.
+The higher 95% UI of `r UI.top.bot[2]` would mean a marked increase in model mistriage, compared to the point estimate of `r substr(format(highest.comp.mistriage), 1, 5)`.
 

### 2019-05-30 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd
Add files via upload

@@
 ## LOAD FUNCTIONS, ENVIROMENT AND TABLES
+  ## Source functions
   source("functions/WorkTableCreator.R")
-## Original results and UI tables
+  ## Original results
   results <- readRDS("output/original.results.Rds")
+  ## Compiled results with UI
   result.tables <- readRDS("result.tables.Rds")
+  ## Compiled results with UI as data frame
   result.tables.data.frame <- readRDS("result.tables.data.frame.Rds")
@@
   # Table 2 = result.tables, cropped to only show validation
-  table.2 <- lapply(result.tables.data.frame, function(x) x[1:3,])
-  table.2 <- lapply(table.2, function(result) {
-             kable(result, format = "markdown")
-             })
+  table.2 <- results.all.data[1:3,]
+  rownames(table.2) <- c("Mistriage", "Undertriage", "Overtriage")
   # Table 3 = Cropped to only show transfer
-  table.3 <- lapply(result.tables.data.frame, function(x) x[4:6,])
-  table.3 <- lapply(table.3, function(result) {
-             kable(result, format = "markdown")
-             })
+  table.3 <- results.all.data[4:6,]
+  rownames(table.3) <- c("Mistriage", "Undertriage", "Overtriage")
   # Table 4 = Cropped to only show transfer differance
-  table.4 <- lapply(result.tables.data.frame, function(x) x[7:9,])
-  table.4 <- lapply(table.4, function(result) {
-             kable(result, format = "markdown")
-             })
+  table.4 <- results.all.data[7:9,]
+  rownames(table.4) <- c("Mistriage", "Undertriage", "Overtriage")
 ```
@@
 
-![**Figure 1** High- and low volume centre data set. Initial split based on number of patients. Temporal split made using date of traumatic event.](fig1.jpg)
+![](fig1.jpg)  
+**Figure 1: High- and low volume centre data set.** Initial split based on number of patients. Temporal split made using date of traumatic event.
 
@@
 
+**Table 1. Sample characteristics of multiple imputed data.** Data is presented as medians with interquartile range (IQR) or counts with % as applicable.
 ```{r, echo = FALSE, results = "asis"} 
@@
 
-**Table 2** Model validation performance  
+**Table 2. Model validation performance, i.e. performance when applied to the sample in which the model was created.** Data is presented as point estimates with 95% Uncertainty Intervals (95% UI).
 ```{r, echo = FALSE, results = "asis"} 
-## Table 2 = Cropped to only show validation
-table.2
+## Table 2
+kable(table.2[,1:2], format = "markdown")
+kable(table.2[,3:4], format = "markdown")
+kable(table.2[,5:6], format = "markdown")
 ```
@@
 
-**Table** 3 Model transfer performance  
+**Table 3. Model transfer performance, i.e. performance when transferred to the other sample in the same data set. ** Data is presented as point estimates with 95% Uncertainty Intervals (95% UI).  
 ```{r, echo = FALSE, results = "asis"} 
-## Table 3 = Cropped to only show transfer
-table.3
+kable(table.3[,1:2], format = "markdown")
+kable(table.3[,3:4], format = "markdown")
+kable(table.3[,5:6], format = "markdown")
 ```
@@
 
-**Table 4** Comparison mistriage (Sample validation performance subtracted from transferred performance)  
+**Table 4. Comparison mistriage, i.e. sample validation performance subtracted from transferred performance.** Data is presented as point estimates with 95% Uncertainty Intervals (95% UI).  
 ```{r, echo = FALSE, results = "asis"} 
-## Table 4 = Cropped to only show transfer differance
-table.4
+kable(table.4[,1:2], format = "markdown")
+kable(table.4[,3:4], format = "markdown")
+kable(table.4[,5:6], format = "markdown")
 ```

### 2019-06-02 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 
-## Conclusion (Needs update with code snippets) 
-Model transfer from the metropolitan sample to the non-metropolitan sample led to an increased mistriage rate of ...
-(95% UI ...).
-Mainly due to a marked increase in overtriage.
-Across all samples, 95% UI for model transfer comparison included ....
-Model transfer can potentially lead to increased patient mortality and a misappropriation of resources, due to increased undertriage and overtriage respectively.
+## Conclusion
+Transfer of the `r colnames(highest.comp.mistriage)` model lead to an increased mistriage rate of `r highest.comp.mistriage`.
+This observed effect was associated with considerable uncertainty, meaning mistriage might slightly decrease, but would more likely increase, potentially substantially.
+In many other samples, model transfer led to large levels of uncertainty, and in all samples 95% UI included 0.
+In conclusion, model transfer is associated with varying degrees of uncertainty but can potentially lead to increased mistriage, resulting in increased patient mortality and a misappropriation of resources, due to increased undertriage and overtriage, respectively.
 Recognizing the limitations of this study, the authors believe further studies are warranted due to the potential economic costs and patient consequences of poor model transfers.

### 2019-06-02 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 ---
-title: "Mistriage"
+title: "Transferability of clinical prediction models for early trauma care"
 author: "Martin Henriksson"
-date: "190429"
-output: word_document
+date: "190602"
+output: html_document
+bibliography: library.bib
+csl: bmcemerg.csl
 ---
@@
 # Introduction
-Trauma results in millions of deaths annually across the globe, and road traffic injuries are among the top 10 leading causes of death worldwide (1, 2).
-Each year 9% of the world’s deaths are because of trauma, and predictions indicate that this number is likely to increase (3).
+Trauma results in millions of deaths annually across the globe, and road traffic injuries are among the top 10 leading causes of death worldwide [@RN1; @RN21].
+Each year 9% of the world’s deaths are because of trauma, and predictions indicate that this number is likely to increase [@RN42].
 
@@
 Patient data, vital signs and scene information is transferred to the receiving hospital.
-This information is evaluated using a system to determine the level of trauma and prepare adequate resources (4), dictating if a full or limited trauma team is activated (5).
+This information is evaluated using a system to determine the level of trauma and prepare adequate resources [@RN43], dictating if a full or limited trauma team is activated [@RN44].
 

### 2019-06-04 ManuscriptMarkdown.Rmd
Add files via upload

@@
 title: "Transferability of clinical prediction models for early trauma care"
-author: "Martin Henriksson"
-date: "190602"
+author: 
+- Martin Gerdin Wärnberg
+- Martin Henriksson
+- Johanna Berg
+- Dell Saulnier
+date: "June 04, 2019"
+documentclass: "article"
 output: html_document
+header-includes:
+   - \usepackage{hyperref}
 bibliography: library.bib
@@
 ---
-
 ```{r setup, include=FALSE}
@@
 ```
+---
+abstract: |
+  **Background:** Trauma results in millions of deaths annually.
+Trauma care involves many levels of the health care system, and clinical prediction models are often used to triage patients.
+Previous research has indicated transferring these models might adversely affect model performance.
+
+  **Methods:** Patients age 15 or above registered between the years 2011 and 2016 in the SweTrau registry were included.
+Three sets of data were created: High and low volume centres, metropolitan and non-metropolitan centres, multi and single centre data.
+A logistic regression model was developed.
+Model performance was evaluated in all data sets using mistriage rate, undertriage rate and overtriage rate.
+Multiple imputation using chained equation was used to handle missing data.
+Model performance was reported as medians with 95% uncertainty interval across imputed data sets.
+
+  **Results:** `r total.number.of.patients ` patients were included.
+Model validation median mistriage rate with 95% UI ranged from `r lowest.val.mistriage` to `r highest.val.mistriage`.
+and transferred model mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage` across all samples.
+Worst performance after transfer compared to receiving sample validation performance was found in the `r colnames(highest.comp.mistriage)` sample with a median mistriage difference of `r highest.comp.mistriage`.
+
+  **Conclusions:** Model transfer from the `r colnames(highest.comp.mistriage)` to the other sample in the data set lead to an increased mistriage rate.
+Mainly due to a marked increase in overtriage.
+In all samples, model transfer was associated with high levels of uncertainty.
+
+  **Keywords:** Trauma, Clinical prediction model, Transferability, Mistriage, Undertriage, Overtriage
+---
+
 # Introduction
@@
 Such models are continuously being developed and studied, examining different parameters and different contexts.
-Models differ in quality and characteristics, but generally perform well predicting survival (6).
+Models differ in quality and characteristics, but generally perform well predicting survival [@RN45].
 Many of these models are developed in a single standardized setting such as a large metropolitan area or a major trauma centre but are then implemented in different settings, such as non-trauma centres.
@@
 What is not fully understood is how this transfer affects model performance.
-Previous research has shown that model performance in terms of calibration can be adversely affected (7).
+Previous research has shown that model performance in terms of calibration can be adversely affected [@RN27].
 However, the aforementioned research studied model transfers between substantially different settings and did not assess more clinically relevant performance measures, such as misclassification.
@@
 It is estimated that up to 10% of all major trauma patients suffer some form of disability.
-The societal effects of trauma are also substantial; with the loss of work days comparable to those lost from cardiovascular and malignant diseases combined (5).
+The societal effects of trauma are also substantial; with the loss of work days comparable to those lost from cardiovascular and malignant diseases combined [@RN23].
 
@@
 In large, more rural parts of the country, trauma patients are instead transported to the nearest hospital for stabilization.
-Once stabilized, they may then be transported to hospitals with more advanced trauma care capacity (19).
+Once stabilized, they may then be transported to hospitals with more advanced trauma care capacity [@RN34].
 Most hospitals in Sweden use some form of system to categorize the level of trauma.
-These systems are usually based on patient vital signs and injury mechanism, as reported by the EMS, or as registered in the emergency department (16).
+These systems are usually based on patient vital signs and injury mechanism, as reported by the EMS, or as registered in the emergency department [@RN90].
 
@@
 Today 95,5% (52 of 55) of Swedish hospitals record trauma cases in SweTrau.
-Currently the registry includes 55 000 cases, and the board encourages its use in both academic research and more local quality improvement initiatives (20).
+Currently the registry includes 55 000 cases, and the board encourages its use in both academic research and more local quality improvement initiatives [@RN28].
 
 The SweTrau inclusion criteria are: Traumatic event with subsequent activation of hospital trauma protocol, admitted patients with NISS (New Injury Severity Score) > 15 and patients transferred to the hospital within 7 days of traumatic event with NISS > 15.
-SweTrau excludes patients if the only traumatic event is chronic subdural hematoma or if hospital trauma protocol is activated without traumatic event (21).
+SweTrau excludes patients if the only traumatic event is chronic subdural hematoma or if hospital trauma protocol is activated without traumatic event [@RN29].
 
@@
 The eligibility criteria were patients registered in SweTrau, age 15 or above.
-This age was decided as the study aims to study adult trauma and not paediatric trauma which differs in physiology, triage and initial care (22).
-The Swedish guidelines for trauma activation define children as age <15 (16).
-Also, several guidelines and protocols listed in ATLS use age ≥15 as cut-off (9).
+This age was decided as the study aims to study adult trauma and not paediatric trauma which differs in physiology, triage and initial care [@RN46].
+The Swedish guidelines for trauma activation define children as age <15 [@RN90].
+Also, several guidelines and protocols listed in ATLS use age ≥15 as cut-off [@RN43].
 Patient age was obtained from the SweTrau register.
@@
 The clinical prediction models included the predictors SBP, RR and GCS on arrival to hospital.
-The rationale for including these three predictors is that they are part of many established clinical prediction models for early trauma care, such as the Revised Trauma Score (23).
+The rationale for including these three predictors is that they are part of many established clinical prediction models for early trauma care, such as the Revised Trauma Score [@RN35].
 
@@
 ### Study outcome
-ISS > 15 was used as the gold standard to define trauma severity as major trauma, and hence patients with ISS ≤ 15 were considered minor trauma (24).
+ISS > 15 was used as the gold standard to define trauma severity as major trauma, and hence patients with ISS ≤ 15 were considered minor trauma [@RN36].
 Overtriage was defined as the event when a clinical prediction model classified a patient with ISS ≤ 15 as major trauma, and undertriage as the event when a clinical prediction model classified a patient with ISS > 15 as minor trauma.
@@
 The development sample contained 70 events (events being patients who died within 30 days of the trauma) and all non-events (non-events being patient survival 30 days past the trauma) during the same time.
-The rationale for including 70 events is that we needed at least 10 events per free parameter in the logistic regression to obtain stable coefficient estimates (25).
-The validation sample instead contained 100 events and at least 100 non-events, which was suggested as the minimum number by Vergouwe in 2005 for external validation samples (26).
+The rationale for including 70 events is that we needed at least 10 events per free parameter in the logistic regression to obtain stable coefficient estimates [@RN30].
+The validation sample instead contained 100 events and at least 100 non-events, which was suggested as the minimum number by Vergouwe in 2005 for external validation samples [@RN37].
 See figure 1 for example using the high- and low volume centre data set, with high volume centres being those in the top quartile of number of patients registered.
@@
 ### Sequence of analysis
-The programming language R was used for all analyses (27), all progress was uploaded to a GitHub repository (28).
+The programming language R was used for all analyses [@RN31], all progress was uploaded to a GitHub repository [@RN59].
 Analyses were performed in the sequence of model development, model validation and finally model comparison.
@@
 The dependent variable was all cause mortality within 30 days of trauma and independent variables SBP, RR, and GCS modelled as previously described.
-To avoid overfitting the model we used a bootstrap procedure to estimate a linear shrinkage factor that we then applied to the model coefficients (29).
+To avoid overfitting the model we used a bootstrap procedure to estimate a linear shrinkage factor that we then applied to the model coefficients [@RN61].
 The shrunk model was then used to estimate the probability of all cause 30-day mortality in the development sample.
-A gridsearch was performed across estimated probabilities in the development sample to identify the cut-off that optimised overtriage keeping undertriage at less than 5% (11).
+A gridsearch was performed across estimated probabilities in the development sample to identify the cut-off that optimised overtriage keeping undertriage at less than 5% [@RN39].
 This cut-off was then used to classify patients as major or minor trauma.
@@
 ### Missing data
-We used multiple imputation using chained equations, as implemented in the R package mice, to handle missing data (30).
+We used multiple imputation using chained equations, as implemented in the R package mice, to handle missing data [@RN41].
 The number of imputations created for each data set was equal to the highest percentage of missing data in that data set.
@@
 When transferred, the model with the lowest mistriage rate was the `r colnames(lowest.tran.mistriage)` model with a median mistriage rate of `r lowest.tran.mistriage`.
-The model with the highest mistriage rate after transfer was the `r colnames(highest.tran.mistriage)` with a median mistriage rate of `r highest.tran.mistriage`
+The model with the highest mistriage rate after transfer was the `r colnames(highest.tran.mistriage)` with a median mistriage rate of `r highest.tran.mistriage`.
 
@@
 In 2016, Gerdin et al.
-found model transfer to adversely affect model performance in terms of calibration, but that this could be improved by updating the model (18).
+found model transfer to adversely affect model performance in terms of calibration, but that this could be improved by updating the model [@RN27].
 This study did not examine model performance in terms of calibration, neither was the effects of updating the model explored.
@@
 External validation as the evaluation of a clinical prediction model in a setting in which it was not originally developed has been studied more extensively.
-Studies using simulated clinical prediction models based on different predictors showed a decline in model performance when being externally validated (33, 34).
-A study from 2018 examining the external validity of prediction models for coronary artery disease (CAD) (35), showed models significantly underestimating the probability of CAD.
+Studies using simulated clinical prediction models based on different predictors showed a decline in model performance when being externally validated [@RN60; @RN61].
+A study from 2018 examining the external validity of prediction models for coronary artery disease (CAD) [@RN62], showed models significantly underestimating the probability of CAD.
 This study did not find the same dramatic decrease in model performance after transfer.
@@
 This is not optimal, and efforts were made to not deviate in any major way.
-Using GitHub, all code was made public for inspection and to ensure reproducibility (28).
+Using GitHub, all code was made public for inspection and to ensure reproducibility [@RN59].
 

### 2019-06-04 ManuscriptMarkdown.Rmd
Add files via upload

@@
   results.all.data <- WorkTableCreator(result.tables.data.frame)
-  ## Sample characteristics table
-  # table.one <-
-
+  ## Sample characteristics table  (Retrived locally!)
+  library(readxl)
+  table_one_high_low <- read_excel("table one/table one high low.xlsx")
+  table_one_metro_non_metro <- read_excel("table one/table one metro non metro.xlsx")
+  table_one_multi_single <- read_excel("table one/table one multi single.xlsx")
 ## RESULTS
@@
 ```{r, echo = FALSE, results = "asis"} 
-## Table 1 = Sample characteristics table
-##result.tables
+## Table 1
+options(knitr.kable.NA = '')
+kable(table_one_high_low, format = "markdown")
+kable(table_one_metro_non_metro, format = "markdown")
+kable(table_one_multi_single, format = "markdown")
 ``` 
+Abbreviations: GCS glasgow coma scale, SBP systolic blood pressure, RR respiratory rate, ISS injury severity score, NISS new injury severity score
 

### 2019-06-18 MainCodeRun.R
Update main code

@@ -25,7 +25,7 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE) {
     ## Set random seed
     set.seed(-41892)
     ## Import data
-    raw.data <- ImportStudyData("simulated-swetrau-data.csv")
+    raw.data <- ImportStudyData("swetrau-20110101-20160425.csv")
     ## Create study sample from selected variables
     selected.data <- VariableSelection(raw.data)
     ## Select cases with age > 15 or age = NA
@@ -75,7 +75,7 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE) {
 
 }
 start.time <- Sys.time()
-MainCodeRun(test = TRUE, clean.start = TRUE)
+MainCodeRun(test = FALSE, clean.start = FALSE)
 end.time <- Sys.time()
 message("This run took ", difftime(end.time, start.time, units = "hours"), " hours")
 

### 2019-06-18 ManuscriptMarkdown.Rmd

@@
 author: 
-- Martin Gerdin Wärnberg
 - Martin Henriksson
@@
 - Dell Saulnier
-date: "June 04, 2019"
+- Martin Gerdin Wärnberg
+date: "June 18, 2019"
 documentclass: "article"
-output: html_document
+output: word_document
 header-includes:
@@
 abstract: |
-  **Background:** Trauma results in millions of deaths annually.
-Trauma care involves many levels of the health care system, and clinical prediction models are often used to triage patients.
-Previous research has indicated transferring these models might adversely affect model performance.
 
-  **Methods:** Patients age 15 or above registered between the years 2011 and 2016 in the SweTrau registry were included.
-Three sets of data were created: High and low volume centres, metropolitan and non-metropolitan centres, multi and single centre data.
-A logistic regression model was developed.
-Model performance was evaluated in all data sets using mistriage rate, undertriage rate and overtriage rate.
-Multiple imputation using chained equation was used to handle missing data.
-Model performance was reported as medians with 95% uncertainty interval across imputed data sets.
+  **Background:** Trauma results in millions of deaths annually. Clinical
+  prediction models are often used in early trauma care. Previous research has
+  indicated transferring these models might adversely affect model performance.
 
-  **Results:** `r total.number.of.patients ` patients were included.
-Model validation median mistriage rate with 95% UI ranged from `r lowest.val.mistriage` to `r highest.val.mistriage`.
-and transferred model mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage` across all samples.
-Worst performance after transfer compared to receiving sample validation performance was found in the `r colnames(highest.comp.mistriage)` sample with a median mistriage difference of `r highest.comp.mistriage`.
+  **Methods:** Patients age 15 or above registered between the years 2011 and
+  2016 in the Swedish national trauma registry SweTrau were included. Three sets
+  of data were created: High and low volume centres, metropolitan and
+  non-metropolitan centres, and multi and single centre data. Clinical
+  prediction models based on logistic regression were developed in each set of
+  data and transferred within sets. Model performance was evaluated in all data
+  sets using mistriage rate, undertriage rate and overtriage rate. Multiple
+  imputation using chained equation was used to handle missing data. Model
+  performance was reported as medians with 95% uncertainty interval across
+  imputed data sets.
 
-  **Conclusions:** Model transfer from the `r colnames(highest.comp.mistriage)` to the other sample in the data set lead to an increased mistriage rate.
-Mainly due to a marked increase in overtriage.
-In all samples, model transfer was associated with high levels of uncertainty.
+  **Results:** `r total.number.of.patients ` patients were included. Model
+  validation median mistriage rate with 95% UI ranged from `r
+  lowest.val.mistriage` to `r highest.val.mistriage`. and transferred model
+  mistriage rate ranged from `r lowest.tran.mistriage` to `r
+  highest.tran.mistriage` across all samples. Worst performance after transfer
+  compared to receiving sample validation performance was found in the `r
+  colnames(highest.comp.mistriage)` sample with a median mistriage difference of
+  `r highest.comp.mistriage`.
+
+  **Conclusions:** Model transfer from the `r colnames(highest.comp.mistriage)`
+  to the other sample in the data set lead to an increased mistriage
+  rate. Mainly due to a marked increase in overtriage. In all samples, model
+  transfer was associated with high levels of uncertainty.
 
@@
 # Introduction
-Trauma results in millions of deaths annually across the globe, and road traffic injuries are among the top 10 leading causes of death worldwide [@RN1; @RN21].
-Each year 9% of the world’s deaths are because of trauma, and predictions indicate that this number is likely to increase [@RN42].
+Trauma results in millions of deaths annually across the globe, and road traffic
+injuries are among the top 10 leading causes of death worldwide [@RN1;
+@RN21]. Each year 9% of the world’s deaths are because of trauma, and
+predictions indicate that this number is likely to increase [@RN42].
 
-In a typical, high resource setting, the initial management of patients is performed on the scene of the trauma by Emergency Medical Services (EMS).
-Patient data, vital signs and scene information is transferred to the receiving hospital.
-This information is evaluated using a system to determine the level of trauma and prepare adequate resources [@RN43], dictating if a full or limited trauma team is activated [@RN44].
+In a typical, high resource setting, the initial management of patients is
+performed on the scene of the trauma by Emergency Medical Services
+(EMS). Patient data, vital signs and scene information is transferred to the
+receiving hospital. This information is evaluated using a system to determine
+the level of trauma and prepare adequate resources [@RN43], dictating if a full
+or limited trauma team is activated [@RN44].
 
-Systems to determine the level of trauma can be based on clinical prediction models.
-Such models are continuously being developed and studied, examining different parameters and different contexts.
-Models differ in quality and characteristics, but generally perform well predicting survival [@RN45].
-Many of these models are developed in a single standardized setting such as a large metropolitan area or a major trauma centre but are then implemented in different settings, such as non-trauma centres.
+Systems to determine the level of trauma can be based on clinical prediction
+models.  Models differ in quality and characteristics, but generally perform
+well predicting survival [@RN45]. Many of these models are developed in a single
+standardized setting such as a large metropolitan area or a major trauma centre
+but are then implemented in different settings, such as non-trauma centres.
 
-What is not fully understood is how this transfer affects model performance.
-Previous research has shown that model performance in terms of calibration can be adversely affected [@RN27].
-However, the aforementioned research studied model transfers between substantially different settings and did not assess more clinically relevant performance measures, such as misclassification.
+What is not fully understood is how this transfer affects model
+performance. Previous research has shown that model performance in terms of
+calibration can be adversely affected [@RN27]. However, the aforementioned
+research studied model transfers between substantially different settings and
+did not assess more clinically relevant performance measures, such as
+misclassification.
 
-In trauma, misclassification is often referred to as mistriage.
-Triage refers to the classification of severity as minor or major trauma.
-Mistriage can be subdivided into overtriage; incorrect classification of a patient with minor trauma as major trauma, or undertriage; incorrect classification of a patient with major trauma as minor trauma.
-Both are detrimental in terms of patient care and distribution of resources.
+In trauma, misclassification is often referred to as mistriage. Triage refers to
+the classification of severity as minor or major trauma. Mistriage can be
+subdivided into overtriage; incorrect classification of a patient with minor
+trauma as major trauma, or undertriage; incorrect classification of a patient
+with major trauma as minor trauma. Both are detrimental in terms of patient care
+and distribution of resources.
 
-Thus, the effect of model transfers between contexts within the setting of a single health care system, as well as the effect of such transfers on mistriage have not been studied and represent substantial knowledge gaps.
-Therefore, the aim of this study is to assess how transfers of clinical prediction models for early trauma care between different contexts within a single health system affect mistriage rates.
+Thus, the effect of model transfers between contexts within the setting of a
+single health care system, as well as the effect of such transfers on mistriage
+have not been studied and represent substantial knowledge gaps. Therefore, the
+aim of this study is to assess how transfers of clinical prediction models for
+early trauma care between different contexts within a single health system
+affect mistriage rates.
 
@@
 ## Study design
-We conducted a registry-based cohort study.
-Sweden has a nationally encompassing register for registering trauma called SweTrau.
-SweTrau data was used to create clinical prediction models, which were transferred between different contexts to study the effects on mistriage.
+We conducted a registry-based cohort study. Sweden has a nationally encompassing
+register for registering trauma called SweTrau. SweTrau data was used to create
+clinical prediction models, which were transferred between different contexts to
+study the effects on mistriage. The study and analysis plans were made
+publically available before the research was undertaken (Add reference to github
+repo here).
 
 ## Setting
-In Sweden, trauma is the most common cause of death before 44 years of age.
-It is estimated that up to 10% of all major trauma patients suffer some form of disability.
-The societal effects of trauma are also substantial; with the loss of work days comparable to those lost from cardiovascular and malignant diseases combined [@RN23].
+In Sweden, trauma is the most common cause of death before 44 years of age. It
+is estimated that up to 10% of all major trauma patients suffer some form of
+disability. The societal effects of trauma are also substantial; with the loss
+of work days comparable to those lost from cardiovascular and malignant diseases
+combined [@RN23].
 
-The trauma care organization in Sweden is usually clearly defined in major metropolitan areas: trauma patients are transported directly to a predesignated hospital with specific trauma competency.
-In large, more rural parts of the country, trauma patients are instead transported to the nearest hospital for stabilization.
-Once stabilized, they may then be transported to hospitals with more advanced trauma care capacity [@RN34].
-Most hospitals in Sweden use some form of system to categorize the level of trauma.
-These systems are usually based on patient vital signs and injury mechanism, as reported by the EMS, or as registered in the emergency department [@RN90].
+The trauma care organization in Sweden is usually clearly defined in major
+metropolitan areas: trauma patients are transported directly to a predesignated
+hospital with specific trauma competency. In large, more rural parts of the
+country, trauma patients are instead transported to the nearest hospital for
+stabilization. Once stabilized, they may then be transported to hospitals with
+more advanced trauma care capacity [@RN34]. Most hospitals in Sweden use some
+form of system to categorize the level of trauma. These systems are usually
+based on patient vital signs and injury mechanism, as reported by the EMS, or as
+registered in the emergency department [@RN90].
 
-Data was obtained from SweTrau between 2011 and 2016.
-Today 95,5% (52 of 55) of Swedish hospitals record trauma cases in SweTrau.
-Currently the registry includes 55 000 cases, and the board encourages its use in both academic research and more local quality improvement initiatives [@RN28].
+Data was obtained from SweTrau between 2011 and 2016. Today 95,5% (52 of 55) of
+Swedish hospitals record trauma cases in SweTrau. Currently the registry
+includes 55 000 cases, and the board encourages its use in both academic
+research and more local quality improvement initiatives [@RN28].
 
-The SweTrau inclusion criteria are: Traumatic event with subsequent activation of hospital trauma protocol, admitted patients with NISS (New Injury Severity Score) > 15 and patients transferred to the hospital within 7 days of traumatic event with NISS > 15.
-SweTrau excludes patients if the only traumatic event is chronic subdural hematoma or if hospital trauma protocol is activated without traumatic event [@RN29].
+The SweTrau inclusion criteria are: Traumatic event with subsequent activation
+of hospital trauma protocol, admitted patients with NISS (New Injury Severity
+Score) > 15 and patients transferred to the hospital within 7 days of traumatic
+event with NISS > 15. SweTrau excludes patients if the only traumatic event is
+chronic subdural hematoma or if hospital trauma protocol is activated without
+traumatic event [@RN29].
 
 ## Participants 
-The eligibility criteria were patients registered in SweTrau, age 15 or above.
-This age was decided as the study aims to study adult trauma and not paediatric trauma which differs in physiology, triage and initial care [@RN46].
-The Swedish guidelines for trauma activation define children as age <15 [@RN90].
-Also, several guidelines and protocols listed in ATLS use age ≥15 as cut-off [@RN43].
-Patient age was obtained from the SweTrau register.
+The eligibility criteria were patients registered in SweTrau, age 15 or
+above. This age was decided as the study aims to study adult trauma and not
+paediatric trauma which differs in physiology, triage and initial care
+[@RN46]. The Swedish guidelines for trauma activation define children as age <15
+[@RN90]. Also, several guidelines and protocols listed in ATLS use age ≥15 as
+cut-off [@RN43]. Patient age was obtained from the SweTrau register.
 
@@
 ### Model predictors
-The clinical prediction models included the predictors SBP, RR and GCS on arrival to hospital.
-The rationale for including these three predictors is that they are part of many established clinical prediction models for early trauma care, such as the Revised Trauma Score [@RN35].
+The clinical prediction models included the predictors systolic blood pressure
+(SBP), respiratory rate (RR) and Glasgow coma scale (GCS) on arrival to
+hospital. The rationale for including these three predictors is that they are
+part of many established clinical prediction models for early trauma care, such
+as the Revised Trauma Score [@RN35].
 
 ### Model outcome
-The outcome used to develop the clinical prediction models was all cause mortality within 30 days of the trauma.
+The outcome used to develop the clinical prediction models was all cause
+mortality within 30 days of the trauma.
 
 ### Participant characteristics
-To describe the patient cohort we reported age, sex, Injury Severity Score (ISS) and New ISS (NISS).
+To describe the patient cohort we reported age, sex, Injury Severity Score (ISS)
+and New ISS (NISS).
 
 ### Study outcome
-ISS > 15 was used as the gold standard to define trauma severity as major trauma, and hence patients with ISS ≤ 15 were considered minor trauma [@RN36].
-Overtriage was defined as the event when a clinical prediction model classified a patient with ISS ≤ 15 as major trauma, and undertriage as the event when a clinical prediction model classified a patient with ISS > 15 as minor trauma.
-Overtriage rate was defined as the number of overtriaged patients divided by all patients.
-Undertriage rate was defined as the number of undertriaged patients divided by all patients.
-Mistriage rate was defined as the sum of the over- and undertriage rates.
+ISS > 15 was used as the gold standard to define trauma severity as major
+trauma, and hence patients with ISS ≤ 15 were considered minor trauma
+[@RN36]. Overtriage was defined as the event when a clinical prediction model
+classified a patient with ISS ≤ 15 as major trauma, and undertriage as the event
+when a clinical prediction model classified a patient with ISS > 15 as minor
+trauma. Overtriage rate was defined as the number of overtriaged patients
+divided by all patients. Undertriage rate was defined as the number of
+undertriaged patients divided by all patients. Mistriage rate was defined as the
+sum of the over- and undertriage rates.
 
 ## Data sources and measurements
-Model predictors, outcome, participant characteristics and study outcome as outlined above were all obtained from SweTrau.
-The method of measurement for the model predictors is not specified (for example, if SBP is measured using an automated cuff or manually) in the registry entries.
-In Swedish emergency rooms, patient parameters are usually obtained by a registered nurse or assistant nurse and are assumed to be accurate.
-Whether the patient is dead or alive 30 days after trauma is manually entered into the registry locally, by each respective hospital.
-Participant characteristics are obtained from the patient file before being registered in SweTrau.
-NISS and ISS are calculated by hospital personal based on the patient’s injuries.
+Model predictors, outcome, participant characteristics and study outcome as
+outlined above were all obtained from SweTrau. The method of measurement for the
+model predictors is not specified (for example, if SBP is measured using an
+automated cuff or manually) in the registry entries. In Swedish emergency rooms,
+patient parameters are usually obtained by a registered nurse or assistant nurse
+and are assumed to be accurate.  Whether the patient is dead or alive 30 days
+after trauma is manually entered into the registry locally, by each respective
+hospital. Participant characteristics are obtained from the patient file before
+being registered in SweTrau. NISS and ISS are calculated by hospital personal
+based on the patient’s injuries.
 
 ## Bias
-Data analysis was conducted according to a prearranged analysis plan.
-The analysis plan and statistical analysis code was finalised using simulated data.
-These efforts were taken to avoid confirmation bias.
-The analysis plan was reviewed by an experienced statistician and programmer prior to implementation, to ensure objectivity.
-Neither outcome nor variable were blinded when conducting data analysis, which made a structured objective approach important.
+Data analysis was conducted according to a prearranged analysis plan. The
+analysis plan and statistical analysis code was finalised using simulated
+data. These efforts were taken to avoid confirmation bias. The analysis plan was
+reviewed by an experienced statistician and programmer prior to implementation,
+to ensure objectivity. Neither outcome nor variable were blinded when conducting
+data analysis, which made a structured objective approach important.
 
 ## Study size
-All patients matching eligibility criteria listed above.
-Four data sets were used to study the transfer of clinical prediction models, each data set representing a different setting.
-Data set, and sample size considerations are outlined below.
+All patients matching eligibility criteria listed above. Four data sets were
+used to study the transfer of clinical prediction models, each data set
+representing a different setting. Data set, and sample size considerations are
+outlined below.
 
 ## Quantitative variables
-SBP and RR was modelled using restricted cubic splines with four knots whenever possible placed at equally spaced percentiles, and GCS as a continuous linear term.
-When describing the participant characteristics quantitative variables were presented as continuous.
-ISS was presented as dichotomous using ISS > 15 as the cutoff.
+SBP and RR was modelled using restricted cubic splines with four knots whenever
+possible placed at equally spaced percentiles, and GCS as a continuous linear
+term. When describing the participant characteristics quantitative variables
+were presented as continuous. ISS was presented as dichotomous using ISS > 15 as
+the cutoff.
 
@@
 ### Data sets
-The complete SweTrau cohort was split into four sets of data. High and low volume centres, metropolitan and non-metropolitan centres, multi and single centre data and individual centres.
+The complete SweTrau cohort was split into four sets of data. High and low
+volume centres, metropolitan and non-metropolitan centres, multi and single
+centre data and individual centres.
 
 ### High and low volume centres
-Based on number of patients, two samples were derived from this data set. High volume centres were those with in the top quartile of number of patients received. The rest were low volume centres.
+Based on number of patients, two samples were derived from this data set. High
+volume centres were those with in the top quartile of number of patients
+received. The rest were low volume centres.
 
 ### Metropolitan and non-metropolitan centres
-This data set was also split into two samples.
-The metropolitan sample consisted of greater Stockholm, greater Gothenburg and greater Malmö, as defined by statistics Sweden.
-The other sample was patients from non-metropolitan areas.
+This data set was also split into two samples. The metropolitan sample consisted
+of greater Stockholm, greater Gothenburg and greater Malmö, as defined by
+statistics Sweden. The other sample was patients from non-metropolitan areas.
 
 ### Multi and single centre data
-In this data set multiple samples were created.
-Each centre with large enough sample size to develop and validate a model were their own sample.
-The multi-centre sample consisted of the combined data from all centres.
+In this data set multiple samples were created. Each centre with large enough
+sample size to develop and validate a model were their own sample. The
+multi-centre sample consisted of the combined data from all centres.
 
 ### Individual centres
-This data set was also split into multiple samples.
-Each centre with large enough sample size to develop and validate a model constituted its own sample.
-This data set was removed in the final version of the study due to a lack of centres with a large enough sample size.
+This data set was also split into multiple samples. Each centre with large
+enough sample size to develop and validate a model constituted its own
+sample. This data set was removed in the final version of the study due to a
+lack of centres with a large enough sample size.
 
 ### Development and validation sample
-Each data set included at least two samples.
-The samples were then split into two subsamples using a temporal split based on the date of traumatic event.
-The earlier subsample was the development sample, and the later subsample the validation sample.
-The development sample contained 70 events (events being patients who died within 30 days of the trauma) and all non-events (non-events being patient survival 30 days past the trauma) during the same time.
-The rationale for including 70 events is that we needed at least 10 events per free parameter in the logistic regression to obtain stable coefficient estimates [@RN30].
-The validation sample instead contained 100 events and at least 100 non-events, which was suggested as the minimum number by Vergouwe in 2005 for external validation samples [@RN37].
-See figure 1 for example using the high- and low volume centre data set, with high volume centres being those in the top quartile of number of patients registered.
+Each data set included at least two samples. The samples were then split into
+two subsamples using a temporal split based on the date of trauma. The earlier
+subsample was the development sample, and the later subsample the validation
+sample. The development sample contained 70 events (events being patients who
+died within 30 days of the trauma) and all non-events (non-events being patient
+survival 30 days past the trauma) during the same time. The rationale for
+including 70 events is that we needed at least 10 events per free parameter in
+the logistic regression to obtain stable coefficient estimates [@RN30]. The
+validation sample instead contained 100 events and at least 100 non-events,
+which was suggested as the minimum number by Vergouwe in 2005 for external
+validation samples [@RN37]. See figure 1 for example using the high- and low
+volume centre data set, with high volume centres being those in the top quartile
+of number of patients registered. The minimum sample size of development and
+validation samples was 140 and 200 respectively. We performed analyses only on
+data sets for which all samples included at least the minimum number of
+patients.
 
 ![](fig1.jpg)  
-**Figure 1: High- and low volume centre data set.** Initial split based on number of patients. Temporal split made using date of traumatic event.
-
-The minimum sample size of development and validation samples was 140 and 200 respectively.
-We performed analyses only on data sets for which all samples included at least the minimum number of patients.
+**Figure 1: High- and low volume centre data set.** Initial split based on
+number of patients. Temporal split made using date of traumatic event.
 
 ### Sequence of analysis
-The programming language R was used for all analyses [@RN31], all progress was uploaded to a GitHub repository [@RN59].
-Analyses were performed in the sequence of model development, model validation and finally model comparison.
-These steps were repeated in each data set.
-Below we used the transfer of a model from a high-volume sample to a low volume sample as an example to describe the complete procedure.
+The programming language R was used for all analyses [@RN31], and all code has
+been made publically available [@RN59]. Analyses were performed in the sequence
+of model development, model validation and finally model comparison. These steps
+were repeated in each data set. Below we used the transfer of a model from a
+high-volume sample to a low volume sample as an example to describe the complete
+procedure.
 
 ### Model development
-In the model development step a clinical prediction model was developed in the high-volume centre development sample.
-The model was developed using logistic regression as implemented in the R function glm.
-The dependent variable was all cause mortality within 30 days of trauma and independent variables SBP, RR, and GCS modelled as previously described.
-To avoid overfitting the model we used a bootstrap procedure to estimate a linear shrinkage factor that we then applied to the model coefficients [@RN61].
-The shrunk model was then used to estimate the probability of all cause 30-day mortality in the development sample.
-A gridsearch was performed across estimated probabilities in the development sample to identify the cut-off that optimised overtriage keeping undertriage at less than 5% [@RN39].
-This cut-off was then used to classify patients as major or minor trauma.
+In the model development step a clinical prediction model was developed in the
+high-volume centre development sample. The model was developed using logistic
+regression as implemented in the R function glm. The dependent variable was all
+cause mortality within 30 days of trauma and independent variables SBP, RR, and
+GCS modelled as previously described. To avoid overfitting the model we used a
+bootstrap procedure to estimate a linear shrinkage factor that we then applied
+to the model coefficients [@RN61]. The shrunk model was then used to estimate
+the probability of all cause 30-day mortality in the development sample. A
+gridsearch was performed across estimated probabilities in the development
+sample to identify the cut-off that optimised overtriage keeping undertriage at
+less than 5% [@RN39]. This cut-off was then used to classify patients as major
+or minor trauma.
 
 ### Model validation
-In the model validation step, the model performance was assessed in the high-volume centre validation sample and in the low volume centre validation data.
-First the model was used to estimate the probability of all cause 30-day mortality in each sample.
-Then the probability cutoff identified in the development sample was applied to the validation samples, patients were classified as major or minor trauma, and model performance was estimated.
+In the model validation step, the model performance was assessed in the
+high-volume centre validation sample and in the low volume centre validation
+data. First the model was used to estimate the probability of all cause 30-day
+mortality in each sample. Then the probability cutoff identified in the
+development sample was applied to the validation samples, patients were
+classified as major or minor trauma, and model performance was estimated.
 
 ### Model comparison
-Finally, in the model comparison step, the difference in model performance between the high and low volume centre validation samples was calculated.
-Empirical bootstrap was used to estimate 95% confidence intervals (CI) around performance and differences in performance estimates.
-Both bootstrap procedures used 1000 bootstrap samples drawn with replacement of the same size as the original samples.
-The three steps of model development, model validation, and model comparison were repeated in all four sets of data.
+Finally, in the model comparison step, the difference in model performance
+between the high and low volume centre validation samples was
+calculated. Empirical bootstrap was used to estimate 95% confidence intervals
+(CI) around performance and differences in performance estimates. Both bootstrap
+procedures used 1000 bootstrap samples drawn with replacement of the same size
+as the original samples. The three steps of model development, model validation,
+and model comparison were repeated in all four sets of data.
 
 ### Performance measures
-Model performance was assessed in terms of over-, under-, and mistriage rates as defined above.
+Model performance was assessed in terms of over-, under-, and mistriage rates as
+defined above.
 
 ### Missing data
-We used multiple imputation using chained equations, as implemented in the R package mice, to handle missing data [@RN41].
-The number of imputations created for each data set was equal to the highest percentage of missing data in that data set.
-Quantitative variables were imputed using predictive mean matching and qualitative variables were imputed using logistic regression.
-SBP and RR was transformed as restricted cubic splines before imputation and imputed as just another variable.
-All analyses outlined above were then conducted separately in each imputed dataset.
-We present the combined results as the median point estimate across imputations along with an empirical bootstrap of the 25th and 75th percentiles across imputations, i.e.
-the lower bound of the presented interval is the lower bound of a 95% CI of the 25th percentile and the upper bound is the upper bound of a 95% CI of the 75th percentile.
-This combined CI was referred to as an Uncertainty Interval (UI) and was used to express the added uncertainty associated with the imputation procedure to handle missing data, as such it is more conservative than a standard 95% CI.
+We used multiple imputation using chained equations, as implemented in the R
+package mice, to handle missing data [@RN41]. The number of imputations created
+for each data set was equal to the highest percentage of missing data in that
+data set. Quantitative variables were imputed using predictive mean matching and
+qualitative variables were imputed using logistic regression. SBP and RR was
+transformed as restricted cubic splines before imputation and imputed as just
+another variable. All analyses outlined above were then conducted separately in
+each imputed dataset. We present the combined results as the median point
+estimate across imputations along with an empirical bootstrap of the 25th and
+75th percentiles across imputations, i.e. the lower bound of the presented
+interval is the lower bound of a 95% CI of the 25th percentile and the upper
+bound is the upper bound of a 95% CI of the 75th percentile. This combined CI
+was referred to as an Uncertainty Interval (UI) and was used to express the
+added uncertainty associated with the imputation procedure to handle missing
+data, as such it is more conservative than a standard 95% CI.
 
 ## Ethical considerations
-This study is approved by the regional ethics review board in Stockholm, Sweden.
-Ethical review numbers are 2015/426-31 and 2016/461-32.
-All patients included recived letters detainling SweTrau inclusion, and the possibility of their data being used for scientific purposes.
-It was assumed that the patients who consented to research want to contribute to this improvement of care, and as thus by performing this study the authors hope to honour their wishes.
+This study is approved by the regional ethics review board in Stockholm,
+Sweden. Ethical review numbers are 2015/426-31 and 2016/461-32. All patients
+included recived letters detainling SweTrau inclusion, and the possibility of
+their data being used for scientific purposes. It was assumed that the patients
+who consented to research want to contribute to this improvement of care, and as
+thus by performing this study the authors hope to honour their wishes.
 
 # Results
-We analysed data from `r total.number.of.patients ` trauma patients to investigate the effects of transfer on clinical prediction models (Table 1).
-The total number of missing observations across all variables was `r total.number.of.na ` in the entire study cohort.
-The sample with the highest percentage of missing observations was the `r highest.percent.missing.name ` sample with `r highest.percent.missing.number `% incomplete observations.
-The variable with the highest number of missing values was `r name.variable.na `, with `r number.variable.na ` missing values, which constituted `r round(percent.variable.na) `% of the total values for this variable.
-The percentages of missing values for the other model predictors, GCS and SBP was `r round(gcs.na$percentage)`% and `r round(sbp.na$percentage)`% respectively.
+We analysed data from `r total.number.of.patients ` trauma patients to
+investigate the effects of transfer on clinical prediction models (Table 1). The
+total number of missing observations across all variables was `r
+total.number.of.na ` in the entire study cohort. The sample with the highest
+percentage of missing observations was the `r highest.percent.missing.name `
+sample with `r highest.percent.missing.number `% incomplete observations. The
+variable with the highest number of missing values was `r name.variable.na `,
+with `r number.variable.na ` missing values, which constituted `r
+round(percent.variable.na) `% of the total values for this variable. The
+percentages of missing values for the other model predictors, GCS and SBP was `r
+round(gcs.na$percentage)`% and `r round(sbp.na$percentage)`% respectively.
 
@@
 ## Development and validation
-During model development, the number of imputations used for each data set was `r high.low.imp`, `r metro.non.imp` and `r multi.single.imp` for the high and low volume data set, the metropolitan and non-metropolitan data set and the multi and single centre data set respectively.
+During model development, the number of imputations used for each data set was
+`r high.low.imp`, `r metro.non.imp` and `r multi.single.imp` for the high and
+low volume data set, the metropolitan and non-metropolitan data set and the
+multi and single centre data set respectively.
 
-Table 2 shows model validation performance.
-The model with the lowest validation mistriage rate was the `r colnames(lowest.val.mistriage)` model with a median mistriage rate with 95% UI of `r lowest.val.mistriage`.
-Performance in terms of undertriage and overtriage rate for the same model was `r val.undertriage`, and `r val.overtriage` respectively.
-Worst validation performance i.e.
-highest mistriage rate was found in the `r colnames(highest.val.mistriage)` sample, with a model median mistriage rate of `r highest.val.mistriage`.
+Table 2 shows model validation performance. The model with the lowest validation
+mistriage rate was the `r colnames(lowest.val.mistriage)` model with a median
+mistriage rate with 95% UI of `r lowest.val.mistriage`. Performance in terms of
+undertriage and overtriage rate for the same model was `r val.undertriage`, and
+`r val.overtriage` respectively. Worst validation performance i.e. highest
+mistriage rate was found in the `r colnames(highest.val.mistriage)` sample, with
+a model median mistriage rate of `r highest.val.mistriage`.
 
@@
 ## Comparison
-Model performance after transfer was determined for each model being transferred to the other sample in the data set, in all data sets (Table 3).
-When transferred, the model with the lowest mistriage rate was the `r colnames(lowest.tran.mistriage)` model with a median mistriage rate of `r lowest.tran.mistriage`.
-The model with the highest mistriage rate after transfer was the `r colnames(highest.tran.mistriage)` with a median mistriage rate of `r highest.tran.mistriage`.
+Model performance after transfer was determined for each model being transferred
+to the other sample in the data set, in all data sets (Table 3). When
+transferred, the model with the lowest mistriage rate was the `r
+colnames(lowest.tran.mistriage)` model with a median mistriage rate of `r
+lowest.tran.mistriage`. The model with the highest mistriage rate after transfer
+was the `r colnames(highest.tran.mistriage)` with a median mistriage rate of `r
+highest.tran.mistriage`.
 
@@
 
-Table 4 shows the transferred model performance when compared to the validation performance in the sample reciving the transferred model, e.g.
-high volume transfer performance minus low volume validation performance.
+Table 4 shows the transferred model performance when compared to the validation
+performance in the sample reciving the transferred model, e.g. high volume
+transfer performance minus low volume validation performance.
 
-The highest mistriage rate when compared to the reciveing sample validation model performance was found in the `r colnames(highest.comp.mistriage)` model, with a median mistriage rate difference of `r highest.comp.mistriage`, meaning this model performed worse when transferred to the other sample in the same data set.
-In clinical terms, this model transfer means that among 100 trauma patients, `r abs(worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median) * 100` more patients would be wrongly classified as major or minor trauma.
+The highest mistriage rate when compared to the reciveing sample validation
+model performance was found in the `r colnames(highest.comp.mistriage)` model,
+with a median mistriage rate difference of `r highest.comp.mistriage`, meaning
+this model performed worse when transferred to the other sample in the same data
+set. In clinical terms, this model transfer means that among 100 trauma
+patients, `r
+abs(worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median) *
+100` more patients would be wrongly classified as major or minor trauma.
 
@@
 # Discussion
-This study aimed to assess how transfers of clinical prediction models for early trauma care between different contexts within a single health system affect mistriage rates.
-The most notable effects on model performance following model transfer were seen after transferring the `r colnames(highest.comp.mistriage)` model.
-The mistriage rate increased by `r substr(format(highest.comp.mistriage), 1, 5)` when using the `r colnames(highest.comp.mistriage)` model in the other sample in the same data set.
-Mainly contributing to this increase in mistriage was the marked increase in overtriage when performing this model transfer.
-For this transfer the 95% UI ranged from `r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating that our findings are compatible with a `r UI.top.bot[1]` decrease to a `r UI.top.bot[2]` increase in mistriage.
-The negative 95% UI of `r UI.top.bot[1]` would signifiy an improvement in model performance.
-The higher 95% UI of `r UI.top.bot[2]` would mean a marked increase in model mistriage, compared to the point estimate of `r substr(format(highest.comp.mistriage), 1, 5)`.
+This study aimed to assess how transfers of clinical prediction models for early
+trauma care between different contexts within a single health system affect
+mistriage rates. The most notable effects on model performance following model
+transfer were seen after transferring the `r colnames(highest.comp.mistriage)`
+model. The mistriage rate increased by `r substr(format(highest.comp.mistriage),
+1, 5)` when using the `r colnames(highest.comp.mistriage)` model in the other
+sample in the same data set. Mainly contributing to this increase in mistriage
+was the marked increase in overtriage when performing this model transfer. For
+this transfer the 95% UI ranged from `r UI.top.bot[1]` to `r UI.top.bot[2]`,
+indicating that our findings are compatible with a `r UI.top.bot[1]` decrease to
+a `r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r
+UI.top.bot[1]` would signifiy an improvement in model performance. The higher
+95% UI of `r UI.top.bot[2]` would mean a marked increase in model mistriage,
+compared to the point estimate of `r substr(format(highest.comp.mistriage), 1,
+5)`.
 
-The transfer of clinical prediction models in trauma care has not previously been extensively studied.
-In 2016, Gerdin et al.
-found model transfer to adversely affect model performance in terms of calibration, but that this could be improved by updating the model [@RN27].
-This study did not examine model performance in terms of calibration, neither was the effects of updating the model explored.
+The transfer of clinical prediction models in trauma care has not previously
+been extensively studied. In 2016, Gerdin et al. found model transfer to
+adversely affect model performance in terms of calibration, but that this could
+be improved by updating the model [@RN27]. This study did not examine model
+performance in terms of calibration, neither was the effects of updating the
+model explored.
 
-External validation as the evaluation of a clinical prediction model in a setting in which it was not originally developed has been studied more extensively.
-Studies using simulated clinical prediction models based on different predictors showed a decline in model performance when being externally validated [@RN60; @RN61].
-A study from 2018 examining the external validity of prediction models for coronary artery disease (CAD) [@RN62], showed models significantly underestimating the probability of CAD.
-This study did not find the same dramatic decrease in model performance after transfer.
-However, the combined results of this study, the studies on model transfer and the studies on external validation warrants further caution when transferring models between different settings.
+External validation as the evaluation of a clinical prediction model in a
+setting in which it was not originally developed has been studied more
+extensively. Studies using simulated clinical prediction models based on
+different predictors showed a decline in model performance when being externally
+validated [@RN60; @RN61]. A study from 2018 examining the external validity of
+prediction models for coronary artery disease (CAD) [@RN62], showed models
+significantly underestimating the probability of CAD. This study did not find
+the same dramatic decrease in model performance after transfer. However, the
+combined results of this study, the studies on model transfer and the studies on
+external validation warrants further caution when transferring models between
+different settings.
 
 ## Strengths and limitations
-This study design realistically reflects potential transfers: e.g.
-clinical prediction models for trauma are usually developed in large metropolitan centres and then transferred as clinical recommendations to minor non-metropolitan centres in which traumatic injuries are not as common.
+This study design realistically reflects potential transfers: e.g. clinical
+prediction models for trauma are usually developed in large metropolitan centres
+and then transferred as clinical recommendations to minor non-metropolitan
+centres in which traumatic injuries are not as common.
 
-Originally all single centres with a valid number of events were to constitute one data set, but when performing the analysis only one centre provided a sufficient number of events.
-Leading to the loss of the fourth data set: Individual centres, and the comparison between these.
-Possibly, the results this data set might have provided results that could reflect realistic situations as major trauma centres regularly exchange data to improve patient care.
+Originally all single centres with a valid number of events were to constitute
+one data set, but when performing the analysis only one centre provided a
+sufficient number of events. Leading to the loss of the fourth data set:
+Individual centres, and the comparison between these. Possibly, the results this
+data set might have provided results that could reflect realistic situations as
+major trauma centres regularly exchange data to improve patient care.
 
-In Sweden major trauma is relatively uncommon, this might lead to inconsistencies across centres when performing SweTrau registration.
-E.g.
-using NISS instead of ISS, registering SBP by category and not by absolute value or not following up on patient survival.
-This might lead to exaggerated performance differences between the settings.
+In Sweden major trauma is relatively uncommon, this might lead to
+inconsistencies across centres when performing SweTrau registration. E.g. using
+NISS instead of ISS, registering SBP by category and not by absolute value or
+not following up on patient survival. This might lead to exaggerated performance
+differences between the settings.
 
-The models in this study used GCS, RR and SBP as predictors for 30-day mortality.
-Potentially, the use of additional predictors such as age, mechanism of injury, or newer predictors as shock index might increase model performance.
-Furthermore, we used 30-day mortality as an outcome, recognizing that late mortality, time in hospital, functional impairment, and morbidity are also important outcome factors in the context of trauma care.
+The models in this study used GCS, RR and SBP as predictors for 30-day
+mortality. Potentially, the use of additional predictors such as age, mechanism
+of injury, or newer predictors as shock index might increase model
+performance. Furthermore, we used 30-day mortality as an outcome, recognizing
+that late mortality, time in hospital, functional impairment, and morbidity are
+also important outcome factors in the context of trauma care.
 
-The predetermined analysis plan was to be followed in a step-by-step fashion, to avoiding confirmation bias.
-Programming challenges and late-time code inclusions required a flexible approach to this analysis plan, sometimes with revisions of previous steps.
-This is not optimal, and efforts were made to not deviate in any major way.
-Using GitHub, all code was made public for inspection and to ensure reproducibility [@RN59].
+The predetermined analysis plan was to be followed in a step-by-step fashion, to
+avoiding confirmation bias. Programming challenges and late-time code inclusions
+required a flexible approach to this analysis plan, sometimes with revisions of
+previous steps. This is not optimal, and efforts were made to not deviate in any
+major way. Using GitHub, all code was made public for inspection and to ensure
+reproducibility [@RN59].
 
 ## Clinical applications
-The results of this study show that mistriage can potentially increase following model transfer, in this case more prominently when transferring from a metropolitan setting to a non-metropolitan setting.
-This supports the view that model transfers should be done with caution, as performance might decline.
+The results of this study show that mistriage can potentially increase following
+model transfer, in this case more prominently when transferring from a
+metropolitan setting to a non-metropolitan setting. This supports the view that
+model transfers should be done with caution, as performance might decline.
 
-Recognizing these risks and updating models accordingly will lead to increased accuracy during trauma patient triage.
-Potentially saving lives by minimizing undertriage, and likely save resources by minimizing overtriage.
-With registries such as SweTrau, analysis of model transfer could be performed continuously to optimize models currently in use and advise which model transfers might be detrimental (or potentially beneficial).
+Recognizing these risks and updating models accordingly will lead to increased
+accuracy during trauma patient triage. Potentially saving lives by minimizing
+undertriage, and likely save resources by minimizing overtriage. With registries
+such as SweTrau, analysis of model transfer could be performed continuously to
+optimize models currently in use and advise which model transfers might be
+detrimental (or potentially beneficial).
 
 ## Future studies
-Studies on model transfer can be performed in different settings, e.g.
-assessing the transferability between different but similar countries, such as Sweden and Norway.
-Also, increasing the number of patients would increase the power of the results, one way of doing this would be to replicate this study with the latest data in the SweTrau, as many hospitals have joined recently leading to an increased number of patients registered.
+Studies on model transfer can be performed in different settings, e.g. assessing
+the transferability between different but similar countries, such as Sweden and
+Norway. Also, increasing the number of patients would increase the power of the
+results, one way of doing this would be to replicate this study with the latest
+data in the SweTrau, as many hospitals have joined recently leading to an
+increased number of patients registered.
 
-Model transfer could potentially be studied using different methods.
-Quite often, as mentioned, clinical models are transferred from a major metropolitan centre to a smaller rural centre.
-Staff opinions considering how applicable this new model is could be obtained using an interview or questionnaire approach Clinical trials using different models could also be performed, however, appropriate ethical considerations are paramount before performing this type of study.
+Model transfer could potentially be studied using different methods. Quite
+often, as mentioned, clinical models are transferred from a major metropolitan
+centre to a smaller rural centre. Staff opinions considering how applicable this
+new model is could be obtained using an interview or questionnaire approach
+Clinical trials using different models could also be performed, however,
+appropriate ethical considerations are paramount before performing this type of
+study.
 
 ## Conclusion
-Transfer of the `r colnames(highest.comp.mistriage)` model lead to an increased mistriage rate of `r highest.comp.mistriage`.
-This observed effect was associated with considerable uncertainty, meaning mistriage might slightly decrease, but would more likely increase, potentially substantially.
-In many other samples, model transfer led to large levels of uncertainty, and in all samples 95% UI included 0.
-In conclusion, model transfer is associated with varying degrees of uncertainty but can potentially lead to increased mistriage, resulting in increased patient mortality and a misappropriation of resources, due to increased undertriage and overtriage, respectively.
-Recognizing the limitations of this study, the authors believe further studies are warranted due to the potential economic costs and patient consequences of poor model transfers.
+Transfer of the `r colnames(highest.comp.mistriage)` model lead to an increased
+mistriage rate of `r highest.comp.mistriage`. This observed effect was
+associated with considerable uncertainty, meaning mistriage might slightly
+decrease, but would more likely increase, potentially substantially. In many
+other samples, model transfer led to large levels of uncertainty, and in all
+samples 95% UI included 0. In conclusion, model transfer is associated with
+varying degrees of uncertainty but can potentially lead to increased mistriage,
+resulting in increased patient mortality and a misappropriation of resources,
+due to increased undertriage and overtriage, respectively. Recognizing the
+limitations of this study, the authors believe further studies are warranted due
+to the potential economic costs and patient consequences of poor model
+transfers.
 

### 2019-06-18 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
   validation median mistriage rate with 95% UI ranged from `r
-  lowest.val.mistriage` to `r highest.val.mistriage`. and transferred model
+  lowest.val.mistriage` to `r highest.val.mistriage`, and transferred model
   mistriage rate ranged from `r lowest.tran.mistriage` to `r
@@
 study the effects on mistriage. The study and analysis plans were made
-publically available before the research was undertaken (Add reference to github
-repo here).
+publically available before the research was undertaken [@RN59].
 
@@
 includes 55 000 cases, and the board encourages its use in both academic
-research and more local quality improvement initiatives [@RN28].
+research and local quality improvement initiatives [@RN28].
 
@@
 cause mortality within 30 days of trauma and independent variables SBP, RR, and
-GCS modelled as previously described. To avoid overfitting the model we used a
-bootstrap procedure to estimate a linear shrinkage factor that we then applied
+GCS modelled as previously described. To avoid overfitting the model a
+bootstrap procedure was used to estimate a linear shrinkage factor that was then applied
 to the model coefficients [@RN61]. The shrunk model was then used to estimate

### 2019-06-22 functions/TableOneCreator.R
Update TableOneCreator.R

@@ -5,7 +5,7 @@
 TableOneCreator <- function(data.sets) {
   
        Table.Variables <- c("pt_age_yrs", "pt_Gender", "ed_gcs_sum", "ed_sbp_value", "ed_rr_value", "res_survival", "ISS", "NISS", "ISS_over_15", "group")
-       lapply(names(data.sets), function(data.set.name) {
+       table.list <- lapply(names(data.sets), function(data.set.name) {
            data.set <- data.sets[[data.set.name]]
            data.set <- lapply(names(data.set), function(sample.name) {
                sample <- data.set[[sample.name]]
@@ -16,10 +16,11 @@ TableOneCreator <- function(data.sets) {
            table.name <- print(paste0("Characteristics_table_of_", data.set.name))
            CreateSampleCharacteristicsTable(study.sample = combined.data.set,
                                             variables = Table.Variables,
-                                            save.to.disk = TRUE,
+                                            save.to.disk = FALSE,
                                             save.to.results = FALSE,
                                             table.name = table.name,
                                             include.overall = FALSE,
                                             group = "group") 
   })
+return(table.list)
 }

### 2019-06-22 functions/RunStudy.R
Update RunStudy.R

@@ -62,8 +62,8 @@ RunStudy <- function(selected.data, boot = FALSE, test = FALSE) {
     data.sets <- lapply(data.sets, MICEImplement, test = test)
     ## If not a bootstrap run, create table one
     if (boot == FALSE) {
-        ## Create sample characteristics tables, and save to disk
-        TableOneCreator(data.sets)
+        ## Create sample characteristics tables, and save to Results
+        Results$table.one.list <- TableOneCreator(data.sets)
     }
     ## Removal of original data (.imp = 0) from data.sets 
     data.sets <- lapply(data.sets, function(sample) lapply(sample, function(x) {

### 2019-06-22 functions/TableOneCreator.R
Update TableOneCreator.R

@@ -6,21 +6,24 @@ TableOneCreator <- function(data.sets) {
   
        Table.Variables <- c("pt_age_yrs", "pt_Gender", "ed_gcs_sum", "ed_sbp_value", "ed_rr_value", "res_survival", "ISS", "NISS", "ISS_over_15", "group")
        table.list <- lapply(names(data.sets), function(data.set.name) {
-           data.set <- data.sets[[data.set.name]]
-           data.set <- lapply(names(data.set), function(sample.name) {
-               sample <- data.set[[sample.name]]
-               sample$group <- sample.name
-               return(sample)
-           })
-           combined.data.set <- do.call(rbind, data.set)
-           table.name <- print(paste0("Characteristics_table_of_", data.set.name))
-           CreateSampleCharacteristicsTable(study.sample = combined.data.set,
-                                            variables = Table.Variables,
-                                            save.to.disk = FALSE,
-                                            save.to.results = FALSE,
-                                            table.name = table.name,
-                                            include.overall = FALSE,
-                                            group = "group") 
+                            data.set <- data.sets[[data.set.name]]
+                            data.set <- lapply(names(data.set), function(sample.name) {
+                                sample <- data.set[[sample.name]]
+                                sample$group <- sample.name
+                                return(sample)
+                            })
+                            combined.data.set <- do.call(rbind, data.set)
+                            table.name <- print(paste0("Characteristics_table_of_", data.set.name))
+                            CreateSampleCharacteristicsTable(study.sample = combined.data.set,
+                                                             variables = Table.Variables,
+                                                             save.to.disk = FALSE,
+                                                             save.to.results = FALSE,
+                                                             table.name = table.name,
+                                                             include.overall = FALSE,
+                                                             group = "group") 
   })
+names(table.list) <- c("Characteristics_table_of_high.volume.vs.low.volume",
+                       "Characteristics_table_of_metropolitan.vs.non.metropolitan",
+                       "Characteristics_table_of_multi.centre.vs.single.centre")
 return(table.list)
 }

### 2019-06-24 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
   **Results:** `r total.number.of.patients ` patients were included. Model
-  validation median mistriage rate with 95% UI ranged from `r
-  lowest.val.mistriage` to `r highest.val.mistriage`, and transferred model
-  mistriage rate ranged from `r lowest.tran.mistriage` to `r
-  highest.tran.mistriage` across all samples. Worst performance after transfer
-  compared to receiving sample validation performance was found in the `r
-  colnames(highest.comp.mistriage)` sample with a median mistriage difference of
-  `r highest.comp.mistriage`.
+  validation median mistriage rate with 95% UI ranged from `r lowest.val.mistriage` 
+  to `r highest.val.mistriage`, and transferred model
+  mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage` 
+  across all samples. Transfer performance when compared to receiving sample validation 
+  performance resulted in subtantial uncertainty in all samples. Worst comparison 
+  performance after transfer was found in the `r colnames(highest.comp.mistriage)` 
+  sample with a median mistriage difference of `r highest.comp.mistriage`.
 
-  **Conclusions:** Model transfer from the `r colnames(highest.comp.mistriage)`
-  to the other sample in the data set lead to an increased mistriage
-  rate. Mainly due to a marked increase in overtriage. In all samples, model
-  transfer was associated with high levels of uncertainty.
+  **Conclusions:** In all samples, model transfer was associated with high 
+  levels of uncertainty. Notably, model transfer from the `r colnames(highest.comp.mistriage)` 
+  sample to the other sample in the data set lead to an increased mistriage
+  rate, mainly due to a marked increase in overtriage.
 
@@
 investigate the effects of transfer on clinical prediction models (Table 1). The
-total number of missing observations across all variables was `r
-total.number.of.na ` in the entire study cohort. The sample with the highest
+total number of missing observations across all variables was `r total.number.of.na ` 
+in the entire study cohort. The sample with the highest
 percentage of missing observations was the `r highest.percent.missing.name `
@@
 variable with the highest number of missing values was `r name.variable.na `,
-with `r number.variable.na ` missing values, which constituted `r
-round(percent.variable.na) `% of the total values for this variable. The
-percentages of missing values for the other model predictors, GCS and SBP was `r
-round(gcs.na$percentage)`% and `r round(sbp.na$percentage)`% respectively.
+with `r number.variable.na ` missing values, which constituted `r round(percent.variable.na) `
+% of the total values for this variable. The percentages 
+of missing values for the other model predictors, GCS and SBP was `r round(gcs.na$percentage)`% 
+and `r round(sbp.na$percentage)`% respectively.
 
@@
 options(knitr.kable.NA = '')
+# kable(results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume , format = "markdown")
+# kable(results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan , format = "markdown")
+# kable(results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre , format = "markdown")
+
 kable(table_one_high_low, format = "markdown")
@@
 to the other sample in the data set, in all data sets (Table 3). When
-transferred, the model with the lowest mistriage rate was the `r
-colnames(lowest.tran.mistriage)` model with a median mistriage rate of `r
-lowest.tran.mistriage`. The model with the highest mistriage rate after transfer
-was the `r colnames(highest.tran.mistriage)` with a median mistriage rate of `r
-highest.tran.mistriage`.
+transferred, the model with the lowest mistriage rate was the `r colnames(lowest.tran.mistriage)` 
+model with a median mistriage rate of `r lowest.tran.mistriage`. The model with 
+the highest mistriage rate after transfer was the `r colnames(highest.tran.mistriage)` 
+with a median mistriage rate of `r highest.tran.mistriage`.
 
@@
 set. In clinical terms, this model transfer means that among 100 trauma
-patients, `r
-abs(worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median) *
-100` more patients would be wrongly classified as major or minor trauma.
+patients, `r abs(worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median) * 100` 
+more patients would be wrongly classified as major or minor trauma.
 
@@
 trauma care between different contexts within a single health system affect
-mistriage rates. The most notable effects on model performance following model
+mistriage rates. In all data sets, model transfer between samples resulted in 
+substantial uncertainty, with some variation between different data sets. Comparing 
+transferred performance with the reciveing sample validation performance resulted in 
+further uncertainty.
+
+The most notable effects on model performance following model
 transfer were seen after transferring the `r colnames(highest.comp.mistriage)`
-model. The mistriage rate increased by `r substr(format(highest.comp.mistriage),
-1, 5)` when using the `r colnames(highest.comp.mistriage)` model in the other
+model. The mistriage rate increased by `r substr(format(highest.comp.mistriage), 1, 5)` 
+when using the `r colnames(highest.comp.mistriage)` model in the other
 sample in the same data set. Mainly contributing to this increase in mistriage
 was the marked increase in overtriage when performing this model transfer. For
-this transfer the 95% UI ranged from `r UI.top.bot[1]` to `r UI.top.bot[2]`,
-indicating that our findings are compatible with a `r UI.top.bot[1]` decrease to
-a `r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r
-UI.top.bot[1]` would signifiy an improvement in model performance. The higher
+this transfer the 95% UI ranged from `r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating 
+a high level of uncertainty and that our findings are compatible with a `r UI.top.bot[1]` 
+decrease to a `r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r UI.top.bot[1]` 
+would signifiy an improvement in model performance. The higher
 95% UI of `r UI.top.bot[2]` would mean a marked increase in model mistriage,
-compared to the point estimate of `r substr(format(highest.comp.mistriage), 1,
-5)`.
+compared to the point estimate of `r substr(format(highest.comp.mistriage), 1, 5)`.
 
@@
 ## Conclusion
-Transfer of the `r colnames(highest.comp.mistriage)` model lead to an increased
-mistriage rate of `r highest.comp.mistriage`. This observed effect was
+Model transfer resulted in large and varying levels of uncertainty, and in all 
+samples 95% UI included 0. Model transfer can therefore be unpredictable, but can 
+potentially lead to increased mistriage, resulting in increased patient mortality 
+and a misappropriation of resources, due to increased undertriage and overtriage, 
+respectively. Transfer of the `r colnames(highest.comp.mistriage)` model lead to an 
+increased mistriage rate of `r highest.comp.mistriage`. This observed effect was also
 associated with considerable uncertainty, meaning mistriage might slightly
-decrease, but would more likely increase, potentially substantially. In many
-other samples, model transfer led to large levels of uncertainty, and in all
-samples 95% UI included 0. In conclusion, model transfer is associated with
-varying degrees of uncertainty but can potentially lead to increased mistriage,
-resulting in increased patient mortality and a misappropriation of resources,
-due to increased undertriage and overtriage, respectively. Recognizing the
-limitations of this study, the authors believe further studies are warranted due
+decrease, but would more likely increase, potentially substantially. Recognizing 
+the limitations of this study, the authors believe further studies are warranted due
 to the potential economic costs and patient consequences of poor model

### 2019-06-24 ManuscriptMarkdown.Rmd

--- 2019-06-18 ManuscriptMarkdown.Rmd
@@
   **Results:** `r total.number.of.patients ` patients were included. Model
-  validation median mistriage rate with 95% UI ranged from `r
-  lowest.val.mistriage` to `r highest.val.mistriage`, and transferred model
-  mistriage rate ranged from `r lowest.tran.mistriage` to `r
-  highest.tran.mistriage` across all samples. Worst performance after transfer
-  compared to receiving sample validation performance was found in the `r
-  colnames(highest.comp.mistriage)` sample with a median mistriage difference of
-  `r highest.comp.mistriage`.
+  validation median mistriage rate with 95% UI ranged from `r lowest.val.mistriage` 
+  to `r highest.val.mistriage`, and transferred model
+  mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage` 
+  across all samples. Worst performance after transfer
+  compared to receiving sample validation performance was found in the `r colnames(highest.comp.mistriage)` 
+  sample with a median mistriage difference of `r highest.comp.mistriage`.
 
@@
 variable with the highest number of missing values was `r name.variable.na `,
-with `r number.variable.na ` missing values, which constituted `r
-round(percent.variable.na) `% of the total values for this variable. The
-percentages of missing values for the other model predictors, GCS and SBP was `r
-round(gcs.na$percentage)`% and `r round(sbp.na$percentage)`% respectively.
+with `r number.variable.na ` missing values, which constituted `r round(percent.variable.na) `
+% of the total values for this variable. The percentages 
+of missing values for the other model predictors, GCS and SBP was `r round(gcs.na$percentage)`
+% and `r round(sbp.na$percentage)`% respectively.
 
@@
 transfer were seen after transferring the `r colnames(highest.comp.mistriage)`
-model. The mistriage rate increased by `r substr(format(highest.comp.mistriage),
-1, 5)` when using the `r colnames(highest.comp.mistriage)` model in the other
+model. The mistriage rate increased by `r substr(format(highest.comp.mistriage), 1, 5)` 
+when using the `r colnames(highest.comp.mistriage)` model in the other
 sample in the same data set. Mainly contributing to this increase in mistriage
@@
 indicating that our findings are compatible with a `r UI.top.bot[1]` decrease to
-a `r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r
-UI.top.bot[1]` would signifiy an improvement in model performance. The higher
+a `r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r UI.top.bot[1]` 
+would signifiy an improvement in model performance. The higher
 95% UI of `r UI.top.bot[2]` would mean a marked increase in model mistriage,
-compared to the point estimate of `r substr(format(highest.comp.mistriage), 1,
-5)`.
+compared to the point estimate of `r substr(format(highest.comp.mistriage), 1, 5)`.
 

### 2019-06-24 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

--- 2019-06-24 ManuscriptMarkdown.Rmd
@@
 - Martin Gerdin Wärnberg
-date: "June 18, 2019"
+date: "June 24, 2019"
 documentclass: "article"
-output: word_document
+output:
+  word_document:
+    reference_docx: transferstyle.docx
 header-includes:
@@
 Trauma results in millions of deaths annually across the globe, and road traffic
-injuries are among the top 10 leading causes of death worldwide [@RN1;
-@RN21]. Each year 9% of the world’s deaths are because of trauma, and
+injuries are among the top 10 leading causes of death [@RN1; @RN21]. 
+Each year 9% of the world’s deaths are because of trauma, and
 predictions indicate that this number is likely to increase [@RN42].
 
-In a typical, high resource setting, the initial management of patients is
-performed on the scene of the trauma by Emergency Medical Services
-(EMS). Patient data, vital signs and scene information is transferred to the
-receiving hospital. This information is evaluated using a system to determine
-the level of trauma and prepare adequate resources [@RN43], dictating if a full
+In a typical, high resource setting, the initial trauma management is
+performed on the scene by Emergency Medical Services
+(EMS). Patient data and vital signs are transferred to the receiving hospital. 
+This information is then evaluated using a system to determine
+the level of trauma, prepare adequate resources [@RN43] and dictate if a full
 or limited trauma team is activated [@RN44].
@@
 well predicting survival [@RN45]. Many of these models are developed in a single
-standardized setting such as a large metropolitan area or a major trauma centre
-but are then implemented in different settings, such as non-trauma centres.
+standardized setting such as a major trauma centre but are then implemented in 
+different settings, such as non-trauma centres.
 
@@
 ## Study design
-We conducted a registry-based cohort study. Sweden has a nationally encompassing
-register for registering trauma called SweTrau. SweTrau data was used to create
+A registry-based cohort study was conducted. Sweden has a nationally encompassing
+trauma register called SweTrau. SweTrau data was used to create
 clinical prediction models, which were transferred between different contexts to
@@
 ## Setting
-In Sweden, trauma is the most common cause of death before 44 years of age. It
-is estimated that up to 10% of all major trauma patients suffer some form of
-disability. The societal effects of trauma are also substantial; with the loss
+In Sweden, trauma is the most common cause of death before 44 years of age. Up 
+to 10% of all major trauma patients suffer some form of disability. 
+The societal effects of trauma are also substantial, with the loss
 of work days comparable to those lost from cardiovascular and malignant diseases
@@
 
-Data was obtained from SweTrau between 2011 and 2016. Today 95,5% (52 of 55) of
-Swedish hospitals record trauma cases in SweTrau. Currently the registry
-includes 55 000 cases, and the board encourages its use in both academic
-research and local quality improvement initiatives [@RN28].
-
-The SweTrau inclusion criteria are: Traumatic event with subsequent activation
-of hospital trauma protocol, admitted patients with NISS (New Injury Severity
-Score) > 15 and patients transferred to the hospital within 7 days of traumatic
-event with NISS > 15. SweTrau excludes patients if the only traumatic event is
-chronic subdural hematoma or if hospital trauma protocol is activated without
-traumatic event [@RN29].
+Today 95,5% (52 of 55) of Swedish hospitals record trauma cases in SweTrau. 
+Currently the registry includes 55 000 cases, and the board encourages its 
+use in both academic research and local quality improvement initiatives [@RN28].
 
 ## Participants 
-The eligibility criteria were patients registered in SweTrau, age 15 or
-above. This age was decided as the study aims to study adult trauma and not
+The eligibility criteria were patients registered in SweTrau between 2011 and 2016, 
+age 15 or above. This age was decided as the study aims to study adult trauma and not
 paediatric trauma which differs in physiology, triage and initial care
-[@RN46]. The Swedish guidelines for trauma activation define children as age <15
-[@RN90]. Also, several guidelines and protocols listed in ATLS use age ≥15 as
-cut-off [@RN43]. Patient age was obtained from the SweTrau register.
+[@RN46]. Also, the Swedish guidelines for trauma activation define children as age <15
+[@RN90]. Patient age was obtained from the SweTrau register.
 
@@
 ### Model predictors
-The clinical prediction models included the predictors systolic blood pressure
-(SBP), respiratory rate (RR) and Glasgow coma scale (GCS) on arrival to
-hospital. The rationale for including these three predictors is that they are
+The clinical prediction models used the predictors systolic blood pressure
+(SBP), respiratory rate (RR) and Glasgow coma scale (GCS), all on arrival to
+hospital. The rationale for these three predictors is that they are
 part of many established clinical prediction models for early trauma care, such
@@
 The outcome used to develop the clinical prediction models was all cause
-mortality within 30 days of the trauma.
+mortality within 30 days of trauma.
 
@@
 outlined above were all obtained from SweTrau. The method of measurement for the
-model predictors is not specified (for example, if SBP is measured using an
-automated cuff or manually) in the registry entries. In Swedish emergency rooms,
+model predictors is not specified in the registry entries. In Swedish emergency rooms,
 patient parameters are usually obtained by a registered nurse or assistant nurse
 and are assumed to be accurate.  Whether the patient is dead or alive 30 days
-after trauma is manually entered into the registry locally, by each respective
-hospital. Participant characteristics are obtained from the patient file before
-being registered in SweTrau. NISS and ISS are calculated by hospital personal
+after trauma is manually entered into the registry locally by each respective
+hospital. NISS and ISS are calculated by hospital personal
 based on the patient’s injuries.
@@
 data. These efforts were taken to avoid confirmation bias. The analysis plan was
-reviewed by an experienced statistician and programmer prior to implementation,
-to ensure objectivity. Neither outcome nor variable were blinded when conducting
-data analysis, which made a structured objective approach important.
+reviewed by an experienced statistician and programmer prior to implementation. 
+Neither outcome nor variable were blinded when conductingdata analysis, which 
+made a structured objective approach important.
 
@@
 ### Data sets
-The complete SweTrau cohort was split into four sets of data. High and low
-volume centres, metropolitan and non-metropolitan centres, multi and single
-centre data and individual centres.
+The complete SweTrau cohort was split into four sets of data, outlined below.
 
@@
 survival 30 days past the trauma) during the same time. The rationale for
-including 70 events is that we needed at least 10 events per free parameter in
+including 70 events is that at least 10 events per free parameter were needed in
 the logistic regression to obtain stable coefficient estimates [@RN30]. The
@@
 validation samples [@RN37]. See figure 1 for example using the high- and low
-volume centre data set, with high volume centres being those in the top quartile
-of number of patients registered. The minimum sample size of development and
+volume centre data set. The minimum sample size of development and
 validation samples was 140 and 200 respectively. We performed analyses only on
@@
 high-volume centre development sample. The model was developed using logistic
-regression as implemented in the R function glm. The dependent variable was all
-cause mortality within 30 days of trauma and independent variables SBP, RR, and
-GCS modelled as previously described. To avoid overfitting the model a
-bootstrap procedure was used to estimate a linear shrinkage factor that was then applied
-to the model coefficients [@RN61]. The shrunk model was then used to estimate
-the probability of all cause 30-day mortality in the development sample. A
-gridsearch was performed across estimated probabilities in the development
-sample to identify the cut-off that optimised overtriage keeping undertriage at
-less than 5% [@RN39]. This cut-off was then used to classify patients as major
-or minor trauma.
+regression. The dependent variable was all cause mortality within 30 days of 
+trauma and independent variables SBP, RR, andGCS modelled as previously described. 
+To avoid overfitting the model a bootstrap procedure was used to estimate a linear 
+shrinkage factor that was then appliedto the model coefficients [@RN61]. The shrunk 
+model was then used to estimate the probability of all cause 30-day mortality in the 
+development sample. A gridsearch was performed across estimated probabilities in 
+the development sample to identify the cut-off that optimised overtriage keeping 
+undertriage at less than 5% [@RN39]. This cut-off was then used to classify patients 
+as major or minor trauma.
 
@@
 ### Missing data
-We used multiple imputation using chained equations, as implemented in the R
-package mice, to handle missing data [@RN41]. The number of imputations created
+We used multiple imputation using chained equations (mice), to handle missing data [@RN41]. 
+The number of imputations created
 for each data set was equal to the highest percentage of missing data in that
@@
 qualitative variables were imputed using logistic regression. SBP and RR was
-transformed as restricted cubic splines before imputation and imputed as just
-another variable. All analyses outlined above were then conducted separately in
+transformed as restricted cubic splines before imputation. All analyses outlined 
+above were then conducted separately in
 each imputed dataset. We present the combined results as the median point
@@
 # kable(results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre , format = "markdown")
-
 kable(table_one_high_low, format = "markdown")
@@
 mistriage rate was the `r colnames(lowest.val.mistriage)` model with a median
-mistriage rate with 95% UI of `r lowest.val.mistriage`. Performance in terms of
+mistriage rate of `r lowest.val.mistriage`. Performance in terms of
 undertriage and overtriage rate for the same model was `r val.undertriage`, and
@@
 Table 4 shows the transferred model performance when compared to the validation
-performance in the sample reciving the transferred model, e.g. high volume
-transfer performance minus low volume validation performance.
-
+performance in the sample reciving the transferred model.
 The highest mistriage rate when compared to the reciveing sample validation
@@
 transfer were seen after transferring the `r colnames(highest.comp.mistriage)`
-model. The mistriage rate increased by `r substr(format(highest.comp.mistriage), 1, 5)` 
-when using the `r colnames(highest.comp.mistriage)` model in the other
-sample in the same data set. Mainly contributing to this increase in mistriage
-was the marked increase in overtriage when performing this model transfer. For
-this transfer the 95% UI ranged from `r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating 
+model, with a mistriage rate increase by `r substr(format(highest.comp.mistriage), 1, 5)`. 
+Mainly contributing to this was a marked increase in overtriage. 
+For this transfer the 95% UI ranged from `r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating 
 a high level of uncertainty and that our findings are compatible with a `r UI.top.bot[1]` 
@@
 The transfer of clinical prediction models in trauma care has not previously
-been extensively studied. In 2016, Gerdin et al. found model transfer to
+been studied extensively . In 2016, Gerdin et al. found model transfer to
 adversely affect model performance in terms of calibration, but that this could
 be improved by updating the model [@RN27]. This study did not examine model
-performance in terms of calibration, neither was the effects of updating the
+performance in terms of calibration, neither were the effects of updating the
 model explored.
@@
 prediction models for coronary artery disease (CAD) [@RN62], showed models
-significantly underestimating the probability of CAD. This study did not find
-the same dramatic decrease in model performance after transfer. However, the
+significantly underestimating the probability of CAD. The
 combined results of this study, the studies on model transfer and the studies on
-external validation warrants further caution when transferring models between
-different settings.
+external validation suggests caution when transferring models.
 
 ## Strengths and limitations
-This study design realistically reflects potential transfers: e.g. clinical
+This study design realistically reflects potential transfers: clinical
 prediction models for trauma are usually developed in large metropolitan centres
 and then transferred as clinical recommendations to minor non-metropolitan
-centres in which traumatic injuries are not as common.
+centres.
 
 Originally all single centres with a valid number of events were to constitute
-one data set, but when performing the analysis only one centre provided a
-sufficient number of events. Leading to the loss of the fourth data set:
-Individual centres, and the comparison between these. Possibly, the results this
-data set might have provided results that could reflect realistic situations as
-major trauma centres regularly exchange data to improve patient care.
+a data set, but only one centre provided a sufficient number of events. Leading to 
+the loss of the data set: Individual centres. This data set might have provided
+results that could reflect realistic situations as major trauma centres regularly 
+exchange data.
 
-In Sweden major trauma is relatively uncommon, this might lead to
-inconsistencies across centres when performing SweTrau registration. E.g. using
-NISS instead of ISS, registering SBP by category and not by absolute value or
-not following up on patient survival. This might lead to exaggerated performance
-differences between the settings.
-
-The models in this study used GCS, RR and SBP as predictors for 30-day
-mortality. Potentially, the use of additional predictors such as age, mechanism
+This study used GCS, RR and SBP as predictors for 30-day
+mortality. Additional predictors such as age, mechanism
 of injury, or newer predictors as shock index might increase model
-performance. Furthermore, we used 30-day mortality as an outcome, recognizing
+performance. Furthermore, we used 30-day mortality as outcome, recognizing
 that late mortality, time in hospital, functional impairment, and morbidity are
@@
 ## Clinical applications
-The results of this study show that mistriage can potentially increase following
-model transfer, in this case more prominently when transferring from a
-metropolitan setting to a non-metropolitan setting. This supports the view that
-model transfers should be done with caution, as performance might decline.
+This study shows that mistriage can potentially increase following
+model transfer, and more prominently in certain cases. This supports 
+the view that model transfers should be done with caution, as 
+performance might decline.
 
@@
 
-Model transfer could potentially be studied using different methods. Quite
-often, as mentioned, clinical models are transferred from a major metropolitan
-centre to a smaller rural centre. Staff opinions considering how applicable this
-new model is could be obtained using an interview or questionnaire approach
-Clinical trials using different models could also be performed, however,
-appropriate ethical considerations are paramount before performing this type of
-study.
+Model transfer could potentially be studied using different methods. Staff opinions 
+considering how applicable this new model is could be obtained using a questionnaire approach. 
+Clinical trials using different models could also be performed, however, appropriate ethical 
+considerations are paramount before performing this type of study.
 
 ## Conclusion
-Model transfer resulted in large and varying levels of uncertainty, and in all 
-samples 95% UI included 0. Model transfer can therefore be unpredictable, but can 
-potentially lead to increased mistriage, resulting in increased patient mortality 
-and a misappropriation of resources, due to increased undertriage and overtriage, 
-respectively. Transfer of the `r colnames(highest.comp.mistriage)` model lead to an 
-increased mistriage rate of `r highest.comp.mistriage`. This observed effect was also
-associated with considerable uncertainty, meaning mistriage might slightly
+Model transfer resulted in large and varying levels of uncertainty. Model transfer 
+can therefore be unpredictable, and potentially lead to increased mistriage, resulting 
+in increased patient mortality and a misappropriation of resources, due to increased 
+undertriage and overtriage, respectively. Transfer of the `r colnames(highest.comp.mistriage)` 
+model lead to an increased mistriage rate of `r highest.comp.mistriage`. This 
+observed effect was also associated with considerable uncertainty, meaning mistriage might slightly
 decrease, but would more likely increase, potentially substantially. Recognizing 

### 2019-06-26 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
   results <- readRDS("output/original.results.Rds")
-  ## Compiled results with UI
-  result.tables <- readRDS("result.tables.Rds")
   ## Compiled results with UI as data frame
@@
   results.all.data <- WorkTableCreator(result.tables.data.frame)
-  ## Sample characteristics table  (Retrived locally!)
-  library(readxl)
-  table_one_high_low <- read_excel("table one/table one high low.xlsx")
-  table_one_metro_non_metro <- read_excel("table one/table one metro non metro.xlsx")
-  table_one_multi_single <- read_excel("table one/table one multi single.xlsx")
   ## RESULTS
@@
 options(knitr.kable.NA = '')
-# kable(results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume , format = "markdown")
-# kable(results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan , format = "markdown")
-# kable(results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre , format = "markdown")
-kable(table_one_high_low, format = "markdown")
-kable(table_one_metro_non_metro, format = "markdown")
-kable(table_one_multi_single, format = "markdown")
+kable(results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume , format = "markdown")
+kable(results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan , format = "markdown")
+kable(results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre , format = "markdown")
 ``` 

### 2019-06-26 functions/CreateSampleCharacteristicsTable.R
Update CreateSampleCharacteristicsTable.R

@@ -50,247 +50,245 @@ CreateSampleCharacteristicsTable <- function(study.sample,
                                              table.caption = "Sample characteristics",
                                              save.to.disk = FALSE,
                                              file.format = "docx") {
-  ## Load required packages
-  library("tableone")
-  library("knitr")
-  ## Error handling
-  if (!is.data.frame(study.sample))
-    stop ("study.sample has to be a data.frame")
-  if (!is.null(data.dictionary))
-    stop ("data.dictionary has to be NULL")
-  if ((!is.character(group) | !IsLength1(group)) & !is.null(group))
-    stop ("group has to be a character vector of length 1 or NULL")
-  if (!is.character(variables) & !is.null(variables))
-    stop ("variables has to be a character vector or NULL")
-  if (!is.character(exclude.variables) & !is.null(exclude.variables))
-    stop ("exclude.variables has to be a character vector or NULL")
-  if (!is.logical(include.overall) | !IsLength1(include.overall))
-    stop ("include.overall has to be a character vector of length 1")
-  if (!is.logical(include.missing) | !IsLength1(include.missing))
-    stop ("include.missing has to be a character vector of length 1")
-  if (!is.logical(include.complete.data) | !IsLength1(include.complete.data))
-    stop ("include.complete.data has to be a character vector of length 1")    
-  if (!is.numeric(digits) | !IsLength1(digits) | digits < 0)
-    stop ("digits has to be a numeric vector of length 1")
-  if (!is.logical(save.to.results) | !IsLength1(save.to.results))
-    stop ("save.to.results has to be a character vector of length 1")
-  if (!is.character(table.name) | !IsLength1(table.name))
-    stop ("table.name has to be a character vector of length 1")
-  if (!is.character(table.caption) | !IsLength1(table.caption))
-    stop ("table.caption has to be a character vector of length 1")        
-  if (!is.logical(save.to.disk) | !IsLength1(save.to.disk))
-    stop ("save.to.disk has to be a character vector of length 1")
-  if (!(file.format %in% c("docx", "rmd", "pdf")) | !IsLength1(file.format))
-    stop ("file.format has to be one of docx, rmd, or pdf")
-  ## Find out if data.frame is multiple imputed data
-  mi <- FALSE
-  if (all(c(".imp", ".id") %in% colnames(study.sample))) {
-    mi <- TRUE
[…]

### 2019-06-26 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 knitr::opts_chunk$set(echo = TRUE)
+
 ## LOAD FUNCTIONS, ENVIROMENT AND TABLES
@@
   results.all.data <- WorkTableCreator(result.tables.data.frame)
+
 ## RESULTS
-  ## Characteristics paragraph
   ## Total patients and NA
@@
   percent.variable.na <- results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$percentage[results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na == number.variable.na]
-  name.variable.na <- row.names( results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na == number.variable.na,] )
+  name.variable.na <- row.names(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[results$NA.info.variable$multi.centre.vs.single.centre$multi.centre$number.of.na == number.variable.na,])
   gcs.na <- (results$NA.info.variable$multi.centre.vs.single.centre$multi.centre[row.names(results$NA.info.variable$multi.centre.vs.single.centre$multi.centre) == "ed_gcs_sum",])
@@
   worst.transfer <- results$original.stats[results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median == max(results$original.stats$Transferred.mistriage.minus.buddy.local.mistriage.median),]
+
 ## DISCUSSION
@@
   UI.top.bot <- unlist(UI.top.bot.list)
-## Table setup
+
+## TABLE SETUP
   # Table 1 = Sample characteristics table
-  #
+  high.low <- results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume
+  names(high.low) <- c("characteristic", "level", "high volume", "low volume")
+  met.nonmet <- results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan
+  names(met.nonmet) <- c("characteristic", "level", "metropolitan", "non metropolitan")
+  multi.single <- results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre
+  names(multi.single) <- c("characteristic", "level", "multi centre", "single centre")
   # Table 2 = result.tables, cropped to only show validation
@@
 options(knitr.kable.NA = '')
-kable(results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume , format = "markdown")
-kable(results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan , format = "markdown")
-kable(results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre , format = "markdown")
+kable(high.low, format = "markdown")
+kable(met.nonmet, format = "markdown")
+kable(multi.single, format = "markdown")
 ``` 

### 2019-07-01 MainCodeRun.R
Add codebook

@@ -39,9 +39,30 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE) {
     selected.data <- Metrocheck(selected.data)
     ## Mark entries as valid (>170 events) individual centres
     selected.data <- IndividualCentreCheck(selected.data)
+    ## Define codebook
+    codebook <- list(pt_age_yrs = list(full.label = "Patient age",
+                                       abbreviated.label = "Age"),
+                     pt_Gender = list(full.label = "Patient gender",
+                                      abbreviated.label = "Gender"),
+                     ed_gcs_sum = list(full.label = "Glasgow coma scale",
+                                       abbreviated.label = "GCS"),
+                     ed_sbp_value = list(full.label = "Systolic blood pressure",
+                                         abbreviated.label = "SBP"),
+                     ed_rr_value = list(full.label = "Respiratory rate",
+                                        abbreviated.label = "RR"),
+                     res_survival = list(full.label = "30 day survival",
+                                         abbreviated.label = "30d survival"),
+                     ISS = list(full.label = "Injury severity score",
+                                abbreviated.label = "ISS"),
+                     NISS = list(full.label = "New injury severity score",
+                                 abbreviated.label = "NISS"),
+                     ISS_over_15 = list(full.label = "Injury severity score over 15",
+                                        abbreviated.label = "ISS>15"),
+                     group = list(full.label = "Group",
+                                  abbreviated.label = ""))
     ## Get original results
     if (clean.start | !original.results.done)
-        RunStudy(selected.data = selected.data , boot = FALSE, test = test)
+        RunStudy(selected.data = selected.data , codebook = codebook, boot = FALSE, test = test)
     ## Create bootstrap samples
     number.of.bootstrap.samples = 1000
     bootstrap.samples <- list()

### 2019-07-01 functions/RunStudy.R

@@ -5,12 +5,13 @@
 #' medians and IQR.  If bootstraps are used, the function will NOT return values
 #' to the Results.env or pring sample characteristics tables.
 #' @param selected.data Data used to run study. No default.
+#' @param codebook Codebook for pretty table printing. Defaults to NULL.
 #' @param boot Logical. If TRUE, the study will run with 1000
 #'     bootstraps. Defaults to FALSE.
 #' @param test Logical. If TRUE only multiple imputed datasets are created and 5
 #'     bootstraps are used to estimate the linear shrinkage factor. Passed to
 #'     MICEImplement and DevelopmentModelCreator. Defaults to FALSE.
-RunStudy <- function(selected.data, boot = FALSE, test = FALSE) {
+RunStudy <- function(selected.data, codebook = NULL, boot = FALSE, test = FALSE) {
 
     ## Error handling
     if (!is.data.frame(selected.data))
@@ -57,20 +58,19 @@ RunStudy <- function(selected.data, boot = FALSE, test = FALSE) {
     }
     ## Create restricted cubic splines
     data.sets <- lapply(data.sets, function(sample) lapply(sample, RCSplineConvert))
-
     ## Impute missing data
     data.sets <- lapply(data.sets, MICEImplement, test = test)
     ## If not a bootstrap run, create table one
     if (boot == FALSE) {
         ## Create sample characteristics tables, and save to Results
-        Results$table.one.list <- TableOneCreator(data.sets)
+        Results$table.one.list <- TableOneCreator(data.sets, codebook = codebook)
     }
     ## Removal of original data (.imp = 0) from data.sets 
     data.sets <- lapply(data.sets, function(sample) lapply(sample, function(x) {
         no.imp.zero <- x[!(x$.imp == "0"),]
         return(no.imp.zero)
     }))
-    
+   
     ## SPLIT DATA SETS BASED ON IMPUTATION
     split.data.sets <- lapply(data.sets, function(sample) lapply(sample, function(x) {
         x <- split(x, x$.imp)
@@ -79,7 +79,6 @@ RunStudy <- function(selected.data, boot = FALSE, test = FALSE) {
     ## DEVELOPMENT AND VALIDATION
     ## Create Development and validation sample, for each imputation
     split.data.sets <- lapply(split.data.sets, function(sample) lapply(sample, function(imp) (lapply(imp, DevValCreator)))) 
-    
     
     ## CLINICAL PREDICTION MODEL
     ## MODEL DEVELOPMENT

### 2019-07-01 functions/TableOneCreator.R

@@ -2,7 +2,8 @@
 #' 
 #' Creates table one, i.e. a sample characteristics table
 #' @param data.sets The data.set list containing the samples.
-TableOneCreator <- function(data.sets) {
+#' @param codebook Codebook for pretty table printing. Defaults to NULL.
+TableOneCreator <- function(data.sets, codebook = NULL) {
   
        Table.Variables <- c("pt_age_yrs", "pt_Gender", "ed_gcs_sum", "ed_sbp_value", "ed_rr_value", "res_survival", "ISS", "NISS", "ISS_over_15", "group")
        table.list <- lapply(names(data.sets), function(data.set.name) {
@@ -16,6 +17,7 @@ TableOneCreator <- function(data.sets) {
                             table.name <- print(paste0("Characteristics_table_of_", data.set.name))
                             CreateSampleCharacteristicsTable(study.sample = combined.data.set,
                                                              variables = Table.Variables,
+                                                             codebook = codebook,
                                                              save.to.disk = FALSE,
                                                              save.to.results = FALSE,
                                                              table.name = table.name,

### 2019-07-01 functions/TableOneCreator.R
Update TableOneCreator.R

@@ -4,26 +4,46 @@
 #' @param data.sets The data.set list containing the samples.
 TableOneCreator <- function(data.sets) {
   
-       Table.Variables <- c("pt_age_yrs", "pt_Gender", "ed_gcs_sum", "ed_sbp_value", "ed_rr_value", "res_survival", "ISS", "NISS", "ISS_over_15", "group")
-       table.list <- lapply(names(data.sets), function(data.set.name) {
-                            data.set <- data.sets[[data.set.name]]
-                            data.set <- lapply(names(data.set), function(sample.name) {
-                                sample <- data.set[[sample.name]]
-                                sample$group <- sample.name
-                                return(sample)
-                            })
-                            combined.data.set <- do.call(rbind, data.set)
-                            table.name <- print(paste0("Characteristics_table_of_", data.set.name))
-                            CreateSampleCharacteristicsTable(study.sample = combined.data.set,
-                                                             variables = Table.Variables,
-                                                             save.to.disk = FALSE,
-                                                             save.to.results = FALSE,
-                                                             table.name = table.name,
-                                                             include.overall = FALSE,
-                                                             group = "group") 
+  Table.Variables <- c("pt_age_yrs", "pt_Gender", "ed_gcs_sum", "ed_sbp_value", "ed_rr_value", "res_survival", "ISS", "NISS", "ISS_over_15", "group")
+  table.list <- lapply(names(data.sets), function(data.set.name) {
+    data.set <- data.sets[[data.set.name]]
+    data.set <- lapply(names(data.set), function(sample.name) {
+      sample <- data.set[[sample.name]]
+      sample$group <- sample.name
+      return(sample)
+    })
+    combined.data.set <- do.call(rbind, data.set)
+    table.name <- print(paste0("Characteristics_table_of_", data.set.name))
+    CreateSampleCharacteristicsTable(study.sample = combined.data.set,
+                                     variables = Table.Variables,
+                                     codebook = list(pt_age_yrs = list(full.label = "Patient age",
+                                                                       abbreviated.label = "Age"),
+                                                     pt_Gender = list(full.label = "Patient gender", 
+                                                                      abbreviated.label = "Gender"),
+                                                     ed_gcs_sum = list(full.label = "Glasgow coma scale", 
+                                                                       abbreviated.label = "GCS"),
+                                                     ed_sbp_value = list(full.label = "Systolic blood pressure", 
+                                                                         abbreviated.label = "SBP"), 
+                                                     ed_rr_value = list(full.label = "Respiratory rate", 
[…]

### 2019-07-01 functions/CreateSampleCharacteristicsTable.R
Update CreateSampleCharacteristicsTable.R

@@ -2,11 +2,10 @@
 #'
 #' Creates the sample characteristics table. Wrapper of TableOne.
 #' @param study.sample Data frame. The study sample. No default.
-#' @param data.dictionary Not currently used.
 #' @param group Character vector of length 1. The grouping variable. If NULL the
 #'     table is not grouped. Defaults to NULL.
 #' @param variables Character vector. The names of variables to include in the
-#'     table. If NULL all variables in data.dictionary is included. Defaults to
+#'     table. If NULL all variables in study.sample are included. Defaults to
 #'     NULL.
 #' @param exclude.variables Character vector. The names of variables to exclude
 #'     from the table. If NULL no variables are excluded. Defaults to NULL.
@@ -23,6 +22,18 @@
 #'     include.overall.
 #' @param digits Numeric vector of length 1 greater than or equal to 0. Number
 #'     of digits to use when rounding table entries. Defaults to 1.
+#' @param codebook A list or NULL. If a list is provided this list will be
+#'     assumed to be a codebook, in the sense that it is a collection of
+#'     variable descriptions. In other words, each entry is a description of a
+#'     variable. The names of the first level entries of the list are assumed to
+#'     be variable names. Must include at least two entries that define the full
+#'     and abbreviated label of each variable respectively. Defaults to NULL.
+#' @param codebook.options A list. The list can only include the entries
+#'     full.label.entry and abbreviated.label.entry, which in turn should should
+#'     be character vectors of length 1 specifying the names of the full and
+#'     abbreviated label entries in the codebook. Defaults to
+#'     list(full.label.entry = "full.label", abbreviated.label.entry =
+#'     "abbreviated.label"). If codebook is NULL this option is ignored.
 #' @param save.to.results Logical vector of length 1. If TRUE the table object
 #'     is saved to a results file on disk using SaveToResults. Defaults to TRUE.
 #' @param table.name Character vector of length 1. The name of the table when
@@ -37,7 +48,6 @@
 #'     "docx".
 #' @export
 CreateSampleCharacteristicsTable <- function(study.sample,
-                                             data.dictionary = NULL,
                                              group = NULL,
                                              variables = NULL,
                                              exclude.variables = NULL,
@@ -45,250 +55,272 @@ CreateSampleCharacteristicsTable <- function(study.sample,
                                              include.missing = TRUE,
[…]

### 2019-07-01 MainCodeRun.R
Update with new version of CreateSampleCharacteristicsTable

@@ -62,7 +62,7 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE) {
                                   abbreviated.label = ""))
     ## Get original results
     if (clean.start | !original.results.done)
-        RunStudy(selected.data = selected.data , codebook = codebook, boot = FALSE, test = test)
+        RunStudy(selected.data = selected.data, codebook = codebook, boot = FALSE, test = test)
     ## Create bootstrap samples
     number.of.bootstrap.samples = 1000
     bootstrap.samples <- list()
@@ -92,12 +92,9 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE) {
     ## Compile results
     CompileResults()
     ## Render results document
-    rmarkdown::render("Results.Rmd")
+    rmarkdown::render("./ManuscriptMarkdown.Rmd")
 
 }
-start.time <- Sys.time()
-MainCodeRun(test = FALSE, clean.start = FALSE)
-end.time <- Sys.time()
-message("This run took ", difftime(end.time, start.time, units = "hours"), " hours")
+MainCodeRun(test = TRUE, clean.start = TRUE)
 
 

### 2019-07-01 functions/CreateSampleCharacteristicsTable.R

@@ -1,296 +1,0 @@
-#' Create sample characteristics table
-#'
-#' Creates the sample characteristics table. Wrapper of TableOne.
-#' @param study.sample Data frame. The study sample. No default.
-#' @param data.dictionary Not currently used.
-#' @param group Character vector of length 1. The grouping variable. If NULL the
-#'     table is not grouped. Defaults to NULL.
-#' @param variables Character vector. The names of variables to include in the
-#'     table. If NULL all variables in data.dictionary is included. Defaults to
-#'     NULL.
-#' @param exclude.variables Character vector. The names of variables to exclude
-#'     from the table. If NULL no variables are excluded. Defaults to NULL.
-#' @param include.overall Logical vector of length 1. If TRUE an overall column
-#'     is included in the tables. Used only if group is not NULL. Defaults to
-#'     TRUE.
-#' @param include.missing Not currently used. Logical vector of length 1. If
-#'     TRUE a column with the number (%) of missing values in each variable is
-#'     included. Defaults to TRUE.
-#' @param include.complete.data Logical vector of length 1. If TRUE the final
-#'     table has two columns, one with complete cases only and one with multiple
-#'     imputed data. Only used if the data is detected as multiple imputed,
-#'     i.e. includes the variables ".imp" AND ".id". Overrides group and
-#'     include.overall.
-#' @param digits Numeric vector of length 1 greater than or equal to 0. Number
-#'     of digits to use when rounding table entries. Defaults to 1.
-#' @param save.to.results Logical vector of length 1. If TRUE the table object
-#'     is saved to a results file on disk using SaveToResults. Defaults to TRUE.
-#' @param table.name Character vector of length 1. The name of the table when
-#'     passed to SaveToResults and saved to disk.. Deafults to
-#'     "sample.characteristics.table".
-#' @param table.caption Character vector of length 1. The table
-#'     caption. Deafults to "Sample characteristics".
-#' @param save.to.disk Logical vector of length 1. If TRUE the table object is
-#'     saved to disk. Defaults to FALSE.
-#' @param file.format Character vector of length 1. The format in which to save
-#'     the table to disk. Has to be one of c("pdf", "rmd", "docx"). Defaults to
-#'     "docx".
-#' @export
-CreateSampleCharacteristicsTable <- function(study.sample,
-                                             data.dictionary = NULL,
[…]

### 2019-07-01 functions/TableOneCreator.R

@@ -15,14 +15,14 @@ TableOneCreator <- function(data.sets, codebook = NULL) {
                             })
                             combined.data.set <- do.call(rbind, data.set)
                             table.name <- print(paste0("Characteristics_table_of_", data.set.name))
-                            CreateSampleCharacteristicsTable(study.sample = combined.data.set,
-                                                             variables = Table.Variables,
-                                                             codebook = codebook,
-                                                             save.to.disk = FALSE,
-                                                             save.to.results = FALSE,
-                                                             table.name = table.name,
-                                                             include.overall = FALSE,
-                                                             group = "group") 
+                            bengaltiger::CreateSampleCharacteristicsTable(study.sample = combined.data.set,
+                                                                          variables = Table.Variables,
+                                                                          codebook = codebook,
+                                                                          save.to.disk = FALSE,
+                                                                          save.to.results = FALSE,
+                                                                          table.name = table.name,
+                                                                          include.overall = FALSE,
+                                                                          group = "group") 
   })
 names(table.list) <- c("Characteristics_table_of_high.volume.vs.low.volume",
                        "Characteristics_table_of_metropolitan.vs.non.metropolitan",

### 2019-07-01 ManuscriptMarkdown.Rmd

@@
   source("functions/WorkTableCreator.R")
-  ## Original results
+  ## Retrive original results
   results <- readRDS("output/original.results.Rds")
-  ## Compiled results with UI as data frame
+  ## Retrive data frame with compiled results
   result.tables.data.frame <- readRDS("result.tables.data.frame.Rds")
-  ## Create table for ease of data retrival
+  ## Create table from result.tables.data.frame for ease of data retrival in document
   results.all.data <- WorkTableCreator(result.tables.data.frame)
@@
   names(multi.single) <- c("characteristic", "level", "multi centre", "single centre")
+   # Create complete table one (test)
+   complete.table.one <- results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume
+   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan$metropolitan)
+   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan$non.metropolitan)
+   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre$multi.centre)
+   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre$single.centre.11001)
+   names(complete.table.one) <- c("characteristic", "level", "high volume", "low volume", "metropolitan", "non metropolitan", "multi centre", "single centre")
   # Table 2 = result.tables, cropped to only show validation
@@
   rownames(table.2) <- c("Mistriage", "Undertriage", "Overtriage")
+   # Create complete table two (test) 
+   complete.table.two <- results.all.data[1:3,]
+   rownames(complete.table.two) <- c("mistriage", "undertriage", "overtriage")
   # Table 3 = Cropped to only show transfer
@@
   of data were created: High and low volume centres, metropolitan and
-  non-metropolitan centres, and multi and single centre data. Clinical
-  prediction models based on logistic regression were developed in each set of
-  data and transferred within sets. Model performance was evaluated in all data
+  non-metropolitan centres, and multi and single centres. Clinical
+  prediction models were developed usinglogistic regression in each data set 
+  and transferred within sets. Model performance was evaluated in all data
   sets using mistriage rate, undertriage rate and overtriage rate. Multiple
@@
   **Results:** `r total.number.of.patients ` patients were included. Model
-  validation median mistriage rate with 95% UI ranged from `r lowest.val.mistriage` 
+  validation mistriage rate ranged from `r lowest.val.mistriage` 
   to `r highest.val.mistriage`, and transferred model
   mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage` 
-  across all samples. Transfer performance when compared to receiving sample validation 
-  performance resulted in subtantial uncertainty in all samples. Worst comparison 
-  performance after transfer was found in the `r colnames(highest.comp.mistriage)` 
-  sample with a median mistriage difference of `r highest.comp.mistriage`.
+  across all samples. Performance after transfer when compared to receiving sample validation 
+  performance resulted in subtantial uncertainty in all samples. Worst performance after 
+  transfer was found in the `r colnames(highest.comp.mistriage)` 
+  sample with a mistriage difference of `r highest.comp.mistriage`.
 
-  **Conclusions:** In all samples, model transfer was associated with high 
-  levels of uncertainty. Notably, model transfer from the `r colnames(highest.comp.mistriage)` 
+  **Conclusions:** In all samples model transfer was associated with high 
+  levels of uncertainty. Notably, transfer from the `r colnames(highest.comp.mistriage)` 
   sample to the other sample in the data set lead to an increased mistriage
-  rate, mainly due to a marked increase in overtriage.
+  rate, primarily due to increased overtriage.
 
@@
 ![](fig1.jpg)  
-**Figure 1: High- and low volume centre data set.** Initial split based on
+**Fig. 1: High- and low volume centre data set.** Initial split based on
 number of patients. Temporal split made using date of traumatic event.
@@
 
+##### page break
+
 **Table 1. Sample characteristics of multiple imputed data.** Data is presented as medians with interquartile range (IQR) or counts with % as applicable.
@@
 options(knitr.kable.NA = '')
-kable(high.low, format = "markdown")
-kable(met.nonmet, format = "markdown")
-kable(multi.single, format = "markdown")
+# kable(high.low, format = "markdown")
+# kable(met.nonmet, format = "markdown")
+# kable(multi.single, format = "markdown")
+kable(complete.table.one, format = "markdown")
 ``` 
-Abbreviations: GCS glasgow coma scale, SBP systolic blood pressure, RR respiratory rate, ISS injury severity score, NISS new injury severity score
+GCS, glasgow coma scale; SBP, systolic blood pressure; RR, respiratory rate; 30DS, 30 day survival; ISS, injury severity score; NISS, new injury severity score
 
@@
 ## Table 2
-kable(table.2[,1:2], format = "markdown")
-kable(table.2[,3:4], format = "markdown")
-kable(table.2[,5:6], format = "markdown")
+# kable(table.2[,1:2], format = "markdown")
+# kable(table.2[,3:4], format = "markdown")
+# kable(table.2[,5:6], format = "markdown")
+kable(complete.table.two, format = "markdown")
 ```

### 2019-07-02 functions/FuncPack.R
Update FuncPack to get specific bengaltiger version using the commit hash

@@ -4,6 +4,9 @@
 #'     but returned. Defaults to FALSE.
 #' @export
 FuncPack <- function(return.only = FALSE) {
+    ## Get specific bengaltiger version
+    library(devtools)
+    install_github("martingerdin/bengaltiger@c360801126f95d1573ed5e766d743648b51cd8f9")
     
     ## Load required packages
     packages <- c("data.table",
@@ -18,7 +21,7 @@ FuncPack <- function(return.only = FALSE) {
 
     if (!return.only)
         for(package in packages) library(package, character.only = TRUE)
-    
+   
     ## Source functions
     files <- list.files("./Functions", full.names = TRUE, pattern = ".R$")
     functions <- gsub("^./Functions/|.R$", "", files)

### 2019-07-02 ManuscriptMarkdown.Rmd
Comment on manuscript

@@
 mistriage rates. In all data sets, model transfer between samples resulted in
-substantial uncertainty, with some variation between different data sets. Comparing 
-transferred performance with the reciveing sample validation performance resulted in 
-further uncertainty.
+substantial uncertainty in its effect on model performance. Comparing
+transferred performance with the reciveing sample validation performance
+resulted in **further uncertainty** _(Not very clear what you mean here, please
+try to clarify.)_.
 
-The most notable effects on model performance following model
-transfer were seen after transferring the `r colnames(highest.comp.mistriage)`
-model, with a mistriage rate increase by `r substr(format(highest.comp.mistriage), 1, 5)`. 
-Mainly contributing to this was a marked increase in overtriage. 
-For this transfer the 95% UI ranged from `r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating 
-a high level of uncertainty and that our findings are compatible with a `r UI.top.bot[1]` 
-decrease to a `r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r UI.top.bot[1]` 
-would signifiy an improvement in model performance. The higher
-95% UI of `r UI.top.bot[2]` would mean a marked increase in model mistriage,
-compared to the point estimate of `r substr(format(highest.comp.mistriage), 1, 5)`.
+The most notable effects on model performance following model transfer were seen
+after transferring the `r colnames(highest.comp.mistriage)` model. This transfer
+resulted in an increase of the mistriage rate by 
+`r substr(format(highest.comp.mistriage), 1, 5)`.  Mainly contributing to this was
+a marked increase in overtriage.  For this transfer the 95% UI ranged from 
+`r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating a high level of uncertainty and
+that our findings are compatible with a `r UI.top.bot[1]` decrease to a 
+`r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r UI.top.bot[1]`
+would signifiy an improvement in model performance. The higher 95% UI of `r
+UI.top.bot[2]` would mean a marked increase in model mistriage, compared to the
+point estimate of `r substr(format(highest.comp.mistriage), 1, 5)`.
 
 The transfer of clinical prediction models in trauma care has not previously
-been studied extensively . In 2016, Gerdin et al. found model transfer to
+been studied extensively. In 2016, Gerdin et al. found model transfer to
 adversely affect model performance in terms of calibration, but that this could
@@
 prediction models for coronary artery disease (CAD) [@RN62], showed models
-significantly underestimating the probability of CAD. The
-combined results of this study, the studies on model transfer and the studies on
-external validation suggests caution when transferring models.
+significantly underestimating the probability of CAD. The combined results of
+this study, the studies on model transfer and the studies on external validation
+suggests caution when transferring models.
 
 ## Strengths and limitations
-This study design realistically reflects potential transfers: clinical
-prediction models for trauma are usually developed in large metropolitan centres
-and then transferred as clinical recommendations to minor non-metropolitan
-centres.
+The main strength of this study is that its design realistically reflects
+potential model transfers: clinical prediction models for trauma are usually
+developed in large metropolitan centres and then transferred as clinical
+recommendations to minor non-metropolitan centres. The results should therefore
+be of practical importance.
 
-Originally all single centres with a valid number of events were to constitute
-a data set, but only one centre provided a sufficient number of events. Leading to 
-the loss of the data set: Individual centres. This data set might have provided
-results that could reflect realistic situations as major trauma centres regularly 
-exchange data.
+**One limitation is that** _(Try to highlight whether you disucss a strength or
+limitation)_ originally all single centres with a valid number of events were to
+constitute a data set, but only one centre provided a sufficient number of
+events. Leading to the loss of the data set: Individual centres. This data set
+might have provided results that could reflect realistic situations as major
+trauma centres regularly exchange data.
 
@@
 ## Clinical applications
-This study shows that mistriage can potentially increase following
-model transfer, and more prominently in certain cases. This supports 
-the view that model transfers should be done with caution, as 
-performance might decline.
+This study shows that mistriage can potentially increase following model
+transfer, and more prominently in certain cases. This supports the view that
+model transfers should be done with caution, as performance might
+deteriorate. Recognizing this risks and updating models accordingly should lead
+to increased accuracy during trauma patient triage. Potentially saving lives by
+minimizing undertriage, and likely save resources by minimizing overtriage. With
+registries such as SweTrau, analysis of model transfer could be performed
+continuously to optimize models currently in use and advise which model
+transfers might be detrimental (or potentially beneficial).
 
-Recognizing these risks and updating models accordingly will lead to increased
-accuracy during trauma patient triage. Potentially saving lives by minimizing
-undertriage, and likely save resources by minimizing overtriage. With registries
-such as SweTrau, analysis of model transfer could be performed continuously to
-optimize models currently in use and advise which model transfers might be
-detrimental (or potentially beneficial).
+## **Future studies** _(Is this heading required by the journal? If not I suggest we remove these two paragraphs)_
+Studies on model transfer can be performed in different settings, e.g. assessing
+the transferability between different but similar countries. Also, increasing
+the number of patients would increase the power of the results, one way of doing
+this would be to replicate this study with the latest data in the SweTrau, as
+many hospitals have joined recently leading to an increased number of patients
+registered.
 
-## Future studies
-Studies on model transfer can be performed in different settings, e.g. assessing
-the transferability between different but similar countries, such as Sweden and
-Norway. Also, increasing the number of patients would increase the power of the
-results, one way of doing this would be to replicate this study with the latest
-data in the SweTrau, as many hospitals have joined recently leading to an
-increased number of patients registered.
-
-Model transfer could potentially be studied using different methods. Staff opinions 
-considering how applicable this new model is could be obtained using a questionnaire approach. 
-Clinical trials using different models could also be performed, however, appropriate ethical 
-considerations are paramount before performing this type of study.
+Model transfer could potentially be studied using different methods. Staff
+opinions considering how applicable this new model is could be obtained using a
+questionnaire approach.  Clinical trials using different models could also be
+performed, however, appropriate ethical considerations are paramount before
+performing this type of study.
 
 ## Conclusion
-Model transfer resulted in large and varying levels of uncertainty. Model transfer 
-can therefore be unpredictable, and potentially lead to increased mistriage, resulting 
-in increased patient mortality and a misappropriation of resources, due to increased 
-undertriage and overtriage, respectively. Transfer of the `r colnames(highest.comp.mistriage)` 
-model lead to an increased mistriage rate of `r highest.comp.mistriage`. This 
-observed effect was also associated with considerable uncertainty, meaning mistriage might slightly
-decrease, but would more likely increase, potentially substantially. Recognizing 
-the limitations of this study, the authors believe further studies are warranted due
+Model transfer resulted in large and varying levels of uncertainty in its effect
+on mistriage. Model transfer can therefore be unpredictable, and potentially
+lead to increased mistriage, **resulting in increased patient mortality and a
+misappropriation of resources, due to increased undertriage and overtriage,
+respectively** _(I suggest you move this part to clinical applications and that
+you try to express these potential effects a bit more carefully)_. Transfer of
+the `r colnames(highest.comp.mistriage)` model lead to an increased mistriage
+rate of `r highest.comp.mistriage`. This observed effect was also associated
+with considerable uncertainty, meaning mistriage might slightly decrease, but
+would more likely increase, potentially substantially. Recognizing the
+limitations of this study, the authors believe further studies are warranted due
 to the potential economic costs and patient consequences of poor model

### 2019-07-03 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 mistriage rates. In all data sets, model transfer between samples resulted in
-substantial uncertainty in its effect on model performance. Comparing
-transferred performance with the reciveing sample validation performance
-resulted in **further uncertainty** _(Not very clear what you mean here, please
-try to clarify.)_.
+substantial uncertainty in its effect on model performance. When comparing
+transferred performance to the reciveing sample validation performance the
+resulting differance also showed considerable uncertainty.
 
@@
 ## Strengths and limitations
+### Strenhths
 The main strength of this study is that its design realistically reflects
@@
 
-**One limitation is that** _(Try to highlight whether you disucss a strength or
-limitation)_ originally all single centres with a valid number of events were to
-constitute a data set, but only one centre provided a sufficient number of
-events. Leading to the loss of the data set: Individual centres. This data set
-might have provided results that could reflect realistic situations as major
-trauma centres regularly exchange data.
+### Limitations
+One limitation is that originally all single centres with a valid number of 
+events were to constitute a data set, but only one centre provided a sufficient 
+number of events. Leading to the loss of the data set: Individual centres. 
+This data set might have provided results that could reflect realistic 
+situations as major trauma centres regularly exchange data.
 
-This study used GCS, RR and SBP as predictors for 30-day
+Furthermore, this study used GCS, RR and SBP as predictors for 30-day
 mortality. Additional predictors such as age, mechanism
@@
 model transfers should be done with caution, as performance might
-deteriorate. Recognizing this risks and updating models accordingly should lead
+deteriorate.
+
+Recognizing this risks and updating models accordingly should lead
 to increased accuracy during trauma patient triage. Potentially saving lives by
@@
 
-## **Future studies** _(Is this heading required by the journal? If not I suggest we remove these two paragraphs)_
-Studies on model transfer can be performed in different settings, e.g. assessing
-the transferability between different but similar countries. Also, increasing
-the number of patients would increase the power of the results, one way of doing
-this would be to replicate this study with the latest data in the SweTrau, as
-many hospitals have joined recently leading to an increased number of patients
-registered.
-
-Model transfer could potentially be studied using different methods. Staff
-opinions considering how applicable this new model is could be obtained using a
-questionnaire approach.  Clinical trials using different models could also be
-performed, however, appropriate ethical considerations are paramount before
-performing this type of study.
-
 ## Conclusion
 Model transfer resulted in large and varying levels of uncertainty in its effect
-on mistriage. Model transfer can therefore be unpredictable, and potentially
-lead to increased mistriage, **resulting in increased patient mortality and a
-misappropriation of resources, due to increased undertriage and overtriage,
-respectively** _(I suggest you move this part to clinical applications and that
-you try to express these potential effects a bit more carefully)_. Transfer of
-the `r colnames(highest.comp.mistriage)` model lead to an increased mistriage
-rate of `r highest.comp.mistriage`. This observed effect was also associated
-with considerable uncertainty, meaning mistriage might slightly decrease, but
-would more likely increase, potentially substantially. Recognizing the
-limitations of this study, the authors believe further studies are warranted due
-to the potential economic costs and patient consequences of poor model
-transfers.
+on mistriage. Therefore model transfer can be unpredictable, potentially leading 
+to increased mistriage, but in some cases also to a decrease in mistriage. 
+Noticeably, transfer of the `r colnames(highest.comp.mistriage)` model lead to an 
+increased mistriage of `r highest.comp.mistriage`. This observed effect was also 
+associated with considerable uncertainty. Recognizing the limitations of this study, 
+the authors believe further studies are warranted dueto the potential economic 
+costs and patient consequences of poor model transfers.
 

### 2019-07-03 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 - Martin Henriksson
+- Dell Saulnier
 - Johanna Berg
-- Dell Saulnier
 - Martin Gerdin Wärnberg

### 2019-07-17 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 - Martin Gerdin Wärnberg
-date: "June 24, 2019"
+date: "July 03, 2019"
 documentclass: "article"
@@
    names(complete.table.one) <- c("characteristic", "level", "high volume", "low volume", "metropolitan", "non metropolitan", "multi centre", "single centre")
+   complete.table.one$characteristic <- c("n", "Age", "Gender", "", "GCS", "SBP", "RR", "30DS", "", "ISS", "NISS", "ISS>15", "")
   # Table 2 = result.tables, cropped to only show validation
@@
 
-  **Background:** Trauma results in millions of deaths annually. Clinical
-  prediction models are often used in early trauma care. Previous research has
-  indicated transferring these models might adversely affect model performance.
+  **Background:** Trauma results in approximately 3,8 million deaths annually. Clinical
+  prediction models are often used for triage in early trauma care. Previous research has
+  indicated transferring these models between different settings might adversely affect 
+  model performance.
 
@@
   non-metropolitan centres, and multi and single centres. Clinical
-  prediction models were developed usinglogistic regression in each data set 
+  prediction models were developed using logistic regression in each data set 
   and transferred within sets. Model performance was evaluated in all data
@@
 # Introduction
-Trauma results in millions of deaths annually across the globe, and road traffic
-injuries are among the top 10 leading causes of death [@RN1; @RN21]. 
+Trauma results in approximately 3,8 million deaths annually across the globe, 
+and road traffic injuries are among the top 10 leading causes of death [@RN1; @RN21]. 
 Each year 9% of the world’s deaths are because of trauma, and
@@
 reviewed by an experienced statistician and programmer prior to implementation. 
-Neither outcome nor variable were blinded when conductingdata analysis, which 
+Neither outcome nor variable were blinded when conducting data analysis, which 
 made a structured objective approach important.
@@
 ### Data sets
-The complete SweTrau cohort was split into four sets of data, outlined below.
+The complete SweTrau cohort was split into four sets of data, each representing a 
+plausible clinical setting for model transfer. These data sets were further divided 
+into samples, this process is outlined below.
 
 ### High and low volume centres
-Based on number of patients, two samples were derived from this data set. High
-volume centres were those with in the top quartile of number of patients
-received. The rest were low volume centres.
+Based on number of patients received and registred in SweTrau for each centre, two 
+samples were derived from this data set. High volume centres were those with in the 
+top quartile of number of patients registred. The rest were low volume centres.
 
@@
 In this data set multiple samples were created. Each centre with large enough
-sample size to develop and validate a model were their own sample. The
-multi-centre sample consisted of the combined data from all centres.
+sample size (outlined below) to develop and validate a model constituted their own 
+sample. The multi-centre sample consisted of the combined data from all centres.
 
@@
 volume centre data set. The minimum sample size of development and
-validation samples was 140 and 200 respectively. We performed analyses only on
-data sets for which all samples included at least the minimum number of
-patients.
+validation samples were 140 and 200 respectively. We performed analyses only on
+data sets for which all samples included at least the minimum sample size.
 
 ![](fig1.jpg)  
-**Fig. 1: High- and low volume centre data set.** Initial split based on
+**Fig. 1: High and low volume centre data set.** Illustrates creation of development 
+and validation sample in the high and low volume data set. Initial split based on
 number of patients. Temporal split made using date of traumatic event.
@@
 regression. The dependent variable was all cause mortality within 30 days of 
-trauma and independent variables SBP, RR, andGCS modelled as previously described. 
+trauma and independent variables SBP, RR, and GCS modelled as previously described. 
 To avoid overfitting the model a bootstrap procedure was used to estimate a linear 
-shrinkage factor that was then appliedto the model coefficients [@RN61]. The shrunk 
+shrinkage factor that was then applied to the model coefficients [@RN61]. The shrunk 
 model was then used to estimate the probability of all cause 30-day mortality in the 
@@
 
-##### page break
-
 **Table 1. Sample characteristics of multiple imputed data.** Data is presented as medians with interquartile range (IQR) or counts with % as applicable.
@@
 options(knitr.kable.NA = '')
-# kable(high.low, format = "markdown")
-# kable(met.nonmet, format = "markdown")
-# kable(multi.single, format = "markdown")
-kable(complete.table.one, format = "markdown")
+kable(high.low, format = "markdown")
+kable(met.nonmet, format = "markdown")
+kable(multi.single, format = "markdown")
+# kable(complete.table.one, format = "markdown")
 ``` 
@@
 ## Table 2
-# kable(table.2[,1:2], format = "markdown")
-# kable(table.2[,3:4], format = "markdown")
-# kable(table.2[,5:6], format = "markdown")
-kable(complete.table.two, format = "markdown")
+kable(table.2[,1:2], format = "markdown")
+kable(table.2[,3:4], format = "markdown")
+kable(table.2[,5:6], format = "markdown")
+# kable(complete.table.two, format = "markdown")
 ```
@@
 `r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r UI.top.bot[1]`
-would signifiy an improvement in model performance. The higher 95% UI of `r
-UI.top.bot[2]` would mean a marked increase in model mistriage, compared to the
+would signifiy an improvement in model performance. The higher 95% UI of 
+`r UI.top.bot[2]` would mean a marked increase in model mistriage, compared to the
 point estimate of `r substr(format(highest.comp.mistriage), 1, 5)`.
@@
 ## Strengths and limitations
-### Strenhths
+### Strengths
 The main strength of this study is that its design realistically reflects
@@
 
-Recognizing this risks and updating models accordingly should lead
+Recognizing these risks and updating models accordingly should lead
 to increased accuracy during trauma patient triage. Potentially saving lives by

### 2019-07-18 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 - Martin Gerdin Wärnberg
-date: "July 03, 2019"
+date: "July 18, 2019"
 documentclass: "article"
@@
 ``` 
-GCS, glasgow coma scale; SBP, systolic blood pressure; RR, respiratory rate; 30DS, 30 day survival; ISS, injury severity score; NISS, new injury severity score
+GCS, glasgow coma scale; SBP, systolic blood pressure; RR, respiratory rate; ISS, injury severity score; NISS, new injury severity score
 

### 2019-07-29 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 - Martin Gerdin Wärnberg
-date: "July 18, 2019"
+date: "July 29, 2019"
 documentclass: "article"
@@
 # Introduction
-Trauma results in approximately 3,8 million deaths annually across the globe, 
-and road traffic injuries are among the top 10 leading causes of death [@RN1; @RN21]. 
-Each year 9% of the world’s deaths are because of trauma, and
-predictions indicate that this number is likely to increase [@RN42].
+Trauma, defined as an injury to living tissue by an extrinsic agent [@RN63], 
+results in approximately 3,8 million deaths annually across the globe [@RN1; @RN21]. 
+Each year 9% of the world’s deaths are because of trauma, with the leading causes being 
+road traffic accidents, suicide and homicide. Predictions indicate the incidence of all 
+these causes are likely to increase by 2030 [@RN42].
 
-In a typical, high resource setting, the initial trauma management is
+In a typical, high-resource setting, the initial management of trauma is
 performed on the scene by Emergency Medical Services
@@
 
-Systems to determine the level of trauma can be based on clinical prediction
+Systems that determine the level of trauma can be based on clinical prediction
 models. Models differ in quality and characteristics, but generally perform
-well predicting survival [@RN45]. Many of these models are developed in a single
-standardized setting such as a major trauma centre but are then implemented in 
+well at predicting survival [@RN45]. Many models are developed in a single,
+standardized setting, such as a major trauma centre but are then implemented in 
 different settings, such as non-trauma centres.
@@
 the classification of severity as minor or major trauma. Mistriage can be
-subdivided into overtriage; incorrect classification of a patient with minor
-trauma as major trauma, or undertriage; incorrect classification of a patient
-with major trauma as minor trauma. Both are detrimental in terms of patient care
-and distribution of resources.
+subdivided into overtriage, the incorrect classification of a patient with minor
+trauma as major trauma, or undertriage, the incorrect classification of a patient
+with major trauma as minor trauma. Mistriage can ultimatly lead to decreased patient 
+survival, but is also detrimental in terms of patient care and distribution of resources.
 
-Thus, the effect of model transfers between contexts within the setting of a
-single health care system, as well as the effect of such transfers on mistriage
-have not been studied and represent substantial knowledge gaps. Therefore, the
-aim of this study is to assess how transfers of clinical prediction models for
-early trauma care between different contexts within a single health system
-affect mistriage rates.
+Thus, the effect of model transfers between care contexts within a single healthcare system, 
+as well as the effect of such transfers on mistriage,have not been studied and represent 
+substantial knowledge gaps. Therefore, the aim of this study is to assess how transfers 
+of clinical prediction models for early (on arrival to hospital) trauma 
+care between different care contexts within a single health system affect mistriage rates.
 
@@
 trauma register called SweTrau. SweTrau data was used to create
-clinical prediction models, which were transferred between different contexts to
+clinical prediction models, which were transferred between different care contexts to
 study the effects on mistriage. The study and analysis plans were made
-publically available before the research was undertaken [@RN59].
+publicly available before the research was undertaken [@RN59].
 
@@
 
-The trauma care organization in Sweden is usually clearly defined in major
+The trauma care in Sweden is usually clearly defined in major
 metropolitan areas: trauma patients are transported directly to a predesignated
-hospital with specific trauma competency. In large, more rural parts of the
+hospital with specific trauma competency. In more rural parts of the
 country, trauma patients are instead transported to the nearest hospital for
@@
 more advanced trauma care capacity [@RN34]. Most hospitals in Sweden use some
-form of system to categorize the level of trauma. These systems are usually
+form of system to categorize the level of trauma, usually
 based on patient vital signs and injury mechanism, as reported by the EMS, or as
@@
 
-Today 95,5% (52 of 55) of Swedish hospitals record trauma cases in SweTrau. 
-Currently the registry includes 55 000 cases, and the board encourages its 
-use in both academic research and local quality improvement initiatives [@RN28].
+Today 52 of the 55 Swedish hospitals record trauma cases in SweTrau. 
+Currently, the registry includes 55 000 cases, and its use is encouraged in 
+both academic research and local quality improvement initiatives [@RN28].
 
 ## Participants 
-The eligibility criteria were patients registered in SweTrau between 2011 and 2016, 
-age 15 or above. This age was decided as the study aims to study adult trauma and not
+The eligibility criteria were adult patients aged 15 years or above registered in SweTrau 
+between 2011 and 2016. The study aimed to assess adult trauma and not
 paediatric trauma which differs in physiology, triage and initial care
-[@RN46]. Also, the Swedish guidelines for trauma activation define children as age <15
-[@RN90]. Patient age was obtained from the SweTrau register.
+[@RN46]. The Swedish guidelines for trauma activation define children as age <15
+[@RN90].
 
 ## Variables
+### Participant characteristics
+To describe the patient cohort we reported age, sex, Injury Severity Score (ISS)
+and New ISS (NISS).
+
 ### Model predictors
@@
 hospital. The rationale for these three predictors is that they are
-part of many established clinical prediction models for early trauma care, such
+part of multiple established clinical prediction models for early trauma care, such
 as the Revised Trauma Score [@RN35].
@@
 mortality within 30 days of trauma.
-
-### Participant characteristics
-To describe the patient cohort we reported age, sex, Injury Severity Score (ISS)
-and New ISS (NISS).
 
@@
 and are assumed to be accurate.  Whether the patient is dead or alive 30 days
-after trauma is manually entered into the registry locally by each respective
-hospital. NISS and ISS are calculated by hospital personal
+after trauma is manually entered into the registry by each respective
+hospital. NISS and ISS are calculated by hospital personnel
 based on the patient’s injuries.
@@
 reviewed by an experienced statistician and programmer prior to implementation. 
-Neither outcome nor variable were blinded when conducting data analysis, which 
-made a structured objective approach important.
+Neither the outcome nor variables were blinded when during analysis, which 
+made a structured approach important, to ensure objectivity.
 
 ## Study size
-All patients matching eligibility criteria listed above. Four data sets were
+All patients matching eligibility criteria were included. Four data sets were
 used to study the transfer of clinical prediction models, each data set
-representing a different setting. Data set, and sample size considerations are
-outlined below.
+representing a different care context. The data sets, and sample size considerations 
+are outlined below.
 
 ## Quantitative variables
-SBP and RR was modelled using restricted cubic splines with four knots whenever
-possible placed at equally spaced percentiles, and GCS as a continuous linear
-term. When describing the participant characteristics quantitative variables
-were presented as continuous. ISS was presented as dichotomous using ISS > 15 as
-the cutoff.
+GCS was modelled as a continuous linear term, and SBP and RR were modelled using 
+restricted cubic splines with four knots whenever possible, placed at equally spaced 
+percentiles. When describing the participant characteristics, quantitative variables 
+were presented as continuous. ISS was presented as dichotomous using ISS > 15 as the cutoff.
 
@@
 ### Data sets
-The complete SweTrau cohort was split into four sets of data, each representing a 
-plausible clinical setting for model transfer. These data sets were further divided 
+The complete SweTrau cohort was split into three overlapping sets of data, each representing a 
+plausible care context for model transfer. These data sets were further divided 
 into samples, this process is outlined below.
 
-### High and low volume centres
-Based on number of patients received and registred in SweTrau for each centre, two 
-samples were derived from this data set. High volume centres were those with in the 
-top quartile of number of patients registred. The rest were low volume centres.
+### Data set 1: High and low volume centres
+Two samples were derived from the SweTrau cohort, based on the number of patients. 
+High volume centres were those within the top quartile of number of patients received. 
+The rest were low volume centres.
 
-### Metropolitan and non-metropolitan centres
-This data set was also split into two samples. The metropolitan sample consisted
-of greater Stockholm, greater Gothenburg and greater Malmö, as defined by
-statistics Sweden. The other sample was patients from non-metropolitan areas.
+### Data set 2: Metropolitan and non-metropolitan centres
+This data set was also split into two samples: a metropolitan sample consisting of 
+patients from greater Stockholm, greater Gothenburg and greater Malmö, as defined by 
+statistics Sweden, and a second sample of patients from non-metropolitan areas.
 
-### Multi and single centre data
-In this data set multiple samples were created. Each centre with large enough
-sample size (outlined below) to develop and validate a model constituted their own 
-sample. The multi-centre sample consisted of the combined data from all centres.
+### Data set 3: Multi- and single centre data
+In this data set, two samples were created. One centre with a large enough 
+sample size to develop and validate a model was its own sample. The multicentre 
+sample consisted of the combined data from all centres.
 
-### Individual centres
-This data set was also split into multiple samples. Each centre with large
-enough sample size to develop and validate a model constituted its own
-sample. This data set was removed in the final version of the study due to a
-lack of centres with a large enough sample size.
+A fourth dataset was initially created, consisting of individual centres that 
+had a large enough sample size to develop and validate a model. However, this 
+data set was removed before analysis due to a lack of centres 
+with a large enough sample size.
 
 ### Development and validation sample
-Each data set included at least two samples. The samples were then split into
+The samples in each data set were split into
 two subsamples using a temporal split based on the date of trauma. The earlier
@@
 the logistic regression to obtain stable coefficient estimates [@RN30]. The
-validation sample instead contained 100 events and at least 100 non-events,
+validation sample contained 100 events and at least 100 non-events,
 which was suggested as the minimum number by Vergouwe in 2005 for external
-validation samples [@RN37]. See figure 1 for example using the high- and low
-volume centre data set. The minimum sample size of development and
+validation samples [@RN37]. Figure 1 shows the development of the samples using 
+the high- and low volume centre data set as an example. The minimum sample size of development and
 validation samples were 140 and 200 respectively. We performed analyses only on
@@
 ![](fig1.jpg)  
-**Fig. 1: High and low volume centre data set.** Illustrates creation of development 
-and validation sample in the high and low volume data set. Initial split based on
-number of patients. Temporal split made using date of traumatic event.
+**Fig. 1: Sample development in the high and low volume centre data set.** Illustrates 
+creation of development and validation sample in the high and low volume data set. 
+Initial split based on number of patients. Temporal split made using date of traumatic event.
 
@@
 The programming language R was used for all analyses [@RN31], and all code has
-been made publically available [@RN59]. Analyses were performed in the sequence
+been made publicly available [@RN59]. Analyses were performed in the sequence
 of model development, model validation and finally model comparison. These steps
-were repeated in each data set. Below we used the transfer of a model from a
-high-volume sample to a low volume sample as an example to describe the complete
-procedure.
+were repeated in each data set. Below, we use the transfer of a model from the
+high volume centre sample to a low volume centre sample as an example to describe 
+the complete procedure.
 
 ### Model development
-In the model development step a clinical prediction model was developed in the
-high-volume centre development sample. The model was developed using logistic
-regression. The dependent variable was all cause mortality within 30 days of 
-trauma and independent variables SBP, RR, and GCS modelled as previously described. 
-To avoid overfitting the model a bootstrap procedure was used to estimate a linear 
+In the model development step, a clinical prediction model was developed in the
+high volume centre development sample. The model was developed using logistic
+regression. The dependent variable was all-cause mortality within 30 days of 
+trauma and the independent variables SBP, RR, and GCS were modelled as previously described. 
+To avoid overfitting the model, a bootstrap procedure was used to estimate a linear 
 shrinkage factor that was then applied to the model coefficients [@RN61]. The shrunk 
-model was then used to estimate the probability of all cause 30-day mortality in the 
+model was used to estimate the probability of all-cause 30-day mortality in the 
 development sample. A gridsearch was performed across estimated probabilities in 
@@
 In the model validation step, the model performance was assessed in the
-high-volume centre validation sample and in the low volume centre validation
-data. First the model was used to estimate the probability of all cause 30-day
-mortality in each sample. Then the probability cutoff identified in the
+high volume centre validation sample and in the low volume centre validation
+sample. First, the model was used to estimate the probability of all-cause 30-day
+mortality in each sample. Then the probability cut-off identified in the
 development sample was applied to the validation samples, patients were
@@
 ### Model comparison
-Finally, in the model comparison step, the difference in model performance
+Finally, the difference in model performance
 between the high and low volume centre validation samples was
@@
 (CI) around performance and differences in performance estimates. Both bootstrap
-procedures used 1000 bootstrap samples drawn with replacement of the same size
+procedures used 1 000 bootstrap samples drawn with replacement of the same size
 as the original samples. The three steps of model development, model validation,
@@
 ### Missing data
-We used multiple imputation using chained equations (mice), to handle missing data [@RN41]. 
+We used multiple imputation using chained equations (mice) to handle missing data [@RN41]. 
 The number of imputations created
@@
 added uncertainty associated with the imputation procedure to handle missing
-data, as such it is more conservative than a standard 95% CI.
+data; as such, it is more conservative than a standard 95% CI.
 
 ## Ethical considerations
-This study is approved by the regional ethics review board in Stockholm,
-Sweden. Ethical review numbers are 2015/426-31 and 2016/461-32. All patients
-included recived letters detainling SweTrau inclusion, and the possibility of
-their data being used for scientific purposes. It was assumed that the patients
-who consented to research want to contribute to this improvement of care, and as
-thus by performing this study the authors hope to honour their wishes.
+The study was approved by the regional ethics review board in Stockholm, Sweden. 
+Ethical review numbers are 2015/426-31 and 2016/461-32. All patients in SweTrau 
+had received letters detailing their inclusion in the database, and the possibility 
+of their data being used for scientific purposes. It was assumed that the patients 
+who consented to the use of their SweTrau data for research want to contribute to 
+the improvement of care, and thus by performing this study the authors hope to 
+honour their wishes.
 
@@
 variable with the highest number of missing values was `r name.variable.na `,
-with `r number.variable.na ` missing values, which constituted `r round(percent.variable.na) `
-% of the total values for this variable. The percentages 
-of missing values for the other model predictors, GCS and SBP was `r round(gcs.na$percentage)`% 
-and `r round(sbp.na$percentage)`% respectively.
+with `r number.variable.na ` missing values, or `r round(percent.variable.na)`% of 
+the total values for this variable. The percentages of missing values for the other 
+model predictors, was `r round(gcs.na$percentage)`% for GCS and `r round(sbp.na$percentage)`% 
+for SBP.
 
@@
 `r high.low.imp`, `r metro.non.imp` and `r multi.single.imp` for the high and
-low volume data set, the metropolitan and non-metropolitan data set and the
-multi and single centre data set respectively.
+low volume data set, the metropolitan and non-metropolitan data set, and the
+multi- and single centre data set, respectively.
 
@@
 undertriage and overtriage rate for the same model was `r val.undertriage`, and
-`r val.overtriage` respectively. Worst validation performance i.e. highest
-mistriage rate was found in the `r colnames(highest.val.mistriage)` sample, with
+`r val.overtriage` respectively. Worst validation performance (i.e. highest
+mistriage rate) was found in the `r colnames(highest.val.mistriage)` sample, with
 a model median mistriage rate of `r highest.val.mistriage`.

### 2019-07-30 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 - Martin Gerdin Wärnberg
-date: "July 29, 2019"
+date: "July 31, 2019"
 documentclass: "article"
@@
 Trauma, defined as an injury to living tissue by an extrinsic agent [@RN63], 
-results in approximately 3,8 million deaths annually across the globe [@RN1; @RN21]. 
-Each year 9% of the world’s deaths are because of trauma, with the leading causes being 
-road traffic accidents, suicide and homicide. Predictions indicate the incidence of all 
-these causes are likely to increase by 2030 [@RN42].
+results in approximatly 458,5 million cases anually across the globe [@RN999], 
+and an estimated 3,8 million deaths [@RN1; @RN21]. Each year 9% of the world’s 
+deaths are because of trauma, with the leading causes being road traffic accidents, 
+suicide and homicide. Predictions indicate the incidence of these causes are 
+likely to increase by 2030 [@RN42].
 
@@
 ## Setting
-In Sweden, trauma is the most common cause of death before 44 years of age. Up 
-to 10% of all major trauma patients suffer some form of disability. 
-The societal effects of trauma are also substantial, with the loss
+In Sweden, trauma is the most common cause of death before 44 years of age, and 
+10% of major trauma patients suffer some form of disability. 
+The societal effects of trauma are substantial, with the loss
 of work days comparable to those lost from cardiovascular and malignant diseases
@@
 ## Participants 
-The eligibility criteria were adult patients aged 15 years or above registered in SweTrau 
+The eligibility criteria were patients aged 15 years or above registered in SweTrau 
 between 2011 and 2016. The study aimed to assess adult trauma and not
@@
 ![](fig1.jpg)  
-**Fig. 1: Sample development in the high and low volume centre data set.** Illustrates 
-creation of development and validation sample in the high and low volume data set. 
-Initial split based on number of patients. Temporal split made using date of traumatic event.
+**Fig. 1: Sample development in the high and low volume centre data set.**  Initial 
+split based on number of patients. Temporal split made using date of traumatic event.
 
@@
 substantial uncertainty in its effect on model performance. When comparing
-transferred performance to the reciveing sample validation performance the
-resulting differance also showed considerable uncertainty.
+transferred performance to the receiving sample validation performance, the
+resulting differences also showed considerable uncertainty.
 
@@
 `r substr(format(highest.comp.mistriage), 1, 5)`.  Mainly contributing to this was
-a marked increase in overtriage.  For this transfer the 95% UI ranged from 
+a marked increase in overtriage.  For this transfer, the 95% UI ranged from 
 `r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating a high level of uncertainty and
@@
 `r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r UI.top.bot[1]`
-would signifiy an improvement in model performance. The higher 95% UI of 
+would signify an improvement in model performance. The higher 95% UI of 
 `r UI.top.bot[2]` would mean a marked increase in model mistriage, compared to the
 point estimate of `r substr(format(highest.comp.mistriage), 1, 5)`.
+
+Due to centralized trauma care, non-metropolitan centres generelly have fewer 
+resources, and less experience reciving trauma patients. They may rely 
+on clinical prediction models more heavily for accurate trauma triage, and an 
+increased mistriage leads to further strain on already limited resources
+due to increased overtriage. Increased undertriage could also possibly lead to
+increased patient mortality
+
+Recognizing these risks and updating models accordingly should lead
+to increased accuracy during trauma patient triage. Potentially saving lives by
+minimizing undertriage, and likely save resources by minimizing overtriage. With
+registries such as SweTrau, analysis of model transfer could be performed
+continuously to optimize models currently in use and advise which model
+transfers might be detrimental (or potentially beneficial).
 
@@
 this study, the studies on model transfer and the studies on external validation
-suggests caution when transferring models.
+would suggest caution when transferring models.
 
@@
 ### Limitations
-One limitation is that originally all single centres with a valid number of 
+Originally, all single centres with a valid number of 
 events were to constitute a data set, but only one centre provided a sufficient 
-number of events. Leading to the loss of the data set: Individual centres. 
-This data set might have provided results that could reflect realistic 
-situations as major trauma centres regularly exchange data.
+number of events, leading to the loss of the individual centres data set. 
+This data set could have provided interesting results reflecting realistic 
+situations.
 
-Furthermore, this study used GCS, RR and SBP as predictors for 30-day
+This study used GCS, RR and SBP as predictors for 30-day
 mortality. Additional predictors such as age, mechanism
-of injury, or newer predictors as shock index might increase model
+of injury, or newer predictors like shock index might increase model
 performance. Furthermore, we used 30-day mortality as outcome, recognizing
@@
 
-The predetermined analysis plan was to be followed in a step-by-step fashion, to
-avoiding confirmation bias. Programming challenges and late-time code inclusions
-required a flexible approach to this analysis plan, sometimes with revisions of
-previous steps. This is not optimal, and efforts were made to not deviate in any
-major way. Using GitHub, all code was made public for inspection and to ensure
-reproducibility [@RN59].
-
-## Clinical applications
-This study shows that mistriage can potentially increase following model
-transfer, and more prominently in certain cases. This supports the view that
-model transfers should be done with caution, as performance might
-deteriorate.
-
-Recognizing these risks and updating models accordingly should lead
-to increased accuracy during trauma patient triage. Potentially saving lives by
-minimizing undertriage, and likely save resources by minimizing overtriage. With
-registries such as SweTrau, analysis of model transfer could be performed
-continuously to optimize models currently in use and advise which model
-transfers might be detrimental (or potentially beneficial).
+The predetermined analysis plan was to be followed in a step-by-step fashion. 
+Programming challenges and late-time code inclusions required a flexible approach 
+to this analysis plan, sometimes with revisions of previous steps. This is not 
+optimal, and efforts were made to not deviate in any major way. Using GitHub, 
+all code was made public for inspection and to ensure reproducibility [@RN59].
 
@@
 Model transfer resulted in large and varying levels of uncertainty in its effect
-on mistriage. Therefore model transfer can be unpredictable, potentially leading 
+on mistriage. Therefore, model transfer can be unpredictable, potentially leading 
 to increased mistriage, but in some cases also to a decrease in mistriage. 
@@
 associated with considerable uncertainty. Recognizing the limitations of this study, 
-the authors believe further studies are warranted dueto the potential economic 
+the authors believe further studies are warranted due to the potential economic 
 costs and patient consequences of poor model transfers.

### 2019-07-31 MainCodeRun.R
Update analysis code

@@ -67,18 +67,21 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
                      group = list(full.label = "Group",
                                   abbreviated.label = ""))
     ## Get original results
-    if (clean.start | !original.results.done)
-        RunStudy(selected.data = selected.data, codebook = codebook, boot = FALSE, test = test)
+    if (clean.start | !original.results.done) {
+        message ("Estimating original results")
+        RunStudy(selected.data = selected.data, copy.results.to.path = copy.results.to.path, codebook = codebook, boot = FALSE, test = test)
+    }
     ## Create bootstrap samples
     number.of.bootstrap.samples = 1000
     bootstrap.samples <- list()
+    bootstrap.samples.exist <- file.exists("bootstrap.samples.Rds")
     if (test)
         number.of.bootstrap.samples = 5
-    if (!clean.start) 
+    if (!clean.start & bootstrap.samples.exist) 
         bootstrap.samples <- readRDS("bootstrap.samples.Rds")
-    if (!clean.start & length(bootstrap.samples) != number.of.bootstrap.samples)
+    if (!clean.start & length(bootstrap.samples) != number.of.bootstrap.samples & bootstrap.samples.exist)
         stop ("The number of saved bootstrap samples imported because clean.start = FALSE is not the same as the number of bootstrap samples requested.")
-    if (clean.start)
+    if (clean.start | !bootstrap.samples.exist)
         bootstrap.samples <- bengaltiger::CreateBootstrapSamples(selected.data,
                                                                  strata = "res_survival",
                                                                  save.to.disk = TRUE,
@@ -101,6 +104,6 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     rmarkdown::render("./ManuscriptMarkdown.Rmd")
 
 }
-MainCodeRun(test = TRUE, clean.start = TRUE)
+MainCodeRun(test = FALSE, clean.start = FALSE, copy.results.to.path = "~/ownCloud/projects/transfer-effect-mistriage-martin/")
 
 

### 2019-07-31 functions/RunStudy.R

@@ -13,7 +13,7 @@
 #'     MICEImplement and DevelopmentModelCreator. Defaults to FALSE.
 #' @param copy.results.to.path Character or NULL. The path to which the results
 #'     should be copied. Defaults to NULL.
-RunStudy <- function(selected.data, codebook = NULL boot = FALSE, test = FALSE, copy.results.to.path = NULL) {
+RunStudy <- function(selected.data, codebook = NULL, boot = FALSE, test = FALSE, copy.results.to.path = NULL) {
 
 
     ## Error handling
@@ -131,7 +131,7 @@ RunStudy <- function(selected.data, codebook = NULL boot = FALSE, test = FALSE, 
     file.name <- paste0("output/", results.specifier, ".results.Rds") 
     saveRDS(Results, file.name)
     if (!is.null(copy.results.to.path))
-        saveRDS(Results, paste0(path, "/", results.specifier, ".results.Rds"))
+        saveRDS(Results, paste0(copy.results.to.path, file.name))
     return(paste0(file.name, " saved to disk"))
 
 }

### 2019-08-01 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 - Martin Henriksson
-- Dell Saulnier
+- Dell D Saulnier
 - Johanna Berg
@@
 # Introduction
-Trauma, defined as an injury to living tissue by an extrinsic agent [@RN63], 
-results in approximatly 458,5 million cases anually across the globe [@RN999], 
-and an estimated 3,8 million deaths [@RN1; @RN21]. Each year 9% of the world’s 
-deaths are because of trauma, with the leading causes being road traffic accidents, 
-suicide and homicide. Predictions indicate the incidence of these causes are 
+Trauma, defined as physical injuries to a host by outside objects [@RN63], 
+results in approximately 458.5 million cases anually across the globe [@RN999], 
+and an estimated 3.8 million deaths [@RN1; @RN21]. Each year, 9% of the world’s 
+deaths are the result of trauma, with the leading causes being road traffic accidents, 
+suicide, and homicide. Predictions indicate the incidence of these causes is 
 likely to increase by 2030 [@RN42].
@@
 
-Systems that determine the level of trauma can be based on clinical prediction
-models. Models differ in quality and characteristics, but generally perform
-well at predicting survival [@RN45]. Many models are developed in a single,
-standardized setting, such as a major trauma centre but are then implemented in 
-different settings, such as non-trauma centres.
+Systems that determine the level of trauma during early trauma care can be based 
+on clinical prediction models. Models differ in quality and characteristics, but 
+generally perform well at predicting survival [@RN45]. Many models are developed 
+in a single, standardized context, such as a major trauma centre but are then 
+implemented in different context, such as non-trauma centres.
 
@@
 with major trauma as minor trauma. Mistriage can ultimatly lead to decreased patient 
-survival, but is also detrimental in terms of patient care and distribution of resources.
+survival, and is also detrimental to patient care and distribution of resources [@RN44].
 
 Thus, the effect of model transfers between care contexts within a single healthcare system, 
-as well as the effect of such transfers on mistriage,have not been studied and represent 
-substantial knowledge gaps. Therefore, the aim of this study is to assess how transfers 
-of clinical prediction models for early (on arrival to hospital) trauma 
-care between different care contexts within a single health system affect mistriage rates.
+as well as the effect of such transfers on mistriage, have not been studied and represent 
+substantial knowledge gaps. The aim of this study is to assess how transfers 
+of clinical prediction models for early trauma care between different care contexts 
+within a single health system affect mistriage rates.
 
@@
 
-The trauma care in Sweden is usually clearly defined in major
+Trauma care in Sweden is usually clearly defined in major
 metropolitan areas: trauma patients are transported directly to a predesignated
@@
 
-Today 52 of the 55 Swedish hospitals record trauma cases in SweTrau. 
+Today, 52 of the 55 Swedish hospitals record trauma cases in SweTrau. 
 Currently, the registry includes 55 000 cases, and its use is encouraged in 
@@
 ## Participants 
-The eligibility criteria were patients aged 15 years or above registered in SweTrau 
+The eligibility criteria were adult patients aged 15 years or above registered in SweTrau 
 between 2011 and 2016. The study aimed to assess adult trauma and not
@@
 Neither the outcome nor variables were blinded when during analysis, which 
-made a structured approach important, to ensure objectivity.
+made a structured approach important to ensure objectivity.
 
@@
 plausible care context for model transfer. These data sets were further divided 
-into samples, this process is outlined below.
+into samples:
 
@@
 undertriage and overtriage rate for the same model was `r val.undertriage`, and
-`r val.overtriage` respectively. Worst validation performance (i.e. highest
+`r val.overtriage` respectively. The worst validation performance (i.e. highest
 mistriage rate) was found in the `r colnames(highest.val.mistriage)` sample, with
@@
 
-Due to centralized trauma care, non-metropolitan centres generelly have fewer 
-resources, and less experience reciving trauma patients. They may rely 
-on clinical prediction models more heavily for accurate trauma triage, and an 
-increased mistriage leads to further strain on already limited resources
-due to increased overtriage. Increased undertriage could also possibly lead to
-increased patient mortality
+Due to centralized trauma care, non-metropolitan centres generally have fewer 
+resources and less experience receiving trauma patients. They may rely 
+on clinical prediction models more heavily for accurate trauma triage, and 
+an increase in overtriage can lead to further strains on already limited 
+resources. Increased undertriage could possibly lead to increased patient 
+mortality.
 
-Recognizing these risks and updating models accordingly should lead
-to increased accuracy during trauma patient triage. Potentially saving lives by
+Recognizing these risks and updating models accordingly could lead
+to increased accuracy during trauma patient triage, potentially saving lives by
 minimizing undertriage, and likely save resources by minimizing overtriage. With

### 2019-08-03 MainCodeRun.R
Fix issue of missing bootstrap

@@ -22,11 +22,10 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     }
     ## Identify starting point and go from there if clean start is FALSE
     original.results.done <- file.exists("./output/original.results.Rds")
-    start.from.bootstrap.sample <- 1
+    completed.bootstraps <- NULL
     if (!clean.start) {
         bootstraps.done <- list.files("./output/", "^bootstrap\\.sample\\.[0-9]*\\.results.Rds$")
-        bootstrap.index <- max(as.numeric(gsub("bootstrap\\.sample\\.|\\.results\\.Rds", "", bootstraps.done)))
-        start.from.bootstrap.sample <- bootstrap.index + 1
+        completed.bootstraps <- sort(as.numeric(gsub("bootstrap\\.sample\\.|\\.results\\.Rds", "", bootstraps.done)))
     }
     ## Set random seed
     set.seed(-41892)
@@ -72,7 +71,7 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
         RunStudy(selected.data = selected.data, copy.results.to.path = copy.results.to.path, codebook = codebook, boot = FALSE, test = test)
     }
     ## Create bootstrap samples
-    number.of.bootstrap.samples = 1000
+    number.of.bootstrap.samples <- 1000
     bootstrap.samples <- list()
     bootstrap.samples.exist <- file.exists("bootstrap.samples.Rds")
     if (test)
@@ -87,7 +86,9 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
                                                                  save.to.disk = TRUE,
                                                                  return.samples = TRUE,
                                                                  number.of.bootstrap.samples = number.of.bootstrap.samples)
-    bootstrap.samples <- bootstrap.samples[start.from.bootstrap.sample:number.of.bootstrap.samples]
+    if (!is.null(completed.bootstraps))
+        bootstrap.samples <- bootstrap.samples[-completed.bootstraps]
+    message (paste0("Estimating results in bootstrap samples ", paste0((1:number.of.bootstrap.samples)[-completed.bootstraps], collapse = ", ")))
     ## Get bootstrap results
     study.cluster <- makeCluster(detectCores())
     registerDoParallel(study.cluster)

### 2019-08-03 ManuscriptMarkdown.Rmd

@@
 - Martin Gerdin Wärnberg
-date: "July 31, 2019"
+date: "August 01, 2019"
 documentclass: "article"

### 2019-08-14 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

--- 2019-08-01 ManuscriptMarkdown.Rmd
@@
 - Martin Gerdin Wärnberg
-date: "July 31, 2019"
+date: "August 14, 2019"
 documentclass: "article"
@@
 
-  **Background:** Trauma results in approximately 3,8 million deaths annually. Clinical
+  **Background:** Trauma results in approximately 3.8 million deaths annually. Clinical
   prediction models are often used for triage in early trauma care. Previous research has
-  indicated transferring these models between different settings might adversely affect 
+  indicated that transferring these models between different settings might adversely affect 
   model performance.
 
-  **Methods:** Patients age 15 or above registered between the years 2011 and
-  2016 in the Swedish national trauma registry SweTrau were included. Three sets
-  of data were created: High and low volume centres, metropolitan and
-  non-metropolitan centres, and multi and single centres. Clinical
+  **Methods:** Patients aged 15 years or older who were registered between 2011 and
+  2016 in the Swedish national trauma registry, SweTrau, were included. Three sets
+  of data were created: high- and low-volume centres, metropolitan and
+  non-metropolitan centres, and multiple and single centres. Clinical
   prediction models were developed using logistic regression in each data set 
   and transferred within sets. Model performance was evaluated in all data
-  sets using mistriage rate, undertriage rate and overtriage rate. Multiple
-  imputation using chained equation was used to handle missing data. Model
-  performance was reported as medians with 95% uncertainty interval across
+  sets using the mistriage rate, undertriage rate and overtriage rate. Multiple
+  imputation using chained equations was used to handle missing data. Model
+  performance was reported as medians with the 95% uncertainty interval across
   imputed data sets.
 
-  **Results:** `r total.number.of.patients ` patients were included. Model
+  **Results:** A total of `r total.number.of.patients ` patients were included. The model
   validation mistriage rate ranged from `r lowest.val.mistriage` 
-  to `r highest.val.mistriage`, and transferred model
+  to `r highest.val.mistriage`, and the transferred model
   mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage` 
-  across all samples. Performance after transfer when compared to receiving sample validation 
-  performance resulted in subtantial uncertainty in all samples. Worst performance after 
+  across all samples. Performance after transfer when compared to the receiving sample validation 
+  performance resulted in substantial uncertainty in all samples. The worst performance after 
   transfer was found in the `r colnames(highest.comp.mistriage)` 
-  sample with a mistriage difference of `r highest.comp.mistriage`.
+  sample, with a mistriage difference of `r highest.comp.mistriage`.
 
-  **Conclusions:** In all samples model transfer was associated with high 
+  **Conclusions:** In all samples, model transfer was associated with high 
   levels of uncertainty. Notably, transfer from the `r colnames(highest.comp.mistriage)` 
-  sample to the other sample in the data set lead to an increased mistriage
+  sample to the other sample in the data set led to an increased mistriage
   rate, primarily due to increased overtriage.
@@
 Trauma, defined as physical injuries to a host by outside objects [@RN63], 
-results in approximately 458.5 million cases anually across the globe [@RN999], 
-and an estimated 3.8 million deaths [@RN1; @RN21]. Each year, 9% of the world’s 
+accounts for approximately 458.5 million cases annually across the globe [@RN999] 
+and an estimated 3.8 million deaths [@RN1; @RN21]. Each year, 9% of global 
 deaths are the result of trauma, with the leading causes being road traffic accidents, 
-suicide, and homicide. Predictions indicate the incidence of these causes is 
+suicide, and homicide. Predictions indicate that the incidence of these causes is 
 likely to increase by 2030 [@RN42].
 
-In a typical, high-resource setting, the initial management of trauma is
+In a typical high-resource setting, the initial management of trauma is
 performed on the scene by Emergency Medical Services
@@
 This information is then evaluated using a system to determine
-the level of trauma, prepare adequate resources [@RN43] and dictate if a full
+the level of trauma, prepare adequate resources [@RN43] and dictate whether a full
 or limited trauma team is activated [@RN44].
@@
 Systems that determine the level of trauma during early trauma care can be based 
-on clinical prediction models. Models differ in quality and characteristics, but 
+on clinical prediction models. Models differ in quality and characteristics but 
 generally perform well at predicting survival [@RN45]. Many models are developed 
-in a single, standardized context, such as a major trauma centre but are then 
-implemented in different context, such as non-trauma centres.
+in a single, standardized context, such as a major trauma centre, but are then 
+implemented in different contexts, such as non-trauma centres.
 
@@
 In trauma, misclassification is often referred to as mistriage. Triage refers to
-the classification of severity as minor or major trauma. Mistriage can be
-subdivided into overtriage, the incorrect classification of a patient with minor
-trauma as major trauma, or undertriage, the incorrect classification of a patient
-with major trauma as minor trauma. Mistriage can ultimatly lead to decreased patient 
-survival, and is also detrimental to patient care and distribution of resources [@RN44].
+the classification of trauma severity as minor or major. Mistriage can be
+subdivided into overtriage, which is the incorrect classification of a patient with minor
+trauma as one with major trauma, or undertriage, which is the incorrect classification of a patient
+with major trauma as one with minor trauma. Mistriage can ultimately lead to decreased patient 
+survival and is also detrimental to patient care and the distribution of resources [@RN44].
 
@@
 as well as the effect of such transfers on mistriage, have not been studied and represent 
-substantial knowledge gaps. The aim of this study is to assess how transfers 
+substantial knowledge gaps. The aim of this study was to assess how transfers 
 of clinical prediction models for early trauma care between different care contexts 
@@
 A registry-based cohort study was conducted. Sweden has a nationally encompassing
-trauma register called SweTrau. SweTrau data was used to create
-clinical prediction models, which were transferred between different care contexts to
+trauma registry called SweTrau. SweTrau data were used to create
+clinical prediction models, which were then transferred between different care contexts to
 study the effects on mistriage. The study and analysis plans were made
@@
 The societal effects of trauma are substantial, with the loss
-of work days comparable to those lost from cardiovascular and malignant diseases
+of days at work comparable to those lost from cardiovascular and malignant diseases
 combined [@RN23].
@@
 country, trauma patients are instead transported to the nearest hospital for
-stabilization. Once stabilized, they may then be transported to hospitals with
+stabilization. Once stabilized, they may then be transported to a hospital with
 more advanced trauma care capacity [@RN34]. Most hospitals in Sweden use some
-form of system to categorize the level of trauma, usually
-based on patient vital signs and injury mechanism, as reported by the EMS, or as
+system to categorize the level of trauma, usually
+based on patient vital signs and the mechanism of injury, as reported by the EMS or as
 registered in the emergency department [@RN90].
@@
 Today, 52 of the 55 Swedish hospitals record trauma cases in SweTrau. 
-Currently, the registry includes 55 000 cases, and its use is encouraged in 
+Currently, the registry includes 55 000 cases, and its use is encouraged for 
 both academic research and local quality improvement initiatives [@RN28].
@@
 ## Participants 
-The eligibility criteria were adult patients aged 15 years or above registered in SweTrau 
-between 2011 and 2016. The study aimed to assess adult trauma and not
-paediatric trauma which differs in physiology, triage and initial care
-[@RN46]. The Swedish guidelines for trauma activation define children as age <15
-[@RN90].
+The eligibility criteria were adult patients aged 15 years or above registered 
+in SweTrau between 2011 and 2016. The study aimed to assess adult trauma and not
+paediatric trauma, which differs in physiology, triage and initial care
+[@RN46]. The Swedish guidelines for trauma activation define children as 
+individuals aged <15 years [@RN90].
 
@@
 ### Participant characteristics
-To describe the patient cohort we reported age, sex, Injury Severity Score (ISS)
+To describe the patient cohort, we reported age, sex, Injury Severity Score (ISS)
 and New ISS (NISS).
@@
 ### Model predictors
-The clinical prediction models used the predictors systolic blood pressure
-(SBP), respiratory rate (RR) and Glasgow coma scale (GCS), all on arrival to
-hospital. The rationale for these three predictors is that they are
-part of multiple established clinical prediction models for early trauma care, such
-as the Revised Trauma Score [@RN35].
+The clinical prediction models used the following predictors on arrival
+at the hospital: systolic blood pressure (SBP), respiratory rate (RR) and 
+Glasgow coma scale (GCS). The rationale for using these three predictors 
+is that they are part of multiple established clinical prediction models 
+for early trauma care, such as the Revised Trauma Score [@RN35].
 
 ### Model outcome
-The outcome used to develop the clinical prediction models was all cause
+The outcome used to develop the clinical prediction models was all-cause
 mortality within 30 days of trauma.
@@
 ISS > 15 was used as the gold standard to define trauma severity as major
-trauma, and hence patients with ISS ≤ 15 were considered minor trauma
+trauma, and hence, patients with ISS ≤ 15 were considered to have minor trauma
 [@RN36]. Overtriage was defined as the event when a clinical prediction model
-classified a patient with ISS ≤ 15 as major trauma, and undertriage as the event
-when a clinical prediction model classified a patient with ISS > 15 as minor
-trauma. Overtriage rate was defined as the number of overtriaged patients
-divided by all patients. Undertriage rate was defined as the number of
-undertriaged patients divided by all patients. Mistriage rate was defined as the
+classified a patient with ISS ≤ 15 as having major trauma, and undertriage was defined as the event
+when a clinical prediction model classified a patient with ISS > 15 as having minor
+trauma. The overtriage rate was defined as the number of overtriaged patients
+divided by the total number of patients. The undertriage rate was defined as the number of
+undertriaged patients divided by the total number of patients. The mistriage rate was defined as the
 sum of the over- and undertriage rates.
@@
 Model predictors, outcome, participant characteristics and study outcome as
-outlined above were all obtained from SweTrau. The method of measurement for the
+outlined above were all obtained from SweTrau. The method of measurement of the
 model predictors is not specified in the registry entries. In Swedish emergency rooms,
@@
 after trauma is manually entered into the registry by each respective
-hospital. NISS and ISS are calculated by hospital personnel
+hospital. The NISS and ISS are calculated by hospital personnel
 based on the patient’s injuries.
@@
 Data analysis was conducted according to a prearranged analysis plan. The
-analysis plan and statistical analysis code was finalised using simulated
-data. These efforts were taken to avoid confirmation bias. The analysis plan was
+analysis plan and statistical analysis code were finalised using simulated
+data. These efforts were made to avoid confirmation bias. The analysis plan was
 reviewed by an experienced statistician and programmer prior to implementation. 
-Neither the outcome nor variables were blinded when during analysis, which 
+Neither the outcome nor variables were blinded during analysis, which 
 made a structured approach important to ensure objectivity.
@@
 ## Study size
-All patients matching eligibility criteria were included. Four data sets were
-used to study the transfer of clinical prediction models, each data set
-representing a different care context. The data sets, and sample size considerations 
+All patients matching the eligibility criteria were included. Four data sets were
+used to study the transfer of clinical prediction models, with each data set
+representing a different care context. The data sets and sample size considerations 
 are outlined below.
@@
 percentiles. When describing the participant characteristics, quantitative variables 
-were presented as continuous. ISS was presented as dichotomous using ISS > 15 as the cutoff.
+were presented as continuous. ISS was presented as dichotomous, using ISS > 15 as the cutoff.
 
@@
 plausible care context for model transfer. These data sets were further divided 
-into samples:
+into samples as follows:
 
-### Data set 1: High and low volume centres
-Two samples were derived from the SweTrau cohort, based on the number of patients. 
-High volume centres were those within the top quartile of number of patients received. 
-The rest were low volume centres.
+### Data set 1: High- and low-volume centres
+Two samples were derived from the SweTrau cohort based on the number of patients. 
+High-volume centres were those within the top quartile in terms of the number of 
+patients received. The rest were low-volume centres.
 
@@
 patients from greater Stockholm, greater Gothenburg and greater Malmö, as defined by 
-statistics Sweden, and a second sample of patients from non-metropolitan areas.
+Swedish statistics, and a second sample consisting of patients from non-metropolitan areas.
 
-### Data set 3: Multi- and single centre data
+### Data set 3: Multiple and single centre data
 In this data set, two samples were created. One centre with a large enough 
-sample size to develop and validate a model was its own sample. The multicentre 
+sample sizes to develop and validate a model was its own sample. The multiple centres 
 sample consisted of the combined data from all centres.
@@
 A fourth dataset was initially created, consisting of individual centres that 
-had a large enough sample size to develop and validate a model. However, this 
+had large enough sample size to develop and validate a model. However, this 
 data set was removed before analysis due to a lack of centres 
-with a large enough sample size.
+with a sufficiently large sample size.
 
@@
 two subsamples using a temporal split based on the date of trauma. The earlier
-subsample was the development sample, and the later subsample the validation
+subsample was the development sample, and the later subsample was the validation
 sample. The development sample contained 70 events (events being patients who
-died within 30 days of the trauma) and all non-events (non-events being patient
-survival 30 days past the trauma) during the same time. The rationale for
+died within 30 days of the trauma) and all non-events (non-events being patients who
+survived 30 days past the trauma) during the same time. The rationale for
 including 70 events is that at least 10 events per free parameter were needed in
@@
 validation samples [@RN37]. Figure 1 shows the development of the samples using 
-the high- and low volume centre data set as an example. The minimum sample size of development and
-validation samples were 140 and 200 respectively. We performed analyses only on
-data sets for which all samples included at least the minimum sample size.
+the high- and low-volume centre data set as an example. The minimum sample sizes 
+of the development and validation samples were 140 and 200, respectively. 
+We performed analyses only on data sets for which all samples included at 
+least the minimum sample size.
 
 ![](fig1.jpg)  
-**Fig. 1: Sample development in the high and low volume centre data set.**  Initial 
-split based on number of patients. Temporal split made using date of traumatic event.
+**Fig. 1: Sample development in the high- and low-volume centre data set.**  Initial 
+split based on the number of patients. Temporal split made using the date of the trauma.
 
@@
 were repeated in each data set. Below, we use the transfer of a model from the
-high volume centre sample to a low volume centre sample as an example to describe 
+high-volume centre sample to the low-volume centre sample as an example to describe 
 the complete procedure.
@@
 In the model development step, a clinical prediction model was developed in the
-high volume centre development sample. The model was developed using logistic
+high-volume centre development sample. The model was developed using logistic
 regression. The dependent variable was all-cause mortality within 30 days of 
-trauma and the independent variables SBP, RR, and GCS were modelled as previously described. 
+trauma, and the independent variables SBP, RR, and GCS were modelled as previously described. 
 To avoid overfitting the model, a bootstrap procedure was used to estimate a linear 
@@
 model was used to estimate the probability of all-cause 30-day mortality in the 
-development sample. A gridsearch was performed across estimated probabilities in 
-the development sample to identify the cut-off that optimised overtriage keeping 
-undertriage at less than 5% [@RN39]. This cut-off was then used to classify patients 
-as major or minor trauma.
+development sample. A grid search was performed across estimated probabilities in 
+the development sample to identify the cutoff value that optimised overtriage while keeping 
+undertriage at less than 5% [@RN39]. This cutoff value was then used to classify patients 
+as having major or minor trauma.
 
@@
 In the model validation step, the model performance was assessed in the
-high volume centre validation sample and in the low volume centre validation
+high-volume centre validation sample and in the low-volume centre validation
 sample. First, the model was used to estimate the probability of all-cause 30-day
-mortality in each sample. Then the probability cut-off identified in the
+mortality in each sample. Then, the probability cutoff value identified in the
 development sample was applied to the validation samples, patients were
-classified as major or minor trauma, and model performance was estimated.
+classified as having major or minor trauma, and the model performance was estimated.
 
@@
 Finally, the difference in model performance
-between the high and low volume centre validation samples was
-calculated. Empirical bootstrap was used to estimate 95% confidence intervals
-(CI) around performance and differences in performance estimates. Both bootstrap
+between the high- and low-volume centre validation samples was
+calculated. Empirical bootstrapping was used to estimate the 95% confidence intervals
+(CI) around the performance and the differences in the performance estimates. Both bootstrap
 procedures used 1 000 bootstrap samples drawn with replacement of the same size
@@
 We used multiple imputation using chained equations (mice) to handle missing data [@RN41]. 
-The number of imputations created
-for each data set was equal to the highest percentage of missing data in that
-data set. Quantitative variables were imputed using predictive mean matching and
-qualitative variables were imputed using logistic regression. SBP and RR was
-transformed as restricted cubic splines before imputation. All analyses outlined 
-above were then conducted separately in
+The number of imputations created for each data set was equal to the highest 
+percentage of missing data in that data set. Quantitative variables were imputed 
+using predictive mean matching, and qualitative variables were imputed using logistic 
+regression. SBP and RR were transformed as restricted cubic splines before imputation. 
+All analyses outlined above were then conducted separately in
 each imputed dataset. We present the combined results as the median point
@@
 75th percentiles across imputations, i.e. the lower bound of the presented
-interval is the lower bound of a 95% CI of the 25th percentile and the upper
+interval is the lower bound of a 95% CI of the 25th percentile, and the upper
 bound is the upper bound of a 95% CI of the 75th percentile. This combined CI
-was referred to as an Uncertainty Interval (UI) and was used to express the
+was referred to as an uncertainty interval (UI) and was used to express the
 added uncertainty associated with the imputation procedure to handle missing
@@
 The study was approved by the regional ethics review board in Stockholm, Sweden. 
-Ethical review numbers are 2015/426-31 and 2016/461-32. All patients in SweTrau 
-had received letters detailing their inclusion in the database, and the possibility 
+The ethics review numbers are 2015/426-31 and 2016/461-32. All patients in SweTrau 
+received letters detailing their inclusion in the database and the possibility 
 of their data being used for scientific purposes. It was assumed that the patients 
-who consented to the use of their SweTrau data for research want to contribute to 
-the improvement of care, and thus by performing this study the authors hope to 
+who consented to the use of their SweTrau data for research wanted to contribute to 
+the improvement of care, and thus, by performing this study, the authors hope to 
 honour their wishes.

### 2019-08-14 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 percentage of missing observations was the `r highest.percent.missing.name `
-sample with `r highest.percent.missing.number `% incomplete observations. The
-variable with the highest number of missing values was `r name.variable.na `,
+sample, with `r highest.percent.missing.number `% incomplete observations. The
+variable with the highest number of missing values was the `r name.variable.na `,
 with `r number.variable.na ` missing values, or `r round(percent.variable.na)`% of 
 the total values for this variable. The percentages of missing values for the other 
-model predictors, was `r round(gcs.na$percentage)`% for GCS and `r round(sbp.na$percentage)`% 
+model predictors were `r round(gcs.na$percentage)`% for GCS and `r round(sbp.na$percentage)`% 
 for SBP.
 
-**Table 1. Sample characteristics of multiple imputed data.** Data is presented as medians with interquartile range (IQR) or counts with % as applicable.
+**Table 1. Sample characteristics of multiple imputed data.** Data are presented as medians with interquartile range (IQR) or counts with %, as applicable.
 ```{r, echo = FALSE, results = "asis"} 
@@
 ``` 
-GCS, glasgow coma scale; SBP, systolic blood pressure; RR, respiratory rate; ISS, injury severity score; NISS, new injury severity score
+GCS, Glasgow coma scale; SBP, systolic blood pressure; RR, respiratory rate; ISS, injury severity score; NISS, new injury severity score
 
@@
 During model development, the number of imputations used for each data set was
-`r high.low.imp`, `r metro.non.imp` and `r multi.single.imp` for the high and
-low volume data set, the metropolitan and non-metropolitan data set, and the
-multi- and single centre data set, respectively.
+`r high.low.imp`, `r metro.non.imp` and `r multi.single.imp` for the high- and
+low-volume data set, the metropolitan and non-metropolitan data set, and the
+multiple and single centre data set, respectively.
 
-Table 2 shows model validation performance. The model with the lowest validation
-mistriage rate was the `r colnames(lowest.val.mistriage)` model with a median
-mistriage rate of `r lowest.val.mistriage`. Performance in terms of
-undertriage and overtriage rate for the same model was `r val.undertriage`, and
-`r val.overtriage` respectively. The worst validation performance (i.e. highest
+Table 2 shows the model validation performance. The model with the lowest validation
+mistriage rate was the `r colnames(lowest.val.mistriage)` model, with a median
+mistriage rate of `r lowest.val.mistriage`. The performance in terms of
+undertriage and overtriage rate for the same model was `r val.undertriage` and
+`r val.overtriage`, respectively. The worst validation performance (i.e. highest
 mistriage rate) was found in the `r colnames(highest.val.mistriage)` sample, with
@@
 
-**Table 2. Model validation performance, i.e. performance when applied to the sample in which the model was created.** Data is presented as point estimates with 95% Uncertainty Intervals (95% UI).
+**Table 2. Model validation performance, i.e. performance when applied to the sample in which the model was created.** Data are presented as point estimates with 95% uncertainty intervals (95% UI).
 ```{r, echo = FALSE, results = "asis"} 
@@
 ## Comparison
-Model performance after transfer was determined for each model being transferred
-to the other sample in the data set, in all data sets (Table 3). When
+Model performance after transfer was determined for each model after being transferred
+to the other sample in the data set in all data sets (Table 3). When
 transferred, the model with the lowest mistriage rate was the `r colnames(lowest.tran.mistriage)` 
-model with a median mistriage rate of `r lowest.tran.mistriage`. The model with 
-the highest mistriage rate after transfer was the `r colnames(highest.tran.mistriage)` 
+model, with a median mistriage rate of `r lowest.tran.mistriage`. The model with 
+the highest mistriage rate after transfer was the `r colnames(highest.tran.mistriage)` model 
 with a median mistriage rate of `r highest.tran.mistriage`.
 
-**Table 3. Model transfer performance, i.e. performance when transferred to the other sample in the same data set. ** Data is presented as point estimates with 95% Uncertainty Intervals (95% UI).  
+**Table 3. Model transfer performance, i.e. performance when transferred to the other sample in the same data set.** Data are presented as point estimates with 95% uncertainty intervals (95% UI).  
 ```{r, echo = FALSE, results = "asis"} 
@@
 Table 4 shows the transferred model performance when compared to the validation
-performance in the sample reciving the transferred model.
-The highest mistriage rate when compared to the reciveing sample validation
+performance in the sample receiving the transferred model.
+The highest mistriage rate when compared to the receiving sample validation
 model performance was found in the `r colnames(highest.comp.mistriage)` model,
@@
 patients, `r abs(worst.transfer$Transferred.mistriage.minus.buddy.local.mistriage.median) * 100` 
-more patients would be wrongly classified as major or minor trauma.
+more patients would be wrongly classified as having major or minor trauma.
 
-**Table 4. Comparison mistriage, i.e. sample validation performance subtracted from transferred performance.** Data is presented as point estimates with 95% Uncertainty Intervals (95% UI).  
+**Table 4. Comparison mistriage, i.e. sample validation performance subtracted from transferred performance.** Data are presented as point estimates with 95% uncertainty intervals (95% UI).  
 ```{r, echo = FALSE, results = "asis"} 
@@
 mistriage rates. In all data sets, model transfer between samples resulted in
-substantial uncertainty in its effect on model performance. When comparing
+substantial uncertainty in terms of its effect on model performance. When comparing the
 transferred performance to the receiving sample validation performance, the
@@
 
-The most notable effects on model performance following model transfer were seen
+The most notable effects on model performance following model transfer were observed
 after transferring the `r colnames(highest.comp.mistriage)` model. This transfer
-resulted in an increase of the mistriage rate by 
+resulted in an increase in the mistriage rate of 
 `r substr(format(highest.comp.mistriage), 1, 5)`.  Mainly contributing to this was
 a marked increase in overtriage.  For this transfer, the 95% UI ranged from 
-`r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating a high level of uncertainty and
+`r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating a high level of uncertainty, suggesting
 that our findings are compatible with a `r UI.top.bot[1]` decrease to a 
@@
 would signify an improvement in model performance. The higher 95% UI of 
-`r UI.top.bot[2]` would mean a marked increase in model mistriage, compared to the
+`r UI.top.bot[2]` would mean a marked increase in model mistriage compared to the
 point estimate of `r substr(format(highest.comp.mistriage), 1, 5)`.
@@
 to increased accuracy during trauma patient triage, potentially saving lives by
-minimizing undertriage, and likely save resources by minimizing overtriage. With
-registries such as SweTrau, analysis of model transfer could be performed
-continuously to optimize models currently in use and advise which model
+minimizing undertriage, and likely saving resources by minimizing overtriage. With
+registries such as SweTrau, the analysis of model transfer could be performed
+continuously to optimize the models currently in use and advise which model
 transfers might be detrimental (or potentially beneficial).
@@
 The transfer of clinical prediction models in trauma care has not previously
-been studied extensively. In 2016, Gerdin et al. found model transfer to
-adversely affect model performance in terms of calibration, but that this could
+been studied extensively. In 2016, Gerdin et al. found that model transfer
+adversely affected model performance in terms of calibration but that this could
 be improved by updating the model [@RN27]. This study did not examine model
-performance in terms of calibration, neither were the effects of updating the
+performance in terms of calibration, nor were the effects of updating the
 model explored.
@@
 extensively. Studies using simulated clinical prediction models based on
-different predictors showed a decline in model performance when being externally
+different predictors showed a decline in model performance when externally
 validated [@RN60; @RN61]. A study from 2018 examining the external validity of
-prediction models for coronary artery disease (CAD) [@RN62], showed models
-significantly underestimating the probability of CAD. The combined results of
+prediction models for coronary artery disease (CAD) [@RN62] showed that the models
+significantly underestimated the probability of CAD. The combined results of
 this study, the studies on model transfer and the studies on external validation
@@
 This study used GCS, RR and SBP as predictors for 30-day
-mortality. Additional predictors such as age, mechanism
-of injury, or newer predictors like shock index might increase model
-performance. Furthermore, we used 30-day mortality as outcome, recognizing
+mortality. Additional predictors such as age and mechanism
+of injury or newer predictors such as the shock index might increase model
+performance. Furthermore, we used 30-day mortality as an outcome, recognizing
 that late mortality, time in hospital, functional impairment, and morbidity are
-also important outcome factors in the context of trauma care.
+also important outcomes in the context of trauma care.
 
@@
 ## Conclusion
-Model transfer resulted in large and varying levels of uncertainty in its effect
+Model transfer resulted in large and varying levels of uncertainty in in terms of its effect
 on mistriage. Therefore, model transfer can be unpredictable, potentially leading 
-to increased mistriage, but in some cases also to a decrease in mistriage. 
-Noticeably, transfer of the `r colnames(highest.comp.mistriage)` model lead to an 
+to increased mistriage but in some cases also to a decrease in mistriage. 
+Noticeably, transfer of the `r colnames(highest.comp.mistriage)` model led to an 
 increased mistriage of `r highest.comp.mistriage`. This observed effect was also 
 associated with considerable uncertainty. Recognizing the limitations of this study, 
-the authors believe further studies are warranted due to the potential economic 
+the authors believe that further studies are warranted due to the potential economic 
 costs and patient consequences of poor model transfers.

### 2019-08-15 functions/DevValCreator.R
Update DevValCreator.R

@@ -13,8 +13,12 @@ DevValCreator <- function(df) {
         stop ("Input has to be a data.frame")
     ## Selecting events only
     events.only <- df[which(df$res_survival == "Dead"),]
-    ## Sorting by date
-    sorted.events.only <- events.only[order(as.Date(events.only$DateTime_Of_Trauma), decreasing = FALSE),]
+    
+    ## Sorting by date (new line 190815)
+    sorted.events.only <- events.only[order(as.Date(as.character(events.only$DateTime_Of_Trauma, "%Y%m%d")), decreasing = FALSE),]
+    ## Sorting by date (old line)
+    # sorted.events.only <- events.only[order(as.Date(events.only$DateTime_Of_Trauma), decreasing = FALSE),]
+    
     ## Saving cut-off date
     cut.off.date <- sorted.events.only$DateTime_Of_Trauma[70]
     ##Split data frame

### 2019-08-15 MainCodeRun.R
Update MainCodeRun.R

@@ -46,18 +46,18 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     ## Mark entries as valid (>170 events) individual centres
     selected.data <- IndividualCentreCheck(selected.data)
     ## Define codebook
-    codebook <- list(pt_age_yrs = list(full.label = "Patient age",
-                                       abbreviated.label = "Age"),
-                     pt_Gender = list(full.label = "Patient gender",
-                                      abbreviated.label = "Gender"),
+codebook <- list(pt_age_yrs = list(full.label = "Patient age, years",
+                                       abbreviated.label = "Age, years"),
+                     pt_Gender = list(full.label = "Patient sex",
+                                      abbreviated.label = "Sex"),
                      ed_gcs_sum = list(full.label = "Glasgow coma scale",
                                        abbreviated.label = "GCS"),
                      ed_sbp_value = list(full.label = "Systolic blood pressure",
                                          abbreviated.label = "SBP"),
                      ed_rr_value = list(full.label = "Respiratory rate",
                                         abbreviated.label = "RR"),
-                     res_survival = list(full.label = "30 day survival",
-                                         abbreviated.label = "30d survival"),
+                     res_survival = list(full.label = "30-day survival",
+                                         abbreviated.label = "30-day survival"),
                      ISS = list(full.label = "Injury severity score",
                                 abbreviated.label = "ISS"),
                      NISS = list(full.label = "New injury severity score",

### 2019-08-15 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
   high.low <- results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume
-  names(high.low) <- c("characteristic", "level", "high volume", "low volume")
+  names(high.low) <- c("Characteristic", "Level", "High volume", "Low volume")
   met.nonmet <- results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan
-  names(met.nonmet) <- c("characteristic", "level", "metropolitan", "non metropolitan")
+  names(met.nonmet) <- c("Characteristic", "Level", "Metropolitan", "Non-metropolitan")
   multi.single <- results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre
-  names(multi.single) <- c("characteristic", "level", "multi centre", "single centre")
+  names(multi.single) <- c("Characteristic", "Level", "Multiple centres", "Single centre")
    # Create complete table one (test)

### 2019-08-15 functions/WorkTableCreator.R
Update WorkTableCreator.R

@@ -11,7 +11,7 @@ WorkTableCreator <- function(result.tables.data.frame) {
   results.all.data$multi.centre.Measure <- NULL
   results.all.data$single.centre.11001.Measure <- NULL
   ## Set column names
-  colnames(results.all.data) <- c("Measure", "high volume", "low volume", "metropolitan", "non metropolitan", "multi centre", "single centre")
+  colnames(results.all.data) <- c("Measure", "High volume", "Low volume", "Metropolitan", "Non-metropolitan", "Multiple centres", "Single centre")
   ## Retrive names to use for rows
   measure.names <- results.all.data$Measure
   measure.names <- as.vector(measure.names)

### 2019-08-16 MainCodeRun.R
Edit code to catch error in charToDate

@@ -30,7 +30,10 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     ## Set random seed
     set.seed(-41892)
     ## Import data
-    raw.data <- ImportStudyData("swetrau-20110101-20160425.csv")
+    data.file <- "swetrau-20110101-20160425.csv"
+    if (test)
+        data.file <- "simulated-swetrau-data.csv"
+    raw.data <- ImportStudyData(data.file)
     ## Create study sample from selected variables
     selected.data <- VariableSelection(raw.data)
     ## Select cases with age > 15 or age = NA
@@ -86,16 +89,29 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
                                                                  save.to.disk = TRUE,
                                                                  return.samples = TRUE,
                                                                  number.of.bootstrap.samples = number.of.bootstrap.samples)
-    if (!is.null(completed.bootstraps))
+    estimated.bootstraps <- 1:number.of.bootstrap.samples
+    if (!is.null(completed.bootstraps)) {
         bootstrap.samples <- bootstrap.samples[-completed.bootstraps]
-    message (paste0("Estimating results in bootstrap samples ", paste0((1:number.of.bootstrap.samples)[-completed.bootstraps], collapse = ", ")))
+        estimated.bootstraps <- estimated.bootstraps[-completed.bootstraps]
+    }
+    message (paste0("Estimating results in bootstrap samples ", paste0(estimated.bootstraps, collapse = ", ")))
     ## Get bootstrap results
+    errorlog <- "output/errorlog.txt"
+    write("", errorlog)
     study.cluster <- makeCluster(detectCores())
     registerDoParallel(study.cluster)
     foreach(bootstrap.sample = bootstrap.samples,
             .packages = FuncPack(return.only = TRUE)$packages,
-            .export = FuncPack(return.only = TRUE)$functions) %dopar%
-        RunStudy(selected.data = bootstrap.sample, boot = TRUE, test = test, copy.results.to.path = copy.results.to.path)
+            .export = FuncPack(return.only = TRUE)$functions) %dopar% {
+                boot.id <- mean(bootstrap.sample[, ".boot.id"])
+                tryCatch(expr = RunStudy(selected.data = bootstrap.sample,
+                                         boot = TRUE,
+                                         test = test,
+                                         copy.results.to.path = copy.results.to.path),
[…]

### 2019-08-16 functions/MICEImplement.R

@@ -25,8 +25,9 @@ MICEImplement <- function(df.list, test = FALSE) {
     
     ## Set number of imputations
     number.of.imputations <- as.integer(max(unlist(percents.NA)))
+    maxit <- 5
     if (test)
-        number.of.imputations <- 3
+        number.of.imputations <- maxit <- 3
     ## number.of.imputations <- 5
     
     ## Run mice

### 2019-08-17 functions/DevValCreator.R
Add tryCatch

@@ -14,8 +14,14 @@ DevValCreator <- function(df) {
     ## Selecting events only
     events.only <- df[which(df$res_survival == "Dead"),]
     
-    ## Sorting by date (new line 190815)
-    sorted.events.only <- events.only[order(as.Date(as.character(events.only$DateTime_Of_Trauma, "%Y%m%d")), decreasing = FALSE),]
+    ## Sorting by date (new segment 20190817, tryCatch added)
+    sorted.events.only <- tryCatch(expr = events.only[order(as.Date(as.character(events.only$DateTime_Of_Trauma, "%Y%m%d")), decreasing = FALSE),],
+                                   error = function(e) {
+                                       .imp <- unique(df$.imp)
+                                       filename <- paste0("output/failed.imputation.", .imp, ".", gsub(":|-| ", "", format(Sys.time())), ".Rds")
+                                       saveRDS(df, filename)
+                                       stop(paste0("Error on .imp ", .imp, ", saved as ", filename, ": ", e))
+                                   })
     ## Sorting by date (old line)
     # sorted.events.only <- events.only[order(as.Date(events.only$DateTime_Of_Trauma), decreasing = FALSE),]
     

### 2019-08-18 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
   results.all.data <- WorkTableCreator(result.tables.data.frame)
+  ## Set names in in retrival table for correct formatting
+  names(results.all.data) <- c("high-volume", "low-volume", "metropolitan", "non-metropolitan", "multiple centres", "single centre")
 
@@
   na.vector <- unlist(results$NA.info.sample)
-  names(na.vector) <- c("high volume", "high volume", "low volume", "low volume", "metropolitan", "metropolitan", "non metropolitan", "non metropolitan", "multi centre", "multi centre", "single centre", "single centre")
+  names(na.vector) <- c("high-volume", "high-volume", "low-volume", "low-volume", "metropolitan", "metropolitan", "non-metropolitan", "non-metropolitan", "multiple centres", "multiple centres", "single centre", "single centre")
   highest.percent.missing.number <- max(na.vector[na.vector<100])
@@
 Model performance after transfer was determined for each model after being transferred
-to the other sample in the data set in all data sets (Table 3). When
+to the other sample in the data set, in all data sets (Table 3). When
 transferred, the model with the lowest mistriage rate was the `r colnames(lowest.tran.mistriage)` 

### 2019-08-22 functions/MICEImplement.R
Change maximum number of iterations to 3 when test is TRUE

@@ -31,8 +31,8 @@ MICEImplement <- function(df.list, test = FALSE) {
     ## number.of.imputations <- 5
     
     ## Run mice
-    df.list <- lapply(df.list, function(df) { 
-        temporary.data <- mice(df, m = number.of.imputations, maxit = 5) 
+    df.list <- lapply(df.list, function(df) {
+        temporary.data <- mice(df, m = number.of.imputations, maxit = maxit) 
         df <- complete(temporary.data, action = "long", include = TRUE)
         return(df)
     })

### 2019-08-22 functions/FuncPack.R
Change bengaltiger commit reference for installation

@@ -6,7 +6,7 @@
 FuncPack <- function(return.only = FALSE) {
     ## Get specific bengaltiger version
     library(devtools)
-    install_github("martingerdin/bengaltiger@c360801126f95d1573ed5e766d743648b51cd8f9")
+    install_github("martingerdin/bengaltiger@81672faabfb2d18dd66a92a7c0c4f32b2477d2f4")
     
     ## Load required packages
     packages <- c("data.table",

### 2019-08-22 functions/RunStudy.R
Disable use of restricted cubic splines

@@ -64,8 +64,8 @@ RunStudy <- function(selected.data, codebook = NULL, boot = FALSE, test = FALSE,
         Results$NA.info.variable <- NA.info.variable
         rm(NA.info.sample, NA.info.variable)
     }
-    ## Create restricted cubic splines
-    data.sets <- lapply(data.sets, function(sample) lapply(sample, RCSplineConvert))
+    ## Create restricted cubic splines (disabled 20190822)
+    ## data.sets <- lapply(data.sets, function(sample) lapply(sample, RCSplineConvert))
     ## Impute missing data
     data.sets <- lapply(data.sets, MICEImplement, test = test)
     ## If not a bootstrap run, create table one
@@ -90,10 +90,10 @@ RunStudy <- function(selected.data, codebook = NULL, boot = FALSE, test = FALSE,
     
     ## CLINICAL PREDICTION MODEL
     ## MODEL DEVELOPMENT
-    ## Remove all restricted cubic splines
-    split.data.sets <- lapply(split.data.sets, function(sample) lapply(sample, function(imp) (lapply(imp, function(devval) lapply(devval, RCRemover)))))
-    ## Recreate restricted cubic splines using knot locations from development sample in validation sample
-    split.data.sets <- lapply(split.data.sets, function(sample) lapply(sample, function(imp) (lapply(imp, RCSplineConvertDevVal)))) 
+    ## Remove all restricted cubic splines (disabled 20190822)
+    ## split.data.sets <- lapply(split.data.sets, function(sample) lapply(sample, function(imp) (lapply(imp, function(devval) lapply(devval, RCRemover)))))
+    ## Recreate restricted cubic splines using knot locations from development sample in validation sample (disabled 20190822)
+    ## split.data.sets <- lapply(split.data.sets, function(sample) lapply(sample, function(imp) (lapply(imp, RCSplineConvertDevVal)))) 
     ## Create model and apply shrinkage factor for each development sample, save as Development model coefficients
     split.data.sets <- lapply(split.data.sets, function(sample) lapply(sample, function(imp) (lapply(imp, DevelopmentModelCreator, test = test)))) 
     ## Predict 30-day mortality in development sample and create development grid + prediction grid

### 2019-08-22 functions/DataCleaning.R
Set DateTime_Of_Trauma = "" to missing

@@ -15,9 +15,10 @@ DataCleaning <- function(df) {
     if (!is.data.frame(df))
         stop ("Input has to be a data.frame")
     ## Entry replacements
-    df$res_survival <- MyReplace(df$res_survival, 999)
+    df$res_survival <- MyReplace(df$res_survival, 999, NA)
     df$ed_gcs_sum <- MyReplace(df$ed_gcs_sum, 999)
     df$ed_gcs_sum <- MyReplace(df$ed_gcs_sum, 99, 3)
+    df$DateTime_Of_Trauma <- MyReplace(df$DateTime_Of_Trauma, "", NA)
     ## Convert patient sex to factor
     df$pt_Gender <- factor(df$pt_Gender, levels = c(1, 2), labels = c("Male", "Female"))
     ## Convert patient 30-day survival to factor

### 2019-08-22 functions/DataCleaning.R
Modify function description

@@ -5,6 +5,7 @@
 #' res_survival 999 is replaced with NA.
 #' GCS 999 is raplaced with NA.
 #' GCS 99 is replaced with 3.
+#' DateTime_Of_Trauma "" is replaced with NA.
 #' Patient sex converted to factor w/ 2 levels
 #' Patient 30-day survival converted to factor w/ 2 levels
 #' New colum is created: ISS_over_15, dichotomized and converted to factor

### 2019-08-22 functions/DataCleaning.R
Implement categorised vital signs using revised trauma score categories instead of restricted cubic splines

@@ -29,5 +29,8 @@ DataCleaning <- function(df) {
     df$ISS_over_15 <- df$ISS
     df$ISS_over_15 <- ifelse(df$ISS_over_15 > 15, 1, 2)
     df$ISS_over_15 <- factor(df$ISS_over_15, levels = c(1, 2), labels = c("Yes", "No"))
+    ## Categorise systolic blood pressure, respiratory rate and Glasgow come
+    ## scale using Revised Trauma Score cutoffs
+    df[, c("gcs", "sbp", "rr")] <- bengaltiger::GetRevisedTraumaScoreComponents(df, gcs.name = "ed_gcs_sum", sbp.name = "ed_sbp_value", rr.name = "ed_rr_value", return.labels = TRUE)
     return(df)
 }

### 2019-08-22 MainCodeRun.R
Modify codebook to reflect change of vital signs variable names

@@ -48,18 +48,18 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     ## Mark entries as valid (>170 events) individual centres
     selected.data <- IndividualCentreCheck(selected.data)
     ## Define codebook
-codebook <- list(pt_age_yrs = list(full.label = "Patient age, years",
-                                       abbreviated.label = "Age, years"),
-                     pt_Gender = list(full.label = "Patient sex",
-                                      abbreviated.label = "Sex"),
-                     ed_gcs_sum = list(full.label = "Glasgow coma scale",
+    codebook <- list(pt_age_yrs = list(full.label = "Patient age",
+                                       abbreviated.label = "Age"),
+                     pt_Gender = list(full.label = "Patient gender",
+                                      abbreviated.label = "Gender"),
+                     gcs = list(full.label = "Glasgow coma scale",
                                        abbreviated.label = "GCS"),
-                     ed_sbp_value = list(full.label = "Systolic blood pressure",
+                     sbp = list(full.label = "Systolic blood pressure",
                                          abbreviated.label = "SBP"),
-                     ed_rr_value = list(full.label = "Respiratory rate",
+                     rr = list(full.label = "Respiratory rate",
                                         abbreviated.label = "RR"),
-                     res_survival = list(full.label = "30-day survival",
-                                         abbreviated.label = "30-day survival"),
+                     res_survival = list(full.label = "30 day survival",
+                                         abbreviated.label = "30d survival"),
                      ISS = list(full.label = "Injury severity score",
                                 abbreviated.label = "ISS"),
                      NISS = list(full.label = "New injury severity score",

### 2019-08-22 functions/IndividualCentreCheck.R
Update number of events needed for a centre to be a valid single centre

@@ -1,7 +1,7 @@
 #' IndividualCentreCheck
 #'
 #' Creates a table from a data.frame, based on number of deaths within 30 days of trauma.
-#' "Sjukhusnummer" with > 170 deaths within 30 days of trauma are considered valid individual centres.
+#' "Sjukhusnummer" with > 220 deaths within 30 days of trauma are considered valid individual centres.
 #' Adds a new colum called Valid_Individual_Centre
 #' If a "Sjukhusnummer" is a valid individual centre, adds Yes to the new colum, otherwise No.
 #' 
@@ -16,7 +16,7 @@ IndividualCentreCheck <- function(df) {
   
     ## Evaluates if Sjukhuskod is valid individual centre
     x <- as.data.frame(table(df$Sjukhuskod, df$res_survival, exclude = c(2, NA, NaN, 999, "Alive")))
-    y <- as.vector(x$Var1[x$Freq >= 170])
+    y <- as.vector(x$Var1[x$Freq >= 220])
   
     df$Valid_Individual_Centre <- ifelse(df$Sjukhuskod %in% y, "Yes", "No")
   

### 2019-08-22 functions/DevValCreator.R
Update number of events needed in development sample to 120

@@ -26,7 +26,7 @@ DevValCreator <- function(df) {
     # sorted.events.only <- events.only[order(as.Date(events.only$DateTime_Of_Trauma), decreasing = FALSE),]
     
     ## Saving cut-off date
-    cut.off.date <- sorted.events.only$DateTime_Of_Trauma[70]
+    cut.off.date <- sorted.events.only$DateTime_Of_Trauma[120]
     ##Split data frame
     dev.sample <- df[df$DateTime_Of_Trauma <= cut.off.date, ]
     val.sample <- df[df$DateTime_Of_Trauma > cut.off.date, ]

### 2019-08-23 functions/DataCleaning.R
Move transformation of DateTime_Of_Trauma to DataCleaning and clean up DevValCreator

@@ -20,6 +20,7 @@ DataCleaning <- function(df) {
     df$ed_gcs_sum <- MyReplace(df$ed_gcs_sum, 999)
     df$ed_gcs_sum <- MyReplace(df$ed_gcs_sum, 99, 3)
     df$DateTime_Of_Trauma <- MyReplace(df$DateTime_Of_Trauma, "", NA)
+    df$DateTime_Of_Trauma <- as.Date(as.character(df$DateTime_Of_Trauma, "%Y%m%d"))
     ## Convert patient sex to factor
     df$pt_Gender <- factor(df$pt_Gender, levels = c(1, 2), labels = c("Male", "Female"))
     ## Convert patient 30-day survival to factor

### 2019-08-23 functions/DevValCreator.R

@@ -14,17 +14,15 @@ DevValCreator <- function(df) {
     ## Selecting events only
     events.only <- df[which(df$res_survival == "Dead"),]
     
-    ## Sorting by date (new segment 20190817, tryCatch added)
-    sorted.events.only <- tryCatch(expr = events.only[order(as.Date(as.character(events.only$DateTime_Of_Trauma, "%Y%m%d")), decreasing = FALSE),],
+    ## Sorting by date 
+    sorted.events.only <- tryCatch(expr = events.only[order(events.only$DateTime_Of_Trauma), decreasing = FALSE),],
                                    error = function(e) {
                                        .imp <- unique(df$.imp)
                                        filename <- paste0("output/failed.imputation.", .imp, ".", gsub(":|-| ", "", format(Sys.time())), ".Rds")
                                        saveRDS(df, filename)
                                        stop(paste0("Error on .imp ", .imp, ", saved as ", filename, ": ", e))
                                    })
-    ## Sorting by date (old line)
-    # sorted.events.only <- events.only[order(as.Date(events.only$DateTime_Of_Trauma), decreasing = FALSE),]
-    
+   
     ## Saving cut-off date
     cut.off.date <- sorted.events.only$DateTime_Of_Trauma[120]
     ##Split data frame

### 2019-08-23 functions/DataCleaning.R
Remove original vital signs from data

@@ -33,5 +33,6 @@ DataCleaning <- function(df) {
     ## Categorise systolic blood pressure, respiratory rate and Glasgow come
     ## scale using Revised Trauma Score cutoffs
     df[, c("gcs", "sbp", "rr")] <- bengaltiger::GetRevisedTraumaScoreComponents(df, gcs.name = "ed_gcs_sum", sbp.name = "ed_sbp_value", rr.name = "ed_rr_value", return.labels = TRUE)
+    df[, c("ed_gcs_sum", "ed_sbp_value", "ed_rr_value")] <- NULL
     return(df)
 }

### 2019-08-23 functions/MICEImplement.R
Clean empty lines in imputation code

@@ -28,8 +28,7 @@ MICEImplement <- function(df.list, test = FALSE) {
     maxit <- 5
     if (test)
         number.of.imputations <- maxit <- 3
-    ## number.of.imputations <- 5
-    
+
     ## Run mice
     df.list <- lapply(df.list, function(df) {
         temporary.data <- mice(df, m = number.of.imputations, maxit = maxit) 

### 2019-08-23 ManuscriptMarkdown.Rmd
Exclude patients with missing date and time of trauma and report this number in the manuscript

--- 2019-08-15 ManuscriptMarkdown.Rmd
@@
   total.number.of.na <- results$NA.info.sample$multi.centre.vs.single.centre$multi.centre$total.number.of.na
+  n.missing.data.time.of.trauma <- Results$n.missing.date.time.of.trauma
   ## Sample NA
@@
 [@RN46]. The Swedish guidelines for trauma activation define children as
-individuals aged <15 years [@RN90].
+individuals aged <15 years [@RN90]. Further, patients with missing date and time
+of trauma were excluded.
 
@@
 We analysed data from `r total.number.of.patients ` trauma patients to
-investigate the effects of transfer on clinical prediction models (Table 1). The
-total number of missing observations across all variables was `r total.number.of.na ` 
+investigate the effects of transfer on clinical prediction models (Table 1), after 
+excluding `r n.missing.date.time.of.trauma ` patients with missing date and time of 
+trauma. The total number of missing observations across all variables was `r total.number.of.na ` 
 in the entire study cohort. The sample with the highest

### 2019-08-23 functions/RunStudy.R

@@ -28,6 +28,11 @@ RunStudy <- function(selected.data, codebook = NULL, boot = FALSE, test = FALSE,
     Results <- list()
 
     ## RUN STUDY
+    ## Exclude observations with missing date and time of trauma
+    if (!boot)
+        Results$n.missing.date.time.of.trauma <- sum(is.na(selected.data$DateTime_Of_Trauma))
+    selected.data <- selected.data[!is.na(selected.data$DateTime_Of_Trauma), ]
+    
     ## Create High Volume Sample and Low Volume Sample
     High.Volume.Sample <- selected.data[selected.data$High_Volume_Centre == "Yes", ]
     Low.Volume.Sample <- selected.data[selected.data$High_Volume_Centre == "No", ]

### 2019-08-23 MainCodeRun.R
Add test run line to main code

@@ -121,7 +121,10 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     rmarkdown::render("./ManuscriptMarkdown.Rmd")
 
 }
-MainCodeRun(test = FALSE, clean.start = FALSE, copy.results.to.path = "~/ownCloud/projects/transfer-effect-mistriage-martin/")
+## Uncomment for real run
+## MainCodeRun(test = FALSE, clean.start = FALSE, copy.results.to.path = "~/ownCloud/projects/transfer-effect-mistriage-martin/")
+## Uncomment for test run
+MainCodeRun(test = TRUE, clean.start = TRUE, copy.results.to.path = NULL)
 
 
 

### 2019-08-25 MainCodeRun.R
Model sbp, rr and gcs as continuous linear terms

@@ -31,8 +31,8 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     set.seed(-41892)
     ## Import data
     data.file <- "swetrau-20110101-20160425.csv"
-    if (test)
-        data.file <- "simulated-swetrau-data.csv"
+    ## if (test)
+    ##     data.file <- "simulated-swetrau-data.csv"
     raw.data <- ImportStudyData(data.file)
     ## Create study sample from selected variables
     selected.data <- VariableSelection(raw.data)
@@ -52,11 +52,11 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
                                        abbreviated.label = "Age"),
                      pt_Gender = list(full.label = "Patient gender",
                                       abbreviated.label = "Gender"),
-                     gcs = list(full.label = "Glasgow coma scale",
+                     ed_gcs_sum = list(full.label = "Glasgow coma scale",
                                        abbreviated.label = "GCS"),
-                     sbp = list(full.label = "Systolic blood pressure",
+                     ed_sbp_value = list(full.label = "Systolic blood pressure",
                                          abbreviated.label = "SBP"),
-                     rr = list(full.label = "Respiratory rate",
+                     ed_rr_value = list(full.label = "Respiratory rate",
                                         abbreviated.label = "RR"),
                      res_survival = list(full.label = "30 day survival",
                                          abbreviated.label = "30d survival"),

### 2019-08-25 functions/DataCleaning.R

@@ -30,9 +30,5 @@ DataCleaning <- function(df) {
     df$ISS_over_15 <- df$ISS
     df$ISS_over_15 <- ifelse(df$ISS_over_15 > 15, 1, 2)
     df$ISS_over_15 <- factor(df$ISS_over_15, levels = c(1, 2), labels = c("Yes", "No"))
-    ## Categorise systolic blood pressure, respiratory rate and Glasgow come
-    ## scale using Revised Trauma Score cutoffs
-    df[, c("gcs", "sbp", "rr")] <- bengaltiger::GetRevisedTraumaScoreComponents(df, gcs.name = "ed_gcs_sum", sbp.name = "ed_sbp_value", rr.name = "ed_rr_value", return.labels = TRUE)
-    df[, c("ed_gcs_sum", "ed_sbp_value", "ed_rr_value")] <- NULL
     return(df)
 }

### 2019-08-25 ManuscriptMarkdown.Rmd
Add change in how to model vital signs to manuscript text and add reference

@@
   total.number.of.na <- results$NA.info.sample$multi.centre.vs.single.centre$multi.centre$total.number.of.na
-  n.missing.data.time.of.trauma <- Results$n.missing.date.time.of.trauma
+  n.missing.data.time.of.trauma <- results$n.missing.date.time.of.trauma
   ## Sample NA
@@
 ## Quantitative variables
-GCS was modelled as a continuous linear term, and SBP and RR were modelled using 
-restricted cubic splines with four knots whenever possible, placed at equally spaced 
-percentiles. When describing the participant characteristics, quantitative variables 
-were presented as continuous. ISS was presented as dichotomous, using ISS > 15 as the cutoff.
+GCS, SBP, and RR were modelled as continuous linear terms. In the original study
+plan it was specified that SBP and RR would be modelled using restricted cubic
+splines with four knots. This approach however resulted in an unstable
+imputation model and we therefore decided to simplify the analysis model. ISS
+was presented as dichotomous, using ISS > 15 as the cutoff.
 
@@
 ### Data sets
-The complete SweTrau cohort was split into three overlapping sets of data, each representing a 
-plausible care context for model transfer. These data sets were further divided 
-into samples as follows:
+The complete SweTrau cohort was split into three overlapping sets of data, each
+representing a plausible care context for model transfer. These data sets were
+further divided into samples as follows:
 
@@
 ### Development and validation sample
-The samples in each data set were split into
-two subsamples using a temporal split based on the date of trauma. The earlier
-subsample was the development sample, and the later subsample was the validation
-sample. The development sample contained 70 events (events being patients who
-died within 30 days of the trauma) and all non-events (non-events being patients who
-survived 30 days past the trauma) during the same time. The rationale for
-including 70 events is that at least 10 events per free parameter were needed in
-the logistic regression to obtain stable coefficient estimates [@RN30]. The
-validation sample contained 100 events and at least 100 non-events,
-which was suggested as the minimum number by Vergouwe in 2005 for external
-validation samples [@RN37]. Figure 1 shows the development of the samples using 
-the high- and low-volume centre data set as an example. The minimum sample sizes 
-of the development and validation samples were 140 and 200, respectively. 
-We performed analyses only on data sets for which all samples included at 
-least the minimum sample size.
+The samples in each data set were split into two subsamples using a temporal
+split based on the date of trauma. The earlier subsample was the development
+sample, and the later subsample was the validation sample. The development
+sample contained 75 events (events being patients who died within 30 days of the
+trauma) and all non-events (non-events being patients who survived 30 days past
+the trauma) during the same time. The rationale for including 75 events was that
+25 events per variable have been shown to result in less biased coefficient
+estimates compared to the more standard 10 events per variable [@RN30;
+@Courvoisier2011]. The validation sample contained 100 events and at least 100
+non-events, which was suggested as the minimum number by Vergouwe in 2005 for
+external validation samples [@RN37]. Figure 1 shows the development of the
+samples using the high- and low-volume centre data set as an example. The
+minimum sample sizes of the development and validation samples were 140 and 200,
+respectively.  We performed analyses only on data sets for which all samples
+included at least the minimum sample size.
 
@@
 The programming language R was used for all analyses [@RN31], and all code has
-been made publicly available [@RN59]. Analyses were performed in the sequence
-of model development, model validation and finally model comparison. These steps
+been made publicly available [@RN59]. Analyses were performed in the sequence of
+model development, model validation and finally model comparison. These steps
 were repeated in each data set. Below, we use the transfer of a model from the
-high-volume centre sample to the low-volume centre sample as an example to describe 
-the complete procedure.
+high-volume centre sample to the low-volume centre sample as an example to
+describe the complete procedure.
 
@@
 regression. The dependent variable was all-cause mortality within 30 days of
-trauma, and the independent variables SBP, RR, and GCS were modelled as previously described. 
-To avoid overfitting the model, a bootstrap procedure was used to estimate a linear 
-shrinkage factor that was then applied to the model coefficients [@RN61]. The shrunk 
-model was used to estimate the probability of all-cause 30-day mortality in the 
-development sample. A grid search was performed across estimated probabilities in 
-the development sample to identify the cutoff value that optimised overtriage while keeping 
-undertriage at less than 5% [@RN39]. This cutoff value was then used to classify patients 
-as having major or minor trauma.
+trauma, and the independent variables SBP, RR, and GCS were modelled as
+previously described.  To avoid overfitting the model, a bootstrap procedure was
+used to estimate a linear shrinkage factor that was then applied to the model
+coefficients [@RN61]. The shrunk model was used to estimate the probability of
+all-cause 30-day mortality in the development sample. A grid search was
+performed across estimated probabilities in the development sample to identify
+the cutoff value that optimised overtriage while keeping undertriage at less
+than 5% [@RN39]. This cutoff value was then used to classify patients as having
+major or minor trauma.
 
@@
 ### Model comparison
-Finally, the difference in model performance
-between the high- and low-volume centre validation samples was
-calculated. Empirical bootstrapping was used to estimate the 95% confidence intervals
-(CI) around the performance and the differences in the performance estimates. Both bootstrap
-procedures used 1 000 bootstrap samples drawn with replacement of the same size
-as the original samples. The three steps of model development, model validation,
-and model comparison were repeated in all four sets of data.
+Finally, the difference in model performance between the high- and low-volume
+centre validation samples was calculated. Empirical bootstrapping was used to
+estimate the 95% confidence intervals (CI) around the performance and the
+differences in the performance estimates. Both bootstrap procedures used 1 000
+bootstrap samples drawn with replacement of the same size as the original
+samples. The three steps of model development, model validation, and model
+comparison were repeated in all four sets of data.
 
@@
 ### Missing data
-We used multiple imputation using chained equations (mice) to handle missing data [@RN41]. 
-The number of imputations created for each data set was equal to the highest 
-percentage of missing data in that data set. Quantitative variables were imputed 
-using predictive mean matching, and qualitative variables were imputed using logistic 
-regression. SBP and RR were transformed as restricted cubic splines before imputation. 
-All analyses outlined above were then conducted separately in
-each imputed dataset. We present the combined results as the median point
-estimate across imputations along with an empirical bootstrap of the 25th and
-75th percentiles across imputations, i.e. the lower bound of the presented
-interval is the lower bound of a 95% CI of the 25th percentile, and the upper
-bound is the upper bound of a 95% CI of the 75th percentile. This combined CI
-was referred to as an uncertainty interval (UI) and was used to express the
-added uncertainty associated with the imputation procedure to handle missing
-data; as such, it is more conservative than a standard 95% CI.
+We used multiple imputation using chained equations (mice) to handle missing
+data [@RN41].  The number of imputations created for each data set was equal to
+the highest percentage of missing data in that data set. Quantitative variables
+were imputed using predictive mean matching, and qualitative variables were
+imputed using logistic regression.  All analyses outlined above were then
+conducted separately in each imputed dataset. We present the combined results as
+the median point estimate across imputations along with an empirical bootstrap
+of the 25th and 75th percentiles across imputations, i.e. the lower bound of the
+presented interval is the lower bound of a 95% CI of the 25th percentile, and
+the upper bound is the upper bound of a 95% CI of the 75th percentile. This
+combined CI was referred to as an uncertainty interval (UI) and was used to
+express the added uncertainty associated with the imputation procedure to handle
+missing data; as such, it is more conservative than a standard 95% CI.
 

### 2019-08-25 functions/DevValCreator.R
Use 75 events in development data

@@ -3,7 +3,7 @@
 #' Inputs a data.frame (e.g. "high.volume")
 #' Splits the data frame based on the date of trauma to development 
 #' sample (early) and validation sample (late).
-#' Development sample will contain 70 events(res_survival == "Dead").
+#' Development sample will contain 75 events(res_survival == "Dead").
 #' Validation sample will contain the rest of the events.
 #' @param df Dataframe. A study sample.
 DevValCreator <- function(df) {
@@ -15,7 +15,7 @@ DevValCreator <- function(df) {
     events.only <- df[which(df$res_survival == "Dead"),]
     
     ## Sorting by date 
-    sorted.events.only <- tryCatch(expr = events.only[order(events.only$DateTime_Of_Trauma), decreasing = FALSE),],
+    sorted.events.only <- tryCatch(expr = events.only[order(events.only$DateTime_Of_Trauma, decreasing = FALSE),],
                                    error = function(e) {
                                        .imp <- unique(df$.imp)
                                        filename <- paste0("output/failed.imputation.", .imp, ".", gsub(":|-| ", "", format(Sys.time())), ".Rds")
@@ -24,7 +24,7 @@ DevValCreator <- function(df) {
                                    })
    
     ## Saving cut-off date
-    cut.off.date <- sorted.events.only$DateTime_Of_Trauma[120]
+    cut.off.date <- sorted.events.only$DateTime_Of_Trauma[75]
     ##Split data frame
     dev.sample <- df[df$DateTime_Of_Trauma <= cut.off.date, ]
     val.sample <- df[df$DateTime_Of_Trauma > cut.off.date, ]

### 2019-08-25 functions/IndividualCentreCheck.R
Modify number events needed to be eligible as single centre to 175

@@ -16,7 +16,7 @@ IndividualCentreCheck <- function(df) {
   
     ## Evaluates if Sjukhuskod is valid individual centre
     x <- as.data.frame(table(df$Sjukhuskod, df$res_survival, exclude = c(2, NA, NaN, 999, "Alive")))
-    y <- as.vector(x$Var1[x$Freq >= 220])
+    y <- as.vector(x$Var1[x$Freq >= 175])
   
     df$Valid_Individual_Centre <- ifelse(df$Sjukhuskod %in% y, "Yes", "No")
   

### 2019-08-25 functions/MICEImplement.R
Modify imputation model and add naive error checks

@@ -31,8 +31,13 @@ MICEImplement <- function(df.list, test = FALSE) {
 
     ## Run mice
     df.list <- lapply(df.list, function(df) {
-        temporary.data <- mice(df, m = number.of.imputations, maxit = maxit) 
+        df$ISS_over_15 <- NULL
+        df$Sjukhuskod <- as.factor(df$Sjukhuskod)
+        temporary.data <- mice(df, m = number.of.imputations, maxit = maxit)
+        print(temporary.data$loggedEvents)
         df <- complete(temporary.data, action = "long", include = TRUE)
+        df$ISS_over_15 <- table(factor(df$ISS > 15, labels = c("No", "Yes")))
+        print(sum(is.na(df[df[, ".imp"] != 0, ])))
         return(df)
     })
 

### 2019-08-28 MainCodeRun.R
Optimise to run on cluster with OpenMPI

@@ -11,6 +11,10 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
                         copy.results.to.path = NULL) {
     
     ## INITIALIZING
+    ## Start the MPI cluster to prevent slaves to execute master code
+    library(doMPI)
+    study.cluster <- startMPIcluster()
+    registerDoMPI(study.cluster)
     ## Load required packages and source functions
     ## FuncPack() first needs to be sourced to run
     source("./functions/FuncPack.R")
@@ -97,9 +101,7 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     message (paste0("Estimating results in bootstrap samples ", paste0(estimated.bootstraps, collapse = ", ")))
     ## Get bootstrap results
     errorlog <- "output/errorlog.txt"
-    write("", errorlog)
-    study.cluster <- makeCluster(detectCores())
-    registerDoParallel(study.cluster)
+    write(Sys.time(), errorlog)
     foreach(bootstrap.sample = bootstrap.samples,
             .packages = FuncPack(return.only = TRUE)$packages,
             .export = FuncPack(return.only = TRUE)$functions) %dopar% {
@@ -112,7 +114,8 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
                                                    errorlog,
                                                    append = TRUE))
             }
-    stopCluster(study.cluster)
+    closeCluster(study.cluster)
+    mpi.quit()
     ## Report all analyses completed
     message("All analyses completed")
     ## Compile results

### 2019-08-28 functions/FuncPack.R

@@ -7,20 +7,25 @@ FuncPack <- function(return.only = FALSE) {
     ## Get specific bengaltiger version
     library(devtools)
     install_github("martingerdin/bengaltiger@81672faabfb2d18dd66a92a7c0c4f32b2477d2f4")
+    library(bengaltiger)
     
     ## Load required packages
     packages <- c("data.table",
                   "mice",
                   "Hmisc",
                   "boot",
-                  "bengaltiger",
-                  "doParallel",
                   "foreach",
                   "knitr",
                   "rmarkdown")
 
-    if (!return.only)
-        for(package in packages) library(package, character.only = TRUE)
+    if (!return.only) {
+        for(package in packages) {
+            tryCatch(expr = library(package, character.only = TRUE),
+                     error = function(e) install.packages(package, repos = "https://cloud.r-project.org"),
+                     finally = library(package, character.only = TRUE))
+        }
+        
+    }
    
     ## Source functions
     files <- list.files("./Functions", full.names = TRUE, pattern = ".R$")

### 2019-08-28 runOnCluster.sh

@@ -1,0 +1,10 @@
+#!/bin/sh
+#SBATCH -t 02:00:00
+#SBATCH -N 1
+#SBATCH --tasks-per-node=20
+#SBATCH --exclusive
+#SBATCH --mail-user=martingerdin@gmail.com
+#SBATCH --mail-type=END
+
+mpirun R --slave -f MainCodeRun.R
+

### 2019-09-05 MainCodeRun.R
Add log

@@ -15,6 +15,7 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     library(doMPI)
     study.cluster <- startMPIcluster(verbose = TRUE, logdir = "log")
     registerDoMPI(study.cluster)
+    write(paste0("Log ", as.character(Sys.time()), " \n"), "log.out")
     ## Load required packages and source functions
     ## FuncPack() first needs to be sourced to run
     source("./functions/FuncPack.R")
@@ -100,8 +101,6 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     }
     message (paste0("Estimating results in bootstrap samples ", paste0(estimated.bootstraps, collapse = ", ")))
     ## Get bootstrap results
-    errorlog <- "output/errorlog.txt"
-    write(Sys.time(), errorlog)
     foreach(bootstrap.sample = bootstrap.samples,
             .packages = FuncPack(return.only = TRUE)$packages,
             .export = FuncPack(return.only = TRUE)$functions) %dopar% {
@@ -110,8 +109,8 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
                                          boot = TRUE,
                                          test = test,
                                          copy.results.to.path = copy.results.to.path),
-                         error = function(e) write(paste0("bootstrap sample ", boot.id, ": ", e),
-                                                   errorlog,
+                         error = function(e) write(paste0("bootstrap sample ", boot.id, ": ", e, " \n"),
+                                                   "log.out",
                                                    append = TRUE))
             }
     closeCluster(study.cluster)
@@ -125,6 +124,6 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     mpi.quit()
 }
 ## Uncomment for real run
-MainCodeRun(test = FALSE, clean.start = FALSE, copy.results.to.path = NULL)
+## MainCodeRun(test = FALSE, clean.start = FALSE, copy.results.to.path = NULL)
 ## Uncomment for test run
-## MainCodeRun(test = TRUE, clean.start = TRUE, copy.results.to.path = NULL)
+MainCodeRun(test = TRUE, clean.start = TRUE, copy.results.to.path = NULL)

### 2019-09-05 functions/RunStudy.R

@@ -24,6 +24,9 @@ RunStudy <- function(selected.data, codebook = NULL, boot = FALSE, test = FALSE,
     if (!is.null(copy.results.to.path) & !is.character(copy.results.to.path))
         stop ("copy.results.to.path has to be character or NULL")
 
+    ## Log
+    write(paste0("Bootstrap sample ", unique(selected.data[, ".boot.id"]), "started \n"), "log.out", append = TRUE)
+    
     ## Create results list
     Results <- list()
 
@@ -137,6 +140,8 @@ RunStudy <- function(selected.data, codebook = NULL, boot = FALSE, test = FALSE,
     saveRDS(Results, file.name)
     if (!is.null(copy.results.to.path))
         saveRDS(Results, paste0(copy.results.to.path, file.name))
-    return(paste0(file.name, " saved to disk"))
+    message(paste0(file.name, " saved to disk"))
 
+    ## Log
+    write(paste0("Bootstrap sample ", unique(selected.data[, ".boot.id"]), "completed \n"), "log.out", append = TRUE)
 }

### 2019-09-05 MainCodeRun.R
Fix error in how completed bootstraps are calculated

@@ -27,11 +27,10 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     }
     ## Identify starting point and go from there if clean start is FALSE
     original.results.done <- file.exists("./output/original.results.Rds")
-    completed.bootstraps <- NULL
-    if (!clean.start) {
-        bootstraps.done <- list.files("./output/", "^bootstrap\\.sample\\.[0-9]*\\.results.Rds$")
-        completed.bootstraps <- sort(as.numeric(gsub("bootstrap\\.sample\\.|\\.results\\.Rds", "", bootstraps.done)))
-    }
+    bootstraps.done <- list.files("./output/", "^bootstrap\\.sample\\.[0-9]*\\.results.Rds$")
+    completed.bootstraps <- sort(as.numeric(gsub("bootstrap\\.sample\\.|\\.results\\.Rds", "", bootstraps.done)))
+    if (length(completed.bootstraps) == 0)
+        completed.bootstraps <- NULL
     ## Set random seed
     set.seed(-41892)
     ## Import data

### 2019-09-05 testRunOnCluster.sh
Add shell script to test on cluster

@@ -1,0 +1,12 @@
+#!/bin/sh
+#SBATCH -t 01:00:00
+#SBATCH -N 1
+#SBATCH --tasks-per-node=20
+#SBATCH --exclusive
+#SBATCH --qos=test
+#SBATCH --mail-user=martingerdin@gmail.com
+#SBATCH --mail-type=END
+
+module load GCC/7.3.0-2.30 OpenMPI/3.1.1 R/3.5.1
+
+mpirun --bind-to core R --slave -f MainCodeRun.R

### 2019-09-05 MainCodeRun.R
Add additional use cases

@@ -122,7 +122,11 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     ## Quit MPI
     mpi.quit()
 }
-## Uncomment for real run
-## MainCodeRun(test = FALSE, clean.start = FALSE, copy.results.to.path = NULL)
 ## Uncomment for test run
 MainCodeRun(test = TRUE, clean.start = TRUE, copy.results.to.path = NULL)
+
+## Uncomment for first real run
+## MainCodeRun(test = FALSE, clean.start = TRUE, copy.results.to.path = NULL)
+
+## Uncomment to resume real run
+## MainCodeRun(test = FALSE, clean.start = TRUE, copy.results.to.path = NULL)

### 2019-09-05 MainCodeRun.R
Add log line to original results

@@ -75,6 +75,7 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     ## Get original results
     if (clean.start | !original.results.done) {
         message ("Estimating original results")
+        write("Estimating original results \n", "log.out")
         RunStudy(selected.data = selected.data, copy.results.to.path = copy.results.to.path, codebook = codebook, boot = FALSE, test = test)
     }
     ## Create bootstrap samples

### 2019-09-05 MainCodeRun.R
Change clean.start parameter when resuming real run

@@ -130,4 +130,4 @@ MainCodeRun(test = TRUE, clean.start = TRUE, copy.results.to.path = NULL)
 ## MainCodeRun(test = FALSE, clean.start = TRUE, copy.results.to.path = NULL)
 
 ## Uncomment to resume real run
-## MainCodeRun(test = FALSE, clean.start = TRUE, copy.results.to.path = NULL)
+## MainCodeRun(test = FALSE, clean.start = FALSE, copy.results.to.path = NULL)

### 2019-09-05 functions/RunStudy.R
Correct log lines when analysing bootstrap samples

@@ -25,7 +25,8 @@ RunStudy <- function(selected.data, codebook = NULL, boot = FALSE, test = FALSE,
         stop ("copy.results.to.path has to be character or NULL")
 
     ## Log
-    write(paste0("Bootstrap sample ", unique(selected.data[, ".boot.id"]), "started \n"), "log.out", append = TRUE)
+    if (boot)
+        write(paste0("Bootstrap sample ", unique(selected.data[, ".boot.id"]), "started \n"), "log.out", append = TRUE)
     
     ## Create results list
     Results <- list()
@@ -143,5 +144,6 @@ RunStudy <- function(selected.data, codebook = NULL, boot = FALSE, test = FALSE,
     message(paste0(file.name, " saved to disk"))
 
     ## Log
-    write(paste0("Bootstrap sample ", unique(selected.data[, ".boot.id"]), "completed \n"), "log.out", append = TRUE)
+    if (boot)
+        write(paste0("Bootstrap sample ", unique(selected.data[, ".boot.id"]), "completed \n"), "log.out", append = TRUE)
 }

### 2019-09-06 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
   total.number.of.na <- results$NA.info.sample$multi.centre.vs.single.centre$multi.centre$total.number.of.na
-  n.missing.data.time.of.trauma <- results$n.missing.date.time.of.trauma
+  n.missing.date.time.of.trauma <- results$n.missing.date.time.of.trauma
   ## Sample NA

### 2019-09-07 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 ## TABLE SETUP
+  # Capitalize names before table creation
+  names(results.all.data) <- c("High-volume", "Low-volume", "Metropolitan", "Non-metropolitan", "Multiple centres", "Single centre")
   # Table 1 = Sample characteristics table
@@
 
-  **Conclusions:** In all samples, model transfer was associated with high 
-  levels of uncertainty. Notably, transfer from the `r colnames(highest.comp.mistriage)` 
-  sample to the other sample in the data set led to an increased mistriage
-  rate, primarily due to increased overtriage.
+  **Conclusions:** In all samples, model transfer was associated with some degree of uncertainty. 
+  Models developed in samples with a high number of patients performed the best after transfer. Affirming 
+  this, the worst performance after transfer was the low-volume model transferred to the high-volume 
+  sample, resulting in a significant mistriage increase of `r highest.comp.mistriage`, this was
+  primarily due to increased overtriage.
 
@@
 samples using the high- and low-volume centre data set as an example. The
-minimum sample sizes of the development and validation samples were 140 and 200,
+minimum sample sizes of the development and validation samples were 150 and 200,
 respectively.  We performed analyses only on data sets for which all samples
@@
 trauma care between different contexts within a single health system affect
-mistriage rates. In all data sets, model transfer between samples resulted in
-substantial uncertainty in terms of its effect on model performance. When comparing the
-transferred performance to the receiving sample validation performance, the
-resulting differences also showed considerable uncertainty.
+mistriage rates. In many data sets, varying degrees of uncertainty were observed 
+when comparing the transferred performance to the receiving sample validation performance.
 
-The most notable effects on model performance following model transfer were observed
+The most notable effect on model performance following model transfer were observed
 after transferring the `r colnames(highest.comp.mistriage)` model. This transfer
@@
 `r substr(format(highest.comp.mistriage), 1, 5)`. Mainly contributing to this was
-a marked increase in overtriage.  For this transfer, the 95% UI ranged from 
-`r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating a high level of uncertainty, suggesting
-that our findings are compatible with a `r UI.top.bot[1]` decrease to a 
-`r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r UI.top.bot[1]`
-would signify an improvement in model performance. The higher 95% UI of 
-`r UI.top.bot[2]` would mean a marked increase in model mistriage compared to the
-point estimate of `r substr(format(highest.comp.mistriage), 1, 5)`.
+an increase in overtriage.  For this transfer, the 95% UI ranged from 
+`r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating a high level of uncertainty. 
+In contrast, the transfer of the high-volume model to the low-volume sample 
+reduced mistriage, this was primarily due to reduced overtriage.
 
-Due to centralized trauma care, non-metropolitan centres generally have fewer 
-resources and less experience receiving trauma patients. They may rely 
-on clinical prediction models more heavily for accurate trauma triage, and 
-an increase in overtriage can lead to further strains on already limited 
-resources. Increased undertriage could possibly lead to increased patient 
-mortality.
+Low-volume centres do by definition receive fewer trauma patients, and therefore 
+less data is available to establish accurate clinical prediction models. It is 
+reasonable to assume this is the reason for the marked increase in mistriage 
+when transferring the low-volume model to the high-volume sample. This also 
+explains why the low-volume model and the single-centre model showed the highest 
+levels of mistriage during model validation (i.e. when applied to the sample in 
+which the model was created). Further confirming this, the transfer of the 
+high-volume model to the low-volume sample reduces mistriage.
 
-Recognizing these risks and updating models accordingly could lead
+In a clinical context, this seems to suggest clinical prediction models for early 
+trauma care should be developed in settings with a high number of trauma patients. 
+These models could then be transferred to settings with less trauma patients. Low-volume 
+centres generally have fewer resources and less experience receiving trauma patients, 
+these centres could therefore benefit by using models developed in centres with more 
+experience receiving trauma patients
+
+Recognizing settings with high mistriage and updating models accordingly could lead
 to increased accuracy during trauma patient triage, potentially saving lives by
@@
 potential model transfers: clinical prediction models for trauma are usually
-developed in large metropolitan centres and then transferred as clinical
-recommendations to minor non-metropolitan centres. The results should therefore
-be of practical importance.
+developed in large metropolitan centres with a high volume of patients and 
+then transferred as clinical recommendations to minor low-volume centres. The 
+results should therefore be of practical importance.
 
@@
 ## Conclusion
-Model transfer resulted in large and varying levels of uncertainty in in terms of its effect
-on mistriage. Therefore, model transfer can be unpredictable, potentially leading 
-to increased mistriage but in some cases also to a decrease in mistriage. 
-Noticeably, transfer of the `r colnames(highest.comp.mistriage)` model led to an 
-increased mistriage of `r highest.comp.mistriage`. This observed effect was also 
-associated with considerable uncertainty. Recognizing the limitations of this study, 
+Model transfer resulted in varying levels of uncertainty in in terms of its effect
+on mistriage. In general, samples with a high number of patients had the lowest mistriage 
+both during validation and after transfer to the other sample in the data set. The opposite 
+was also true, and the transfer of the `r colnames(highest.comp.mistriage)` 
+model led to an increased mistriage of `r highest.comp.mistriage`, noting that this observation 
+was associated with considerable uncertainty. Recognizing the limitations of this study, 
 the authors believe that further studies are warranted due to the potential economic 

### 2019-09-07 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 - Martin Gerdin Wärnberg
-date: "August 14, 2019"
+date: "September 07, 2019"
 documentclass: "article"

### 2019-09-08 MainCodeRun.R
Update MainCodeRun.R

@@ -52,18 +52,18 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     ## Mark entries as valid (>170 events) individual centres
     selected.data <- IndividualCentreCheck(selected.data)
     ## Define codebook
-    codebook <- list(pt_age_yrs = list(full.label = "Patient age",
-                                       abbreviated.label = "Age"),
-                     pt_Gender = list(full.label = "Patient gender",
-                                      abbreviated.label = "Gender"),
+    codebook <- list(pt_age_yrs = list(full.label = "Patient age, years",
+                                       abbreviated.label = "Age, years"),
+                     pt_Gender = list(full.label = "Patient sex",
+                                      abbreviated.label = "Sex"),
                      ed_gcs_sum = list(full.label = "Glasgow coma scale",
                                        abbreviated.label = "GCS"),
                      ed_sbp_value = list(full.label = "Systolic blood pressure",
                                          abbreviated.label = "SBP"),
                      ed_rr_value = list(full.label = "Respiratory rate",
                                         abbreviated.label = "RR"),
-                     res_survival = list(full.label = "30 day survival",
-                                         abbreviated.label = "30d survival"),
+                     res_survival = list(full.label = "30-day survival",
+                                         abbreviated.label = "30-day survival"),
                      ISS = list(full.label = "Injury severity score",
                                 abbreviated.label = "ISS"),
                      NISS = list(full.label = "New injury severity score",

### 2019-09-09 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

--- 2019-09-07 ManuscriptMarkdown.Rmd
@@
 - Martin Gerdin Wärnberg
-date: "August 14, 2019"
+date: "September 09, 2019"
 documentclass: "article"
@@
 trauma care between different contexts within a single health system affect
-mistriage rates. In many data sets, varying degrees of uncertainty were observed 
-when comparing the transferred performance to the receiving sample validation performance.
+mistriage rates. Varying degrees of uncertainty were observed when comparing 
+the transferred performance to the receiving sample validation performance.
 
-The most notable effect on model performance following model transfer were observed
+The most notable effect on model performance following model transfer was observed
 after transferring the `r colnames(highest.comp.mistriage)` model. This transfer
-resulted in an increase in the mistriage rate of 
-`r substr(format(highest.comp.mistriage), 1, 5)`. Mainly contributing to this was
+resulted in an increased mistriage rate of `r substr(format(highest.comp.mistriage), 1, 4)`.
+Mainly contributing to this was
 an increase in overtriage.  For this transfer, the 95% UI ranged from 
@@
 In contrast, the transfer of the high-volume model to the low-volume sample 
-reduced mistriage, this was primarily due to reduced overtriage.
+reduced mistriage, primarily due to reduced overtriage.
 
@@
 which the model was created). Further confirming this, the transfer of the 
-high-volume model to the low-volume sample reduces mistriage.
+high-volume model to the low-volume sample reduced mistriage.
 
-In a clinical context, this seems to suggest clinical prediction models for early 
+In a clinical context, this suggests that clinical prediction models for early 
 trauma care should be developed in settings with a high number of trauma patients.
-These models could then be transferred to settings with less trauma patients. Low-volume 
-centres generally have fewer resources and less experience receiving trauma patients, 
-these centres could therefore benefit by using models developed in centres with more 
-experience receiving trauma patients
+These models could then be transferred to settings with less trauma patients, which 
+should then result in a decreased mistriage rate in those settings.
 
-Recognizing settings with high mistriage and updating models accordingly could lead
-to increased accuracy during trauma patient triage, potentially saving lives by
-minimizing undertriage, and likely saving resources by minimizing overtriage. With
-registries such as SweTrau, the analysis of model transfer could be performed
-continuously to optimize the models currently in use and advise which model
-transfers might be detrimental (or potentially beneficial).
+In 2018, Granström et al. found that using a criteria-based triage system based on a 
+clinical prediction model was effective in reducing overtriage without increasing 
+mortality [@RN44]. Therefore, improving clinical prediction models to reduce mistriage is
+clinically beneficial, potentially saving lives by minimizing undertriage, and likely 
+saving resources by minimizing overtriage. Possibly, with registries as SweTrau, the analysis of 
+model transfer could be performed continuously to optimize the models currently in use 
+and advise which model transfers might be detrimental (or potentially beneficial).
 
@@
 
-External validation as the evaluation of a clinical prediction model in a
-setting in which it was not originally developed has been studied more
+External validation - the evaluation of a clinical prediction model in a
+setting in which it was not originally developed - has been studied more
 extensively. Studies using simulated clinical prediction models based on
@@
 this study, the studies on model transfer and the studies on external validation
-would suggest caution when transferring models.
+would suggest indiscriminate model transfers could lead to detrimental clinical
+consequences, but also that clinically beneficial model transfers could be
+identified.
 
@@
 ### Strengths
-The main strength of this study is that its design realistically reflects
+The study design realistically reflects
 potential model transfers: clinical prediction models for trauma are usually

### 2019-09-09 functions/CompileResults.R
Update CompileResults.R

@@ -71,7 +71,7 @@ CompileResults <- function() {
         colnames(point.estimates) <- colnames(lbs) <- colnames(ubs) <- colnames
         combined <- data.frame(t(rbind(point.estimates, lbs, ubs)))
         colnames(combined) <- c("pe", "lb", "ub")
-        combined[] <- lapply(combined, function(column) sprintf("%.3f", column))
+        combined[] <- lapply(combined, function(column) sprintf("%.2f", column))
         rownames <- rownames(combined)
         combined <- data.frame(with(combined, paste0(pe, " (", lb, " - ", ub, ")")))
         combined <- cbind(rownames, combined)

### 2019-09-10 MainCodeRun.R
Update default run

@@ -124,10 +124,10 @@ MainCodeRun <- function(test = FALSE, clean.start = TRUE,
     mpi.quit()
 }
 ## Uncomment for test run
-MainCodeRun(test = TRUE, clean.start = TRUE, copy.results.to.path = NULL)
+## MainCodeRun(test = TRUE, clean.start = TRUE, copy.results.to.path = NULL)
 
 ## Uncomment for first real run
 ## MainCodeRun(test = FALSE, clean.start = TRUE, copy.results.to.path = NULL)
 
 ## Uncomment to resume real run
-## MainCodeRun(test = FALSE, clean.start = FALSE, copy.results.to.path = NULL)
+MainCodeRun(test = FALSE, clean.start = FALSE, copy.results.to.path = NULL)

### 2019-09-10 functions/FuncPack.R
Update bengaltiger version

@@ -6,7 +6,7 @@
 FuncPack <- function(return.only = FALSE) {
     ## Get specific bengaltiger version
     library(devtools)
-    ## install_github("martingerdin/bengaltiger@81672faabfb2d18dd66a92a7c0c4f32b2477d2f4")
+    ## install_github("martingerdin/bengaltiger@aea49457aa68a4d00c0045cb22db582f09727a18")
     library(bengaltiger)
     
     ## Load required packages

### 2019-09-10 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 - Martin Gerdin Wärnberg
-date: "September 09, 2019"
+date: "September 10, 2019"
 documentclass: "article"
@@
 These models could then be transferred to settings with less trauma patients, which 
-should then result in a decreased mistriage rate in those settings.
-
-In 2018, Granström et al. found that using a criteria-based triage system based on a 
-clinical prediction model was effective in reducing overtriage without increasing 
-mortality [@RN44]. Therefore, improving clinical prediction models to reduce mistriage is
-clinically beneficial, potentially saving lives by minimizing undertriage, and likely 
-saving resources by minimizing overtriage. Possibly, with registries as SweTrau, the analysis of 
-model transfer could be performed continuously to optimize the models currently in use 
-and advise which model transfers might be detrimental (or potentially beneficial).
+should then result in a decreased mistriage rate in those settings. In 2018, Granström 
+et al. found that using a criteria-based triage system (i.e. a clinical prediction model) 
+for trauma triage was effective in reducing overtriage without increasing mortality [@RN44]. 
+Therefore, improving clinical prediction models to reduce mistriage is clinically beneficial, 
+potentially saving lives by minimizing undertriage, and likely saving resources by 
+minimizing overtriage. Possibly, with registries as SweTrau, the analysis of model 
+transfer could be performed continuously to optimize the models currently in use and 
+advise which model transfers might be detrimental (or potentially beneficial).
 

### 2019-09-10 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

--- 2019-09-09 ManuscriptMarkdown.Rmd
@@
 
-**Table 2. Model validation performance, i.e. performance when applied to the sample in which the model was created.** Data are presented as point estimates with 95% uncertainty intervals (95% UI).
+**Table 2.
+Model validation performance, i.e.
+performance when applied to the validation subsample of the sample in which the model was created.** Data are presented as point estimates with 95% uncertainty intervals (95% UI).
 ```{r, echo = FALSE, results = "asis"} 
@@
 
-**Table 3. Model transfer performance, i.e. performance when transferred to the other sample in the same data set.** Data are presented as point estimates with 95% uncertainty intervals (95% UI).  
+**Table 3.
+Model transfer performance, i.e.
+performance when transferred to the other validation subsample in the same data set.** Data are presented as point estimates with 95% uncertainty intervals (95% UI).
 ```{r, echo = FALSE, results = "asis"} 

### 2019-09-18 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 - Martin Gerdin Wärnberg
-date: "September 09, 2019"
+date: "September 18, 2019"
 documentclass: "article"
@@
 undertriage and overtriage rate for the same model was `r val.undertriage` and
-`r val.overtriage`, respectively. The worst validation performance (i.e. highest
-mistriage rate) was found in the `r colnames(highest.val.mistriage)` sample, with
+`r val.overtriage`, respectively. The highest validation mistriage rate (i.e. worst
+validation performance) was found in the `r colnames(highest.val.mistriage)` sample, with
 a model median mistriage rate of `r highest.val.mistriage`.
@@
 Model performance after transfer was determined for each model after being transferred
-to the other sample in the data set, in all data sets (Table 3). When
+to the validation subsample in the other sample, in the data set (Table 3). When
 transferred, the model with the lowest mistriage rate was the `r colnames(lowest.tran.mistriage)` 
@@
 Table 4 shows the transferred model performance when compared to the validation
-performance in the sample receiving the transferred model.
+performance in the validation subsample receiving the transferred model.
 The highest mistriage rate when compared to the receiving sample validation

### 2019-09-24 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

@@
 - Martin Gerdin Wärnberg
-date: "September 18, 2019"
+date: "September 24, 2019"
 documentclass: "article"
@@
 
-  **Background:** Trauma results in approximately 3.8 million deaths annually. Clinical
+  **Background:** Trauma results in approximately four million deaths annually. Clinical
   prediction models are often used for triage in early trauma care. Previous research has
@@
 
-  **Results:** A total of `r total.number.of.patients ` patients were included. The model
+  **Results:** A total of `r total.number.of.patients ` patients were included. Model
   validation mistriage rate ranged from `r lowest.val.mistriage` 
   to `r highest.val.mistriage`, and the transferred model
-  mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage` 
-  across all samples. Performance after transfer when compared to the receiving sample validation 
-  performance resulted in substantial uncertainty in all samples. The worst performance after 
-  transfer was found in the `r colnames(highest.comp.mistriage)` 
-  sample, with a mistriage difference of `r highest.comp.mistriage`.
+  mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage`, 
+  across all samples. The worst performance after transfer was the low-volume model transferred 
+  to the high-volume sample, resulting in a significant mistriage rate increase of `r highest.comp.mistriage`, 
+  primarily due to increased overtriage.
 
   **Conclusions:** In all samples, model transfer was associated with some degree of uncertainty. 
-  Models developed in samples with a high number of patients performed the best after transfer. Affirming 
-  this, the worst performance after transfer was the low-volume model transferred to the high-volume 
-  sample, resulting in a significant mistriage increase of `r highest.comp.mistriage`, this was
-  primarily due to increased overtriage.
+  Models developed in samples with high numbers of patients performed best after transfer, and the
+  model developed in the low-volume sample performed worst after transfer.
 
@@
 accounts for approximately 458.5 million cases annually across the globe [@RN999] 
-and an estimated 3.8 million deaths [@RN1; @RN21]. Each year, 9% of global 
+and around four million deaths [@RN1; @RN21]. Each year, 9% of global 
 deaths are the result of trauma, with the leading causes being road traffic accidents, 

### 2019-09-28 ManuscriptMarkdown.Rmd

@@
 - Martin Gerdin Wärnberg
-date: "September 24, 2019"
+date: "September 28, 2019"
 documentclass: "article"
@@
 
-  **Background:** Trauma results in approximately four million deaths annually. Clinical
+  **Abstract**
+
+  *Background:* Trauma results in approximately four million deaths annually. Clinical
   prediction models are often used for triage in early trauma care. Previous research has
@@
 
-  **Methods:** Patients aged 15 years or older who were registered between 2011 and
+  *Methods:* Patients aged 15 years or older who were registered between 2011 and
   2016 in the Swedish national trauma registry, SweTrau, were included. Three sets
@@
 
-  **Results:** A total of `r total.number.of.patients ` patients were included. Model
+  *Results:* A total of `r total.number.of.patients ` patients were included. Model
   validation mistriage rate ranged from `r lowest.val.mistriage` 
@@
 
-  **Conclusions:** In all samples, model transfer was associated with some degree of uncertainty. 
+  *Conclusions:* In all samples, model transfer was associated with some degree of uncertainty. 
   Models developed in samples with high numbers of patients performed best after transfer, and the

### 2019-10-08 ManuscriptMarkdownAnon.Rmd
Add files via upload

--- 2019-09-28 ManuscriptMarkdown.Rmd
@@
 title: "Transferability of clinical prediction models for early trauma care"
-author: 
-- Martin Henriksson
-- Dell D Saulnier
-- Johanna Berg
-- Martin Gerdin Wärnberg
-date: "September 28, 2019"
+date: "October 08, 2019"
 documentclass: "article"
@@
   *Background:* Trauma results in approximately four million deaths annually. Clinical
-  prediction models are often used for triage in early trauma care. Previous research has
-  indicated that transferring these models between different settings might adversely affect 
-  model performance.
+  prediction models are often used for triage in early trauma care. Transferring these 
+  models between different settings might adversely affect model performance.
 
-  *Methods:* Patients aged 15 years or older who were registered between 2011 and
-  2016 in the Swedish national trauma registry, SweTrau, were included. Three sets
-  of data were created: high- and low-volume centres, metropolitan and
+  *Methods:* Patients aged 15 years or older, registered between 2011-2016 
+  in the Swedish national trauma registry, were included. Three data sets 
+  were created: high- and low-volume centres, metropolitan and
   non-metropolitan centres, and multiple and single centres. Clinical
-  prediction models were developed using logistic regression in each data set 
-  and transferred within sets. Model performance was evaluated in all data
-  sets using the mistriage rate, undertriage rate and overtriage rate. Multiple
-  imputation using chained equations was used to handle missing data. Model
-  performance was reported as medians with the 95% uncertainty interval across
-  imputed data sets.
+  prediction models were developed using logistic regression, and transferred within sets. 
+  Model performance was evaluated in all data sets using the mistriage rate, undertriage 
+  rate and overtriage rate. Model performance was reported as medians with the 95% uncertainty 
+  interval across imputed data sets.
 
-  *Results:* A total of `r total.number.of.patients ` patients were included. Model
-  validation mistriage rate ranged from `r lowest.val.mistriage` 
-  to `r highest.val.mistriage`, and the transferred model
+  *Results:* `r total.number.of.patients ` patients were included. Transferred model
   mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage`, 
-  across all samples. The worst performance after transfer was the low-volume model transferred 
-  to the high-volume sample, resulting in a significant mistriage rate increase of `r highest.comp.mistriage`, 
-  primarily due to increased overtriage.
+  across all samples. The worst performance after transfer was the low-volume model 
+  transferred to the high-volume sample, resulting in a significant mistriage rate 
+  increase of `r highest.comp.mistriage`.
 
-  *Conclusions:* In all samples, model transfer was associated with some degree of uncertainty. 
-  Models developed in samples with high numbers of patients performed best after transfer, and the
-  model developed in the low-volume sample performed worst after transfer.
+  *Conclusions:* In all samples, model transfer was associated with some uncertainty. 
+  Models developed in samples with high numbers of patients generally performed best after 
+  transfer, and the model developed in the low-volume sample performed worst after transfer.
 

### 2019-10-09 ManuscriptMarkdownAJE.Rmd
Update and rename ManuscriptMarkdownAnon.Rmd to ManuscriptMarkdownAJE.Rmd

--- 2019-10-08 ManuscriptMarkdownAnon.Rmd
@@
 title: "Transferability of clinical prediction models for early trauma care"
-date: "October 08, 2019"
+date: "October 09, 2019"
 documentclass: "article"
@@
   *Background:* Trauma results in approximately four million deaths annually. Clinical
-  prediction models are often used for triage in early trauma care. Transferring these 
-  models between different settings might adversely affect model performance.
+  prediction models are often used for triage in early trauma care. This study aimed to 
+  assess how transfers of clinical prediction models for early trauma 
+  care between different care contexts within a single health system affected mistriage rates.
 
@@
   prediction models were developed using logistic regression, and transferred within sets. 
-  Model performance was evaluated in all data sets using the mistriage rate, undertriage 
+  Model performance was evaluated in all data sets using mistriage rate, undertriage 
   rate and overtriage rate. Model performance was reported as medians with the 95% uncertainty 
@@
   *Results:* `r total.number.of.patients ` patients were included. Transferred model
-  mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage`, 
-  across all samples. The worst performance after transfer was the low-volume model 
-  transferred to the high-volume sample, resulting in a significant mistriage rate 
-  increase of `r highest.comp.mistriage`.
+  mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage`. 
+  Transfer of the low-volume model to the high-volume sample resulted in a significant 
+  mistriage rate increase of `r highest.comp.mistriage`.
 
-  *Conclusions:* In all samples, model transfer was associated with some uncertainty. 
-  Models developed in samples with high numbers of patients generally performed best after 
-  transfer, and the model developed in the low-volume sample performed worst after transfer.
+  *Conclusions:* Models developed in samples with high numbers of patients generally performed best after 
+  transfer, and the model developed in the low-volume sample performed significantly worse after transfer.
 
